### 2023-04-15 paper/main.tex
uploaded paper

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
    \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2018

% ready for submission
% \usepackage{neurips_2018}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
\usepackage[preprint]{nips_2018}

% to compile a camera-ready version, add the [final] option, e.g.:
     %\usepackage[final]{neurips_2018}

% to avoid loading the natbib package, add option nonatbib:
%     \usepackage[nonatbib]{neurips_2018}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{natbib}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage[ruled]{algorithm2e}
\usepackage{bbm}
\usepackage{subcaption}
\usepackage{caption}
\usepackage{array}
\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}

\SetKwInput{kwInit}{Init}
\newtheorem*{remark}{Remark}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newtheorem{theorem}{Theorem}
%\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}

\newcommand{\RSA}{\mathbb{R}^{|\mathcal{S}|\times|\mathcal{A}|}}
\newcommand{\RSAvector}{\mathbb{R}^{|\mathcal{S}||\mathcal{A}| \times 1}}
\newcommand{\RS}{\mathbb{R}^{|\mathcal{S}|}}
\newcommand{\Rvector}[1]{\mathbb{R}^{#1}}
\newcommand{\Rmatrix}[2]{\mathbb{R}^{#1 \times #2}}
\newcommand{\AdaBracket}[1]{\left(#1\right)}
\newcommand{\AdaRectBracket}[1]{\left[#1\right]}
\newcommand{\AdaAngleBracket}[1]{\left\langle#1\right\rangle}
\newcommand{\AdaAngleProduct}[2]{\left\langle#1, #2\right\rangle}
\newcommand{\chiAny}[3]{\chi^{#1}\AdaBracket{#2||#3}}
\newcommand{\expectation}[2]{\mathbb{E}_{#1}\AdaRectBracket{#2}}
\newcommand{\qlog}{$q$-logarithm }
\newcommand{\qexp}{$q$-exponential }
\newcommand{\fdiv}{$f$-divergence }
\newcommand{\qstarlog}{$q^*$-logarithm }
\newcommand{\qstarexp}{$q^*$-exponential }
\newcommand{\tsallis}[1]{S_q(#1)}
\newcommand{\sparse}[1]{S_2(#1)}
\newcommand{\qKL}{D^{q}_{\!K\!L}\!\AdaBracket{\pi || \bar{\pi}}}
\newcommand{\qKLindex}[2]{D^{q}_{\!K\!L}\!\AdaBracket{\pi_{#1}|| \pi_{#2}}}
\newcommand{\qstarKLindex}[2]{D^{q^*}_{\!K\!L}\!\AdaBracket{\pi_{#1}|| \pi_{#2}}}
\newcommand{\qKLany}[2]{D^{q}_{\!K\!L}\!\left(#1 \left|  \right| #2 \right)}
\newcommand{\qKLstarany}[2]{D^{q^*}_{\!K\!L}\!\left(#1 \left|  \right| #2 \right)}
\newcommand{\KL}{D_{\!K\!L}\!\AdaBracket{\pi || \bar{\pi}}}
\newcommand{\KLindex}[2]{D_{\!K\!L}\!\AdaBracket{\pi_{#1}|| \pi_{#2}}}
\newcommand{\KLany}[2]{D_{\!K\!L}\!\left(#1 \left|  \right| #2 \right)}
\newcommand{\entropy}{\mathcal{H}\left( \pi \right)}
\newcommand{\entropyany}[1]{\mathcal{H}\left( #1 \right)}
\newcommand{\entropyIndex}[1]{\mathcal{H}\left( \pi_{#1} \right)}
\newcommand{\fdivany}[2]{D_{f}\!\left(#1 \left|  \right| #2 \right)}
\newcommand{\logq}[1]{\ln_{q}\!#1}
\newcommand{\logstarq}[1]{\ln_{q^*}\!#1}
\newcommand{\expq}[1]{\exp_{q}\!#1}
\newcommand{\expstarq}[1]{\exp_{q^*}\!#1}
\newcommand{\eq}[1]{Eq.\,(#1)}
\newcommand{\taulogqratio}[3]{\mathbb{M}_{#1,#2} + \ln_{#3}{\pi}_{#1}}
\newcommand{\uniformA}{\left|\mathcal{A}\right|}
\newcommand{\normalization}[1]{\tau\psi\AdaBracket{ \frac{Q_{#1}}{\tau}}}
\newcommand{\renyi}{Rényi }
\newcommand{\renyidivany}[2]{D_\text{\renyi}\!\!\AdaBracket{ #1 || #2}}
\newcommand{\transition}{\gamma P_{\pi}}
\newcommand{\greedy}[1]{\argmax_{\pi}\!\AdaAngleProduct{\pi}{#1}}
\newcommand{\greedyH}[1]{\argmax_{\pi}\!\AdaAngleProduct{\pi}{#1 - \tau\ln\pi}}
\newcommand{\greedyOmega}[1]{\argmax_{\pi}\!\AdaBracket{\AdaAngleProduct{\pi}{#1} - \tau\Omega(\pi)}}
\newcommand{\greedyTKL}[1]{\argmax_{\pi}\!\AdaAngleProduct{\pi}{#1 - \ln_{{q}}\!\frac{\pi}{\pi_k}}}
\newcommand{\BellmanKL}[2]{T^{\lambda,0}_{\pi_{#1}|\pi_{#2}}}
\newcommand{\BellmanKLGreek}[3]{T^{#1,0}_{\pi_{#2}|\pi_{#3}}}
\newcommand{\BellmanKLbar}{T^{\lambda, 0}_{\pi|\bar{\pi}}}
\newcommand{\BellmanH}[2]{r + \gamma P\!\AdaAngleProduct{\pi_{#1}}{Q_{#2} -
\tau\ln\pi_{#1}}}
\newcommand{\BellmanOmega}[2]{r + \gamma P\!\AdaBracket{ \AdaAngleProduct{\pi_{#1}}{Q_{#2}} - \tau\Omega(\pi_{#1})}}
\newcommand{\BellmanTKL}[2]{r + \gamma P\!\AdaAngleProduct{\pi_{#1}}{\TsallisQ{#2} - \ln_{q}\!\frac{\pi_{#1}}{\pi_{#2}}}}
\newcommand{\BellmanImplicit}[2]{\tilde{T}^{0,#1}_{\pi_{#2}}}
\newcommand{\BellmanBoth}[4]{{T}^{#1,#2}_{\pi_{#3}|\pi_{#4}}}
\newcommand{\BellmanG}[1]{T^{\alpha}_{\pi_{#1}}}
\newcommand{\Bellman}[1]{r + \gamma P_* Q_{#1} }
\newcommand{\sumJK}[1]{\sum_{j=#1}^{k}}
\newcommand{\datasetQ}{Q_{\pi_{\mathcal{D}}}}
\newcommand{\datasetOptimalQ}{Q_{*, \pi_{\mathcal{D}}}}
\newcommand{\datasetPolicy}{\pi_{\mathcal{D}}}

\title{Offline Reinforcement Learning with Tsallis Regularization}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{%
  Department of Computing Science\\
  Alberta University\\
  Pittsburgh, PA 15213 \\
  \texttt{hippo@cs.cranberry-lemon.edu} \\
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}
    We show that the Tsallis regularization is well suited to offline reinforcement learning, where its main drawback insufficient exploration vanishes.
    Furthermore, Tsallis entropy induces sparsemax policies which have a natural interpretation of regarding actions not present in the dataset as the truncated actions, hence no support constraint as is done in In-Sample softmax is necessary.
\end{abstract}

\section{Introduction}

\section{Background}

We focus on discrete-time discounted Markov Decision Processes (MDPs) expressed by the tuple $(\mathcal{S}, \mathcal{A}, P,  r,  \gamma)$, where $\mathcal{S}$ and $\mathcal{A}$ denote state space and finite action space, respectively. 
$d(\cdot)$ denotes the initial state distribution.
$P(\cdot |s,a)$ denotes transition probability over the state space given state-action pair $(s,a)$,  
and $r(s,a)$ defines the reward associated with that transition. 
When time step $t$ is of concern, we write $r_t := r(s_t, a_t)$.
$\gamma \in (0,1)$ is the discount factor.
A policy $\pi(\cdot|s)$ is a mapping from the state space to distributions over actions.
% The stationary state distribution induced by $\pi$ can then be defined as $d^{\pi}_{\mu}(s)\!:=\!(1-\gamma)\sum^{\infty}_{t=0}\gamma^{t}P(s_t=s|\pi, \mu)$.
We define the state value function at a specific state $s$ following policy $\pi$  as  $V_{\pi}(s) = \mathbb{E}\AdaRectBracket{\sum_{t=0}^{\infty} r_t | s }$.
Likewise, we define the state-action value function given initial action $a_0$ as  $Q_{\pi}(s,a) = \mathbb{E}\AdaRectBracket{\sum_{t=0}^{\infty} r_t | s_0=s, a_0 = a }$,
% \begin{align}
%     Q_{\pi}(s,a) = \mathbb{E}\AdaRectBracket{\sum_{t=0}^{\infty} r_t | s_0=s, a_0 = a } ,
% \end{align}
where the expectation is with respect to the policy $\pi$ and transition probability $P$.

A standard approach to find the optimal value function $Q_*$ is value iteration. To define the formulas for value iteration, it will be convenient to write these functions as vectors, $V_{\pi} \in \RS$ and $Q_{\pi} \!\in\! \RSA$. 
For notational convenience, we define the inner product for any two functions $F_1,F_2\in\RSA$  as $\AdaAngleProduct{F_1}{F_2} \in \RS$, where we only take the inner product over actions. 
% MARTHA: mention this when it is used
%$\boldsymbol{1}$ denotes an all-one vector whose dimension should be clear from the context.
The Bellman operator acting upon any function $Q \!\in\! \RSA$ can be defined as: $T_{\pi}Q := r + \transition Q$, where $P_{\pi}Q \!\in\! \RSA \!:=\! \mathbb{E}_{s'\sim P(\cdot|s,a),a'\sim\pi(\cdot|s')}\!\AdaRectBracket{Q(s',a')} \!=\! P\!\AdaAngleProduct{\pi}{Q}$.
When $\pi$ is greedy with respect to $Q$, we have the Bellman optimality operator defined by $T_*Q := r + \gamma P_* Q$, $P_*Q \!=\! \mathbb{E}_{s'\sim P(\cdot|s,a)}\!\AdaRectBracket{\max_{a'}Q(s',a')}$. 
The above definitions should be understood as component-wise.
Repeatedly applying the Bellman operator $T_{\pi}Q$ converges to the unique fixed point $Q_{\pi}$, and $T_*Q$ it converges to the optimal action value function $Q_*:=Q_{\pi^*}$.
% Typically the Bellman operator is applied $m \geq 1$ times, and
% $m=1,\infty$ corresponds value iteration and policy iteration, respectively.
% $1\!<\! m < \!\infty\!$ implies the use of approximate modified policy iteration \citep{scherrer15-AMPI}.
This basic recursion can be modified with the addition of a regularizer $\Omega(\pi)$:
\begin{align}
    \begin{cases}
        \pi_{k+1} = \greedy{Q_{k}}, & \\
        Q_{k+1} =  \AdaBracket{ T_{\pi_{k+1}} Q_k }^m , &
    \end{cases}
    \label{eq:PI}
\quad
    \begin{cases}
        \pi_{k+1} = \greedyOmega{Q_{k}}, & \\
        Q_{k+1} = \AdaBracket{T_{\pi_{k+1, \Omega}}Q_k}^m &
    \end{cases}
\end{align}
% MARTHA: What? why would m = infty be policy iteration? this formulas has T_*
%
% \begin{align}
%   \begin{cases}
%       \pi_{k+1} = \greedyOmega{Q_{k}}, & \\
%       Q_{k+1} = \AdaBracket{T_{\pi_{k+1, \Omega}}Q_k}^m &
%   \end{cases}
%   \label{eq:regularizedPI}
% \end{align}
where $T_{\pi_{k+1}, \Omega}Q_k \!=\! \BellmanOmega{k+1}{k}$ is the regularized Bellman operator \citep{geist19-regularized}. This modified recursion is guaranteed to converge if $\Omega$ is concave in $\pi$. For standard (Shannon) entropy regularization, we use $\Omega(\pi) = -\entropy = \AdaAngleProduct{\pi}{\ln\pi}$. The resulting optimal policy has $\pi_{k+1} \propto \exp\AdaBracket{\tau^{-1}Q_k}$, where $\propto$ indicates \emph{proportional to} up to a constant not depending on actions.

Another popular choice is KL divergence $\Omega(\pi)=\KLany{\pi}{\mu} = \AdaAngleProduct{\pi}{\ln{\pi} \!-\!\ln{\mu}}$, which is more general since we recover Shannon entropy when we choose $\mu$ to be a uniform distribution, i.e. $\frac{1}{|\mathcal{A}|}$. In this work, when we say KL regularization, we mean the standard choice of setting $\mu = \pi_k$, the estimate from the previous update.
Therefore, $D_\text{KL}$ serves as a penalty to penalize aggressive policy changes.
The optimal policy in this case takes the form $\pi_{k+1}\propto\pi_k\exp\AdaBracket{\tau^{-1} Q_k}$.
By induction, we can show this KL-regularized optimal policy $\pi_{k+1}$ is a softmax over a uniform average over the history of action value estimates \cite{vieillard2020leverage}:
\begin{align}
  \pi_{k+1} \propto \pi_k \exp\AdaBracket{\tau^{-1} Q_k}  \propto \cdots \propto \exp\AdaBracket{\! \tau^{-1}\sumJK{1}Q_j}
  .
  \label{eq:average}
\end{align}
Using KL regularization has been shown to be theoretically superior to entropy regularization, in terms of error tolerance \cite{azar2012dynamic,vieillard2020leverage,Kozuno-2022-KLGenerativeMinMaxOptimal,chan2021-greedification}.

The definitions of $\mathcal{H}(\cdot)$ and $D_\text{KL}(\cdot||\cdot)$ rely on the standard logarithm and its inverse (the exponential) and both induce softmax policies as an exponential over (weighted) action-values \citep{hiriart2004-convexanalysis,Nachum2020-RLFenchelRockafellar}. 
Convergence properties of the resulting regularized algorithms have been well studied \citep{kozunoCVI,geist19-regularized,vieillard2020leverage}.
In this paper, we investigate Tsallis entropy and Tsallis KL divergence as the regularizer, which generalize Shannon entropy and KL divergence respectively.

\section{$q$-statistics and Tsallis regularization}

A more general form of entropy known as Tsallis entropy is proposed by \cite{TsallisEntropy},  defined based on the \qlog that is referred to as \emph{deformed logarithm} \cite{tsallis2009introduction}.
Let us revisit the definition of \qlog its unique inverse function $q$-exponential (analogous to the exponential function).
We can define the Tsallis entropy $S_q(\pi)$ in a similar way to Shannon entropy with \qlog:
\begin{align}
    \begin{split}
        q \in  \mathbb{R}, \quad \logq{x} = \frac{x^{q-1} - 1}{q-1}, \quad \expq{x} = \AdaRectBracket{1 + (q - 1)x}^{\frac{1}{q-1}}_{+},\quad \tsallis{\pi} = -\AdaAngleProduct{\pi}{\logq{\pi}}.
    \end{split}
\end{align}
When $q=2$, we recover the Tsallis sparse entropy \cite{Martins16-sparsemax}, which achieves the sparsity by projecting onto the probability simplex \cite{Blondel-2020LearningFenchelYoundLoss}.
In RL, it has been investigated by \cite{Lee2018-TsallisRAL,Lee2020-generalTsallisRSS} and it is known that the optimal regularized policy takes the form:
\begin{align*}
    \pi_{k+1}(a|s) = &\AdaRectBracket{\frac{Q_k(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_k(s, \cdot)}{\tau}}}_{+} = \exp_2 \AdaBracket{\frac{Q_k(s,a)}{\tau q} - \psi\AdaBracket{\frac{Q_k(s, \cdot)}{\tau q}}}, \\
    &\qquad \qquad \psi\AdaBracket{\frac{Q_{\pi}(s,\cdot)}{\tau}} \doteq \frac{\sum_{a\in K(s)} \frac{Q(s,a)}{\tau} - 1 }{|K{(s)}|},
\end{align*}
where $[\cdot]_{+} = \max \{\cdot, 0\}$,  $K(s)$ is the set of highest-value actions satisfying $1 \!+\! i\frac{Q(s,a_{(i)})}{\tau} \!>\! \sum_{j=1}^{i}\frac{Q(s,a_{(j)})}{\tau}$., with $a_{(j)}$ denotes the $j$-th largest action.

Very recently \citet{zhu2023generalized} proposed in RL to exploit Tsallis KL divergence which generalizes Tsallis entropy.  
They proved that the regularized optimal policy takes a similar form to \eq{\ref{eq:average}}:
% Recall that, this definition is different from the statistical physics definition \citep{Furuichi2004-fundamentals-qKL}:
\begin{align}
    \begin{split}
        \qKLany{\pi}{\mu} = \AdaAngleProduct{\pi}{\logq{\frac{\pi}{\mu}}}, \quad \pi_{k+1} = \pi_{k} \exp_q\AdaBracket{\frac{Q_k}{\tau} - \psi\AdaBracket{\frac{Q_k}{\tau}} }.
    \end{split}
    \label{eq:tkl_policy}
\end{align}
\eq{\ref{eq:tkl_policy}} is important to our application of offline RL where we show the in-sample softmax becomes in-sample sparsemax.
{\color{red} (going to put some figures here to show $q$-statistics behavior and illustrate how the sparsemax policies truncate actions.)}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{img/policy_truncate.png}
    \caption{The sparsemax operator acting upon Gaussian and Boltzmann policies by truncating actions with values lower than $\psi$. 
    The truncated actions can be seen as not present in the offline dataset.
    }
\end{figure}

\section{In-Sample Softmax for Offline RL}

To alleviate the out-of-distribution error, \citet{Fujimoto2019-InSampleMax} proposed the in-sample Bellman optimality equation to update only for actions present in the dataset:
\begin{align}
    \datasetOptimalQ = r + \gamma P_{*, \datasetPolicy} \datasetOptimalQ = r(s,a) + \gamma \expectation{s'\sim P(\cdot | s,a)}{\max_{a': \datasetPolicy(a'|s') > 0} \datasetOptimalQ(s',a')}.
\end{align}
Later, \citet{Xiao2023-InSampleSoftmax} proposed in-sample softmax to better estimate the policy inside the bracket since in the continuous case the hard max operator might be difficult to solve for.
By adding negative entropy regularization $\Omega(\pi) = \tau \entropy$ and imposing the data support constraint, the in-sample softmax Bellman optimality equation has the following evaluation step:
\begin{align}
    \datasetOptimalQ = r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \sum_{a': \datasetPolicy(a'|s') > 0} \!\!\!\!\! \pi(a'|s') \AdaBracket{\datasetOptimalQ(s',a') - \tau\entropy(\cdot|s')}}.
    \label{eq:insample_softmax}
\end{align}
By choosing the maximizer policy the above equation takes the form:
\begin{align}
    \datasetOptimalQ = r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{\tau \ln \!\!\!\!\!\sum_{a': \datasetPolicy(a'|s') > 0} \!\!\!\!\! \exp\AdaBracket{\tau^{-1} \datasetOptimalQ(s',a')}}.
\end{align}
As the dataset support constraint poses a challenge to implementation, \citet{Xiao2023-InSampleSoftmax} proposed to transform the summation into an expectation to avoid directly computing the constraint: 
\begin{align}
    \begin{split}
    \sum_{a': \datasetPolicy(a'|s') > 0} \!\!\!\!\!\!\exp\AdaBracket{\tau^{-1}\datasetOptimalQ (s',a')} &= \!\!\!\!\!\!\sum_{a': \datasetPolicy(a'|s') > 0} \frac{\datasetPolicy(a'|s')}{\datasetPolicy(a'|s')} \exp\AdaBracket{\tau^{-1}\datasetOptimalQ (s',a')} \\
    &= \expectation{a'\sim\datasetPolicy(\cdot | s')}{\exp\AdaBracket{\tau^{-1}\datasetOptimalQ (s',a') - \ln\datasetPolicy(a'|s')}}.
    \end{split}
\end{align}
The expectation is approximated by Monte-Carlo sampling actions from the dataset.
Since the term $\exp\AdaBracket{\tau^{-1}\datasetOptimalQ (s',a')}$ appears also in the regularized softmax policy, in-sample softmax updates the policy towards 
\begin{align}
    \pi_{\datasetPolicy, k+1} \propto  \datasetPolicy(a|s) \exp\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,a)}{\tau} - \ln\hat{\pi}_{\mathcal{D}}(a|s)},
    \label{eq:inac_policy}
\end{align}
where $\hat{\pi}_{\mathcal{D}}$ inside the exponential function is learned to imitate the behavior policy to avoid $\pi_{\mathcal{D}} = 0$ leading to an unbounded log-policy.

We interpret \eq{\ref{eq:inac_policy}} from another perspective.
the in-sample softmax policy can be decomposed into two terms:
the first term $\datasetPolicy(a|s) \exp\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,a)}{\tau} }$ acts as a KL-regularized policy with respect to the behavior policy (see \eq{\ref{eq:average}}); 
the second term $\exp\AdaBracket{-\ln\hat{\pi}_{\mathcal{D}}(a|s)}$ can be seen as induced by another regularization $\Omega(\pi) = -\AdaAngleProduct{\pi}{\ln{\hat{\pi}_{\mathcal{D}}}}$, the cross entropy between the in-sample softmax policy and the behavior policy.

\section{Tsallis Regularized Policy for Offline RL}

The key to both in-sample max \cite{Fujimoto2019-InSampleMax} and in-sample softmax \cite{Xiao2023-InSampleSoftmax} is the dataset support constraint.
In this paper, we propose to exploit the properties of sparsemax policies to naturally satisfy this constraint.
In a nutshell, we assume the actions not present in the dataset are with low probability and are exactly these actions that would have been truncated if by an application of the sparsemax operator.

Let us recall the definition of the sparsemax policy takes the form $\pi_{k+1}(a|s) = \AdaRectBracket{\frac{Q_k(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_k(s, \cdot)}{\tau}}}_{+}$. 
That is, actions with values lower than $\psi$ are set to zero probability;
actions with values greater than $\psi$ are included in the set $K(s)$.
If we follow the assumption that the actions present in the dataset are these actions in $K(s)$, we can replace the dataset support constraint $a:\datasetPolicy(a|s) > 0$ by the constraint $a\in K(s)$.

Let us write out the Bellman optimality equation similar to \eq{\ref{eq:insample_softmax}}, but replacing $\entropy$ to the Tsallis sparse entropy $S_2(\pi)$:
\begin{align}
    \begin{split}
    \datasetOptimalQ &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \sum_{a'} \pi(a'|s') \AdaBracket{\datasetOptimalQ(s',a') - \tau S_2(\pi)(\cdot|s')}}. \\ 
    &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \frac{1}{2}\tau \AdaBracket{ \sum_{a' \in K(s)} \AdaBracket{\frac{\datasetOptimalQ(s',a')}{\tau}}^2 - \psi \AdaBracket{\frac{\datasetOptimalQ(s', \cdot)}{\tau}}^2 + 1 }}\\
    &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \frac{1}{2}\tau \AdaBracket{ \sum_{a': \datasetPolicy(a'|s') > 0} \AdaBracket{\frac{\datasetOptimalQ(s',a')}{\tau}}^2 - \psi \AdaBracket{\frac{\datasetOptimalQ(s', \cdot)}{\tau}}^2 + 1 }}.
    \label{eq:insample_sparsemax}
    \end{split}
\end{align}
Here, $\max_{\pi \in \Delta(\mathcal{A})}\pi(a|s)Q(s,a) + \tau S_2(\pi(\cdot|s)) $ attains its maximum at $V(s) = \frac{\tau}{2} \sum_{a \in K(s)}\AdaBracket{ \frac{Q(s,a)}{\tau} }^2 - \psi\AdaBracket{\frac{Q(s, \cdot)}{\tau}}^2 + \frac{\tau}{2}$ \cite{Lee2018-TsallisRAL,Martins16-sparsemax}.
Since we assume the conditions $a:\datasetPolicy(a|s) > 0$ by the constraint $a\in K(s)$ are interchangeable, the dataset support constraint is implicitly satisfied by the Tsallis sparse entropy regularization.
{\color{red}(we could add something here to say actually for all $q\geq 2$ the policies are variants of sparsemax, but not sure this is needed or not.)}

\begin{remark}
    Tsallis entropy regularization has not been popular since its proposal in RL \cite{Lee2018-TsallisRAL}.
    One of the main reasons is the sparsemax policies are not suitable for online RL since the exploration is handicapped resulted from the action truncation.
    However, in offline RL this drawback vanishes, and theoretically it provides a better in-sample algorithm that the constraint is satisfied implicitly. 
\end{remark}

\subsection{Implementation}

Let $\theta, \phi, \omega$ denote the parametrization of networks for $Q$, $\pi_{\datasetPolicy}, \datasetPolicy$, respectively.
In-sample softmax updates the policy towards 
\begin{align}
    \pi_{\datasetPolicy, Q_{\theta}}(a|s) = \datasetPolicy(a|s)\exp\AdaBracket{\frac{Q_{\theta}(s,a) - Z(s)}{\tau} - \ln \pi_{\omega}(a|s)},
    \label{eq:insample_policy}
\end{align}
where $Z(s)$ denotes the normalization constant and is necessary since the policy is updated by minimizing KL divergence:
\begin{align*}
    &\KLany{\pi_{\datasetPolicy, Q_{\theta}}(\cdot|s)}{\pi_{\phi}(\cdot | s)} = \expectation{a \sim \pi_{\datasetPolicy, Q_{\theta}}(\cdot|s) }{\ln \pi_{\datasetPolicy, Q_{\theta}}(a|s) - \ln \pi_{\phi}(a | s)} \\
    &= \expectation{a \sim \datasetPolicy(\cdot|s)}{ - \exp\AdaBracket{\frac{Q_{\theta}(s,a) - Z(s)}{\tau} - \ln \pi_{\omega}(a|s)} \ln\pi_{\phi}(a|s) },
\end{align*} 
where the $\datasetPolicy$ term in $ \pi_{\datasetPolicy, Q_{\theta}}$ is absorbed into the expectation, so the KL divergence loss can be minimized by sampling actions from the offline dataset.

We follow the same setup here, but replacing every appearance of $\ln, \exp$ to their \qlog and $q$-exponential counterpart for $q=2$.
In order to also sample actions from the dataset when updating the policy, we repeat the derivation in \eq{\ref{eq:insample_policy}} but for the sparsemax policy.
Let us define $z_{\datasetPolicy, k, \tau}(s,a) := \frac{Q_{\datasetPolicy, k}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,\cdot)}{\tau}}$ for convenience:
\begin{align}
    \begin{split}
        \pi_{\datasetPolicy, k+1} (a|s) &= \datasetPolicy(a|s) \datasetPolicy(a|s)^{-1} \exp_2{z_{\datasetPolicy, k, \tau}(s,a)} \\
        &= \datasetPolicy(a|s)  \exp_2{\AdaBracket{\ln_2{\frac{1}{\datasetPolicy(a|s)}}}} \exp_2{z_{\datasetPolicy, k, \tau}(s,a)}\\
        &=  \datasetPolicy(a|s) \AdaBracket{ \exp_2\AdaBracket{z_{\datasetPolicy, k, \tau}(s,a) + \ln_2\frac{1}{\datasetPolicy(a|s)}} -  z_{\datasetPolicy, k, \tau}(s,a)  \ln_2{\frac{1}{\datasetPolicy(a|s)}} }.
        \label{eq:tsallis_inac_policy}
    \end{split}
\end{align}
In the last step we made use of the relationship $\AdaBracket{\expq{x}\cdot \expq{y}}^{q-1} = \expq{\AdaBracket{x+y}}^{q-1} + (q-1)^2 xy$ \cite{Yamano2004-properties-qlogexp}.

In the discrete case, the truncation threshold or the normalization function $\psi$ is computed by sorting the action values.
However, in the continuous case we can no longer do that.
Instead, we propose to use a single-point estimate:
\begin{align*}
    &V(s) \approx \frac{1}{2}\tau \AdaBracket{ \AdaBracket{\frac{Q_{\datasetPolicy, k}(s',a')}{\tau}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy, k}(s', \cdot)}{\tau}}^2 + 1} \\
    &\Leftrightarrow \psi\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,\cdot)}{\tau}} \approx \sqrt{\AdaBracket{\frac{Q_{\datasetPolicy, k}(s',a')}{\tau}}^2 - \frac{V(s) - \frac{1}{2}\tau}{\frac{1}{2}\tau}}.
\end{align*}
{\color{red}(need to record this $\psi$ value in the experiment to see how big area of the action Gaussian distribution is truncated.)}
The single-point estimate is similar to the common practice of entropy-regularized actor-critic algorithms such as soft actor-critic \cite{haarnoja-SAC2018} that approximates the value function by a single-point estimate, i.e. $V_{\zeta}(s) \approx Q_{\theta}(s,a) + \ln\pi_{\phi}(a|s)$.

We now summarize the loss functions for implementing the Tsallis in-sample actor-critic algorithm:
\begin{align}
    & \psi\AdaBracket{\frac{Q_{\datasetPolicy, \theta}(s,\cdot)}{\tau}} \leftarrow \sqrt{\AdaBracket{\frac{Q_{\theta}(s,a)}{\tau}}^2 - \frac{V_{\zeta}(s) - \frac{\tau}{2}}{\frac{\tau}{2}} }, \\ 
    & z_{\datasetPolicy, \theta, \tau}(s,a) \leftarrow \frac{Q_{\theta}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\theta}(s,\cdot)}{\tau}},\\
    &\mathcal{L}_\text{actor}(\phi) = -\expectation{s,a \sim \mathcal{D}}{ \AdaBracket{ \exp_2\AdaBracket{z_{\datasetPolicy, \theta, \tau}(s,a) + \ln_2\frac{1}{\pi_{\omega}(a|s)}} -  z_{\datasetPolicy, \theta, \tau}(s,a)  \logq{\frac{1}{\pi_{\omega}(a|s)}} } \ln\pi_{\phi}(a|s) }, \\
    & \mathcal{L}_\text{baseline}(\zeta) = \expectation{s\sim\mathcal{D}, a\sim \pi_{\phi}(s)}{\AdaBracket{ v_{\zeta}(s) - Q_{\theta}(s,a) - \tau\ln_2\pi_{\phi}(a|s)}^2},\\
    & \mathcal{L}_\text{critic}(\theta) = \expectation{s,a,r,s' \sim \mathcal{D}}{\AdaBracket{r + \gamma v_{\zeta}(s') - Q_{\theta}(s,a)}^2 }.
\end{align}

\clearpage

\bibliographystyle{abbrvnat}
\bibliography{library}

\end{document}

### 2023-04-15 parallel/run_config_array.sh
Making some mods, working towards getting experiments up and running again.

@@ -16,12 +16,20 @@
 
 
 source /project/6010404/mkschleg/tsallis_inac/TSALLIS_INAC/bin/activate
+### USE
+# Need to specify job name, array parameters, and number of tasks per node (split so we can fit a good amount of jobs"
+# sbatch -J <job_name> --array=0-10 run_config_array.sh 
+### INPUT
+# $1 = configuration file
+# $2 = base_save_dir
 
-mkdir "parallel-logs/$(SLURM_JOB_NAME)"
+source /project/6010404/mkschleg/tsallis_inac/TSALLIS_INAC/bin/activate
 
-JOB_LOG="parallel-logs/$(SLURM_JOB_NAME)/$(SLURM_ARRAY_TASK_ID).log"
-PARALLEL_CONFIGS = "--delay .2 -j $(SLURM_NTASKS) -- joblog $(JOB_LOG)"
-SRUN_CONFIGS = "-N 1 -n 1 --exclusive" # 1 CPU per job?
+mkdir "parallel-logs/${SLURM_JOB_NAME}"
+
+JOB_LOG="parallel-logs/${SLURM_JOB_NAME}/${SLURM_ARRAY_TASK_ID}.log"
+PARALLEL_CONFIGS="--delay .2 -j ${SLURM_NTASKS} -- joblog ${JOB_LOG}"
+SRUN_CONFIGS="-N 1 -n 1 --exclusive" # 1 CPU per job?
 
 output_str=`python run_from_config.py --id 0 --config $1 --base_save_dir ~/tmp --get_num_jobs`
 NUM_JOBS=`echo "$output_str" | tail -n1`
@@ -41,4 +49,5 @@ echo "NUM TOTAL JOBS: $NUM_JOBS"
 echo "ARRAY TASK: $SLURM_ARRAY_TASK_ID"
 echo "SEQUENCE RAN: $SEQUENCE"
 
-echo $PARALLEL_CONFIGS srun $SRUN_CONFIGS python run_from_config.py --id {1} --config $CONFIG_FILE --base_save_dir  ::: $SEQUENCE
+parallel $PARALLEL_CONFIGS srun $SRUN_CONFIGS python run_from_config.py --id {1} --config $CONFIG_FILE --base_save_dir $2  ::: $SEQUENCE
+

### 2023-04-15 run_from_config.py

@@ -43,7 +43,7 @@ def run_experiment(config_file, job_id, base_save_dir, num_threads):
     random.seed(cfg["run"])
     seed = random.randint(1, 1000000000)
     cfg.set_seed(seed)
-    torch_utils.random_seed(cfg["run"])
+    torch_utils.random_seed(cfg["seed"])
 
     # Save Path
     exp_path = cfg.get_save_dir_and_save_config(

### 2023-04-17 parallel/run_config_array.sh
Fixxing job scripts from CC

@@ -5,8 +5,6 @@
 #SBATCH --signal=USR1@90
 #SBATCH -o job_out/%x_%a.out # Standard output
 #SBATCH -e job_out/%x_%a.err # Standard error
-#SBATCH --nodes=1   # Get a whole node for the array.
-#SBATCH --mem=0     # Total memory for all tasks
 
 ### USE
 # Need to specify job name, array parameters, and number of tasks per node (split so we can fit a good amount of jobs"
@@ -28,7 +26,7 @@ source /project/6010404/mkschleg/tsallis_inac/TSALLIS_INAC/bin/activate
 mkdir "parallel-logs/${SLURM_JOB_NAME}"
 
 JOB_LOG="parallel-logs/${SLURM_JOB_NAME}/${SLURM_ARRAY_TASK_ID}.log"
-PARALLEL_CONFIGS="--delay .2 -j ${SLURM_NTASKS} -- joblog ${JOB_LOG}"
+PARALLEL_CONFIGS="--delay .2 -j ${SLURM_NTASKS} --joblog ${JOB_LOG}"
 SRUN_CONFIGS="-N 1 -n 1 --exclusive" # 1 CPU per job?
 
 output_str=`python run_from_config.py --id 0 --config $1 --base_save_dir ~/tmp --get_num_jobs`

### 2023-04-17 parallel/run_retry.sh

@@ -1,0 +1,31 @@
+#!/bin/bash
+#SBATCH --account=rrg-whitem
+#SBATCH --mail-user=mkschleg@ualberta.ca
+#SBATCH --mail-type=ALL
+#SBATCH --signal=USR1@90
+#SBATCH -o job_out/%x_retry.out # Standard output
+#SBATCH -e job_out/%x_retry.err # Standard error
+
+### USE
+# Need to specify job name, array parameters, and number of tasks per node (split so we can fit a good amount of jobs"
+# sbatch -J <job_name> --array=0-10 run_config_array.sh 
+### INPUT
+# $1 = configuration file
+
+
+source /project/6010404/mkschleg/tsallis_inac/TSALLIS_INAC/bin/activate
+### USE
+# Need to specify job name, array parameters, and number of tasks per node (split so we can fit a good amount of jobs"
+# sbatch -J <job_name> --array=0-10 run_config_array.sh 
+### INPUT
+# $1 = configuration file
+# $2 = base_save_dir
+
+# mkdir "parallel-logs/${SLURM_JOB_NAME}"
+
+JOB_LOG=$1
+PARALLEL_CONFIGS="--delay .2 -j ${SLURM_NTASKS} --joblog ${JOB_LOG}"
+SRUN_CONFIGS="-N 1 -n 1 --exclusive" # 1 CPU per job?
+
+parallel $PARALLEL_CONFIGS --retry-failed
+

### 2023-04-17 core/agent/tsallis_inac.py
important update: removed Psi, now works on Hopper and Walker

@@ -140,19 +140,28 @@ class TsallisInAC(base.Agent):
         min_Q, _, _ = self.get_q_value(states, actions, with_grad=False)
         '''
         Tsallis, 
-        update towards the policy: pi_D exp_q(Q - Psi) pi_D^{-1}
+        update towards the policy: pi_D exp_q(Q) pi_D^{-1}
         the first pi_D is absorbed into expectation, the rest becomes
         [exp_q( Q - Psi + ln_q(pi_D^-1) )^(q-1) + (q-1)^2 (Q-Psi) ln_q(pi_D^-1)]^(1/(q-1))
-        approximate the normalization Psi by sqrt((Q/tau)^2 - (V-0.5*tau)/0.5*tau)
         '''
         with torch.no_grad():
-            value = self.get_state_value(states)
             beh_prob = torch.clip(torch.exp(self.beh_pi.get_logprob(states, actions)), min=self.eps)
-            Psi = torch.sqrt(torch.clip((min_Q/self.tau)**2 - (value - 0.5*self.tau)/0.5*self.tau, min=0))
+            # value = self.get_state_value(states)
+            # Psi = torch.sqrt(torch.clip((min_Q/self.tau)**2 - (value - 0.5*self.tau)/0.5*self.tau, min=0))
         '''
-        x = Q-Psi, y = ln_q pi_D^-1
+        x = Q, y = ln_q pi_D^-1
         '''
-        x = min_Q/self.tau - Psi
+        
+        # x = min_Q/self.tau - Psi
+        '''why we do not use Psi 
+        we are assuming that actions not in the dataset are truncated 
+        by Tsallis behavior policies. So when presented with these actions in the dataset,
+        there is no need to do a second round of truncation.  
+        Uncomment these lines for values, Psi for a comparison. 
+        They don't learn on more difficult envs like Hopper or Walker
+    
+        '''
+        x = min_Q / self.tau
         y = self.ac.pi.logq_x(beh_prob**(-1), self.q)
         tsallis_policy = torch.pow(self.ac.pi.expq_x(x + y, self.q)**(self.q-1) + (self.q - 1)**2 * x * y, 1/(self.q-1))
         clipped = torch.clip(tsallis_policy, self.eps, self.exp_threshold)

### 2023-04-17 core/agent/tsallis_inac.py
important update: removed Psi, now works on Hopper and Walker

@@ -148,19 +148,18 @@ class TsallisInAC(base.Agent):
             beh_prob = torch.clip(torch.exp(self.beh_pi.get_logprob(states, actions)), min=self.eps)
             # value = self.get_state_value(states)
             # Psi = torch.sqrt(torch.clip((min_Q/self.tau)**2 - (value - 0.5*self.tau)/0.5*self.tau, min=0))
-        '''
-        x = Q, y = ln_q pi_D^-1
-        '''
-        
         # x = min_Q/self.tau - Psi
         '''why we do not use Psi 
         we are assuming that actions not in the dataset are truncated 
         by Tsallis behavior policies. So when presented with these actions in the dataset,
         there is no need to do a second round of truncation.  
         Uncomment these lines for values, Psi for a comparison. 
-        They don't learn on more difficult envs like Hopper or Walker
-    
+        They don't learn on more difficult envs like Hopper or Walker    
         '''
+        
+        '''
+        x = Q, y = ln_q pi_D^-1
+        '''                
         x = min_Q / self.tau
         y = self.ac.pi.logq_x(beh_prob**(-1), self.q)
         tsallis_policy = torch.pow(self.ac.pi.expq_x(x + y, self.q)**(self.q-1) + (self.q - 1)**2 * x * y, 1/(self.q-1))

### 2023-04-18 paper/main.tex
uploaded the paper
updated the paper

@@
 
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=\textwidth]{img/q_stat_illus.png}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=\textwidth]{img/tsallis_entropy.png}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=\textwidth]{img/tsallis_kl.png}
+    \end{subfigure}
+    \caption{(Left) Behavior of \qlog and $q$-exponential functions.
+    When $q=1$ they respectively recover the standard logarithm and exponential.
+    (Mid) Tsallis entropy of the Gaussian policy $\mathcal{N}(2, 1)$. (Right) Tsallis KL divergence between two Gaussian policies $\mathcal{N}(2.75, 1)$ and $\mathcal{N}(3.25, 1)$.
+    }
+    \label{fig:q_stats}
+\end{figure}
+
 A more general form of entropy known as Tsallis entropy is proposed by \cite{TsallisEntropy},  defined based on the \qlog that is referred to as \emph{deformed logarithm} \cite{tsallis2009introduction}.
-Let us revisit the definition of \qlog its unique inverse function $q$-exponential (analogous to the exponential function).
+Let us revisit the definition of \qlog its unique inverse function $q$-exponential (analogous to the exponential function), see Figure \ref{fig:q_stats}.
 We can define the Tsallis entropy $S_q(\pi)$ in a similar way to Shannon entropy with \qlog:
@@
     \begin{split}
-        q \in  \mathbb{R}, \quad \logq{x} = \frac{x^{q-1} - 1}{q-1}, \quad \expq{x} = \AdaRectBracket{1 + (q - 1)x}^{\frac{1}{q-1}}_{+},\quad \tsallis{\pi} = -\AdaAngleProduct{\pi}{\logq{\pi}}.
+        q \in  \mathbb{R}, \quad \logq{x} = \frac{x^{q-1} - 1}{q-1}, \quad \expq{x} = \AdaRectBracket{1 + (q - 1)x}^{\frac{1}{q-1}}_{+},\quad \tsallis{\pi} = -\frac{1}{2}\AdaAngleProduct{\pi}{\logq{\pi}}.
     \end{split}
@@
 
+% \begin{figure}
+%     \centering
+%     \begin{subfigure}[b]{0.49\textwidth}
+%         \centering
+%         \includegraphics[width=\textwidth]{img/tsallis_entropy.png}
+%         \label{fig:y equals x}
+%     \end{subfigure}
+%     \hfill
+%     \begin{subfigure}[b]{0.49\textwidth}
+%         \centering
+%         \includegraphics[width=\textwidth]{img/tsallis_kl.png}
+%         \label{fig:three sin x}
+%     \end{subfigure}
+%     \caption{Estimation of policy ratio $\frac{\pi_{k+1}}{\pi_k}$ during learning of Munchausen DQN on \texttt{CartPole-v1} and \texttt{LunarLander-v2}.
+%     Notice the $y$-axis is at the magnitude of $10^{35}$.
+%     }
+%     \label{fig:mdqn_policy_ratio}
+% \end{figure}
+
 Very recently \citet{zhu2023generalized} proposed in RL to exploit Tsallis KL divergence which generalizes Tsallis entropy.  
@@
 \eq{\ref{eq:tkl_policy}} is important to our application of offline RL where we show the in-sample softmax becomes in-sample sparsemax.
-{\color{red} (going to put some figures here to show $q$-statistics behavior and illustrate how the sparsemax policies truncate actions.)}
 
-\begin{figure}[h]
+\begin{figure}[t]
     \centering
-    \includegraphics[width=0.75\textwidth]{img/policy_truncate.png}
+    \begin{subfigure}[b]{0.475\textwidth}
+        \includegraphics[width=\textwidth]{img/scaled_Gaussian_sparsemax.png}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.475\textwidth}
+        \includegraphics[width=\textwidth]{img/scaled_Boltzmann_sparsemax.png}
+    \end{subfigure}
     \caption{The sparsemax operator acting upon Gaussian and Boltzmann policies by truncating actions with values lower than $\psi$. 
@@
     \datasetOptimalQ = r + \gamma P_{*, \datasetPolicy} \datasetOptimalQ = r(s,a) + \gamma \expectation{s'\sim P(\cdot | s,a)}{\max_{a': \datasetPolicy(a'|s') > 0} \datasetOptimalQ(s',a')}.
+    \label{eq:hardmax_offline}
 \end{align}
@@
 
-\section{Tsallis Regularized Policy for Offline RL}
+\section{Forward-Backward Tsallis Learning for Offline RL}
+
+Summary:
+\begin{itemize}
+    \item Forward: Tsallis sparsemax policies generate the dataset.
+This assumption is not very strong since Tsallis entropy generalizes Shannon entropy and the behavior policy class is sufficiently rich.
+    \item Backward: learn the optimal policy with the modified Tsallis entropy regularization. The learned policy differs from the conventional sparsemax policies in that no thresholding is required.
+\end{itemize}
 
@@
 In this paper, we propose to exploit the properties of sparsemax policies to naturally satisfy this constraint.
-In a nutshell, we assume the actions not present in the dataset are with low probability and are exactly these actions that would have been truncated if by an application of the sparsemax operator.
-
-Let us recall the definition of the sparsemax policy takes the form $\pi_{k+1}(a|s) = \AdaRectBracket{\frac{Q_k(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_k(s, \cdot)}{\tau}}}_{+}$. 
-That is, actions with values lower than $\psi$ are set to zero probability;
-actions with values greater than $\psi$ are included in the set $K(s)$.
-If we follow the assumption that the actions present in the dataset are these actions in $K(s)$, we can replace the dataset support constraint $a:\datasetPolicy(a|s) > 0$ by the constraint $a\in K(s)$.
-
-Let us write out the Bellman optimality equation similar to \eq{\ref{eq:insample_softmax}}, but replacing $\entropy$ to the Tsallis sparse entropy $S_2(\pi)$:
+In a nutshell, we assume the behavior policies are Tsallis sparsemax policies, and therefore the actions not present in the dataset are with low probability and are exactly these truncated actions.
+Formally, suppose the dataset is generated by:
+\begin{align}
+    \datasetPolicy(a|s) = \exp_2\AdaBracket{\frac{Q_{\datasetPolicy}(s,a)}{\tau_\mathcal{D}} - \psi\AdaBracket{\frac{Q_{\datasetPolicy}(s, \cdot)}{\tau_\mathcal{D}}}},  \quad \sum_{a\in K_\mathcal{D}(s)} \pi_\mathcal{D}(a|s)= 1,
+\end{align}
+where $\tau_\mathcal{D}$ is a unknown coefficient and $K_\mathcal{D}(s)$ denotes the set of allowable actions.
+Now let us inspect the Bellman equation for the behavior policy:
 \begin{align}
     \begin{split}
-    \datasetOptimalQ &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \sum_{a'} \pi(a'|s') \AdaBracket{\datasetOptimalQ(s',a') - \tau S_2(\pi)(\cdot|s')}}. \\ 
-    &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \frac{1}{2}\tau \AdaBracket{ \sum_{a' \in K(s)} \AdaBracket{\frac{\datasetOptimalQ(s',a')}{\tau}}^2 - \psi \AdaBracket{\frac{\datasetOptimalQ(s', \cdot)}{\tau}}^2 + 1 }}\\
-    &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \frac{1}{2}\tau \AdaBracket{ \sum_{a': \datasetPolicy(a'|s') > 0} \AdaBracket{\frac{\datasetOptimalQ(s',a')}{\tau}}^2 - \psi \AdaBracket{\frac{\datasetOptimalQ(s', \cdot)}{\tau}}^2 + 1 }}.
+    Q_{\datasetPolicy}(s,a) &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \sum_{a'} \datasetPolicy(a'|s') \AdaBracket{Q_{\datasetPolicy}(s',a') - \tau_{\mathcal{D}} S_2(\datasetPolicy)(\cdot|s')}}.
+\\
+    &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \frac{1}{2}\tau_{\mathcal{D}} \AdaBracket{ \sum_{a' \in K_{\mathcal{D}}(s')} \AdaBracket{\frac{Q_{\datasetPolicy}(s',a')}{\tau_{\mathcal{D}}}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy}(s', \cdot)}{\tau_{\mathcal{D}}}}^2 + 1 }}\\
+    &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \frac{1}{2}\tau_{\mathcal{D}} \AdaBracket{ \sum_{a': \datasetPolicy(a'|s') > 0} \AdaBracket{\frac{Q_{\datasetPolicy}(s',a')}{\tau_{\mathcal{D}}}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy}(s', \cdot)}{\tau_{\mathcal{D}}}}^2 + 1 }}.
     \label{eq:insample_sparsemax}
@@
 Here, $\max_{\pi \in \Delta(\mathcal{A})}\pi(a|s)Q(s,a) + \tau S_2(\pi(\cdot|s)) $ attains its maximum at $V(s) = \frac{\tau}{2} \sum_{a \in K(s)}\AdaBracket{ \frac{Q(s,a)}{\tau} }^2 - \psi\AdaBracket{\frac{Q(s, \cdot)}{\tau}}^2 + \frac{\tau}{2}$ \cite{Lee2018-TsallisRAL,Martins16-sparsemax}.
-Since we assume the conditions $a:\datasetPolicy(a|s) > 0$ by the constraint $a\in K(s)$ are interchangeable, the dataset support constraint is implicitly satisfied by the Tsallis sparse entropy regularization.
-{\color{red}(we could add something here to say actually for all $q\geq 2$ the policies are variants of sparsemax, but not sure this is needed or not.)}
+Under the Tsallis behavior policy assumption, the dataset support constraint $a: \datasetPolicy(a|s) > 0$ coincides with the condition $a \in K_{\mathcal{D}}(s)$.
+Moreover, the assumption actually assumes a forward process of learning a set of allowable actions $K_\mathcal{D}(s)$ given $Q_{\datasetPolicy}(s,a), \forall s,a$;
+ and it naturally raises the possibility of \emph{backward Tsallis learning}: given the set of allowable actions $K_\mathcal{D}(s)$, learn an optimal policy that resembles $\datasetPolicy$ in the sense of $\pi_* \preceq \datasetPolicy$, i.e.
+putting probability mass only on the actions in $K_\mathcal{D}$.
+
+ The backward Tsallis learning consists in learning $\pi_{t, \datasetPolicy} \preceq \datasetPolicy$ given $K_\mathcal{D}$.
+However, it should be noted that the learned policy no longer requires a thresholding function $\psi$ since $K_\mathcal{D}$ is given, hence no action need to be truncated.
+Specifically, we propose to modify the learned policy form as 
+\begin{align}
+    \pi_{t+1, \datasetPolicy}(a|s) \propto \exp_2\AdaBracket{\frac{Q_{t, \datasetPolicy}(s,a)}{\tau}},  \quad \sum_{a\in K_{\mathcal{D}}(s)} \pi_{t+1, \datasetPolicy} = 1.
+    \label{eq:proposal_policy}
+\end{align}
+The policy is a variant of categorical distributions, which is less susceptible to numerical issues than exponential functions \cite{Tsai2021-selfsupervisedRelativePredictiveCoding}.
+In fact, $\pi_{t+1, \datasetPolicy}$ is a linear function in $Q_{t, \datasetPolicy}$ when $q=2$.
+Another advantage is one no longer needs to compute the thresholding function $\psi$, which requires sorting in the discrete control setting; it remains unclear how to accurately estimate it for the continuous control problems.
+
+% That is, the backward Bellman optimality equation becomes:
+% \begin{align}
+%     \begin{split}
+%         \datasetOptimalQ(s,a) &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \sum_{a' \in K_\mathcal{D}(s')} \pi_{*, \datasetPolicy}(a'|s') \AdaBracket{\datasetOptimalQ(s',a') - \tau S_2(\pi_{*, \datasetPolicy})(\cdot|s')}}.
+%     \end{split}
+% \end{align}
+% The term inside the expectation is now:
+% \begin{align}
+%     \begin{split}
+%         & \sum_{a' \in K_\mathcal{D}(s')} \pi_{*, \datasetPolicy}(a'|s') \AdaBracket{\datasetOptimalQ(s',a') - \tau S_2(\pi_{*, \datasetPolicy})(\cdot|s')} \\ 
+%          &= \sum_{a' \in K_\mathcal{D}(s')} \pi_{*, \datasetPolicy}(a'|s') \AdaBracket{ \datasetOptimalQ(s',a') + \frac{\tau}{2}(1 - \pi_{*, \datasetPolicy}(a'|s')) } \\
+%          &= \sum_{a' \in K_\mathcal{D}(s')} \pi_{*, \datasetPolicy}(a'|s') \AdaBracket{ \datasetOptimalQ(s',a') - \frac{\tau}{2}\pi_{*, \datasetPolicy}(a'|s') } + \frac{\tau}{2} \\
+%          &= \sum_{a' \in K_\mathcal{D}(s')} \frac{\AdaRectBracket{ \tau^{-1}Q_{t, \datasetPolicy}(s,a)}_+ }{\sum_{b\in K_\mathcal{D}(s')}\tau^{-1} Q_{t, \datasetPolicy}(s,b)}
+%     \end{split}
+% \end{align}
+\textbf{Difference with In-sample Softmax. }
+Our proposal is seemingly similar to in-sample softmax \cite{Xiao2023-InSampleSoftmax}.
+However, the two methods are radically different: in-sample softmax requires satisfying the dataset support constraint during learning to alleviate the out-of-distribution error, and is motivated by the observation that taking softmax is easier than the hard max operator \eq{\ref{eq:hardmax_offline}}.
+On the other hand, we assume the behavior policies are Tsallis policies and as such the actions not present in the dataset are those being truncated.
+This viewpoint leads to the natural satisfaction of the dataset support constraint.
+
+Furthermore, the assumption naturally leads to the choice of Tsallis regularized backward learning which recovers in-sample softmax as a special case when $q=1$.
+Since softmax policies have full support, there always exists non-negligible probabilities for suboptimal actions; while for \eq{\ref{eq:proposal_policy}} the optimal policy can switch between greedy policy when $Q_{t, \datasetPolicy}(s,a) \leq 0, \forall a\neq a^*$ and multimodal policy when some actions have positive values.
+
+{\color{red}(state KL divergence result here if available)}
 
@@
 
-\subsection{Implementation}
+\section{Implementation}
 
@@
 In order to also sample actions from the dataset when updating the policy, we repeat the derivation in \eq{\ref{eq:insample_policy}} but for the sparsemax policy.
-Let us define $z_{\datasetPolicy, k, \tau}(s,a) := \frac{Q_{\datasetPolicy, k}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,\cdot)}{\tau}}$ for convenience:
+% Let us define $q_{\theta}(s,a) := \frac{Q_{\datasetPolicy, k}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,\cdot)}{\tau}}$ for convenience:
 \begin{align}
     \begin{split}
-        \pi_{\datasetPolicy, k+1} (a|s) &= \datasetPolicy(a|s) \datasetPolicy(a|s)^{-1} \exp_2{z_{\datasetPolicy, k, \tau}(s,a)} \\
-        &= \datasetPolicy(a|s)  \exp_2{\AdaBracket{\ln_2{\frac{1}{\datasetPolicy(a|s)}}}} \exp_2{z_{\datasetPolicy, k, \tau}(s,a)}\\
-        &=  \datasetPolicy(a|s) \AdaBracket{ \exp_2\AdaBracket{z_{\datasetPolicy, k, \tau}(s,a) + \ln_2\frac{1}{\datasetPolicy(a|s)}} -  z_{\datasetPolicy, k, \tau}(s,a)  \ln_2{\frac{1}{\datasetPolicy(a|s)}} }.
+        \pi_{\datasetPolicy, k+1} (a|s) &= \datasetPolicy(a|s) \datasetPolicy(a|s)^{-1} \exp_2{Q_{\theta}(s,a)} \\
+        &= \datasetPolicy(a|s)  \exp_2{\AdaBracket{\ln_2{\frac{1}{\datasetPolicy(a|s)}}}} \exp_2{Q_{\theta}(s,a)}\\
+        &=  \datasetPolicy(a|s) \AdaBracket{ \exp_2\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\datasetPolicy(a|s)}} -  Q_{\theta}(s,a)  \ln_2{\frac{1}{\datasetPolicy(a|s)}} }.
         \label{eq:tsallis_inac_policy}
@@
 In the last step we made use of the relationship $\AdaBracket{\expq{x}\cdot \expq{y}}^{q-1} = \expq{\AdaBracket{x+y}}^{q-1} + (q-1)^2 xy$ \cite{Yamano2004-properties-qlogexp}.
+Similar to the discussion after \eq{\ref{eq:inac_policy}}, the Tsallis policy we derived here can be seen as the result from Tsallis KL regularization, plus another regularization that gives rise to the additional term inside the $\exp_2$ function.
 
-In the discrete case, the truncation threshold or the normalization function $\psi$ is computed by sorting the action values.
-However, in the continuous case we can no longer do that.
-Instead, we propose to use a single-point estimate:
-\begin{align*}
-    &V(s) \approx \frac{1}{2}\tau \AdaBracket{ \AdaBracket{\frac{Q_{\datasetPolicy, k}(s',a')}{\tau}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy, k}(s', \cdot)}{\tau}}^2 + 1} \\
-    &\Leftrightarrow \psi\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,\cdot)}{\tau}} \approx \sqrt{\AdaBracket{\frac{Q_{\datasetPolicy, k}(s',a')}{\tau}}^2 - \frac{V(s) - \frac{1}{2}\tau}{\frac{1}{2}\tau}}.
-\end{align*}
-{\color{red}(need to record this $\psi$ value in the experiment to see how big area of the action Gaussian distribution is truncated.)}
-The single-point estimate is similar to the common practice of entropy-regularized actor-critic algorithms such as soft actor-critic \cite{haarnoja-SAC2018} that approximates the value function by a single-point estimate, i.e.
+% In the discrete case, the truncation threshold or the normalization function $\psi$ is computed by sorting the action values.
+% However, in the continuous case we can no longer do that.
+% Instead, we propose to use a single-point estimate:
+% \begin{align*}
+%     &V(s) \approx \frac{1}{2}\tau \AdaBracket{ \AdaBracket{\frac{Q_{\datasetPolicy, k}(s',a')}{\tau}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy, k}(s', \cdot)}{\tau}}^2 + 1} \\
+%     &\Leftrightarrow \psi\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,\cdot)}{\tau}} \approx \sqrt{\AdaBracket{\frac{Q_{\datasetPolicy, k}(s',a')}{\tau}}^2 - \frac{V(s) - \frac{1}{2}\tau}{\frac{1}{2}\tau}}.
+% \end{align*}
+% {\color{red}(need to record this $\psi$ value in the experiment to see how big area of the action Gaussian distribution is truncated.)}
+% The single-point estimate is similar to the common practice of entropy-regularized actor-critic algorithms such as soft actor-critic \cite{haarnoja-SAC2018} that approximates the value function by a single-point estimate, i.e.
 $V_{\zeta}(s) \approx Q_{\theta}(s,a) + \ln\pi_{\phi}(a|s)$.
@@
 \begin{align}
-    & \psi\AdaBracket{\frac{Q_{\datasetPolicy, \theta}(s,\cdot)}{\tau}} \leftarrow \sqrt{\AdaBracket{\frac{Q_{\theta}(s,a)}{\tau}}^2 - \frac{V_{\zeta}(s) - \frac{\tau}{2}}{\frac{\tau}{2}} }, \\ 
-    & z_{\datasetPolicy, \theta, \tau}(s,a) \leftarrow \frac{Q_{\theta}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\theta}(s,\cdot)}{\tau}},\\
-    &\mathcal{L}_\text{actor}(\phi) = -\expectation{s,a \sim \mathcal{D}}{ \AdaBracket{ \exp_2\AdaBracket{z_{\datasetPolicy, \theta, \tau}(s,a) + \ln_2\frac{1}{\pi_{\omega}(a|s)}} -  z_{\datasetPolicy, \theta, \tau}(s,a)  \logq{\frac{1}{\pi_{\omega}(a|s)}} } \ln\pi_{\phi}(a|s) }, \\
+    % & \psi\AdaBracket{\frac{Q_{\datasetPolicy, \theta}(s,\cdot)}{\tau}} \leftarrow \sqrt{\AdaBracket{\frac{Q_{\theta}(s,a)}{\tau}}^2 - \frac{V_{\zeta}(s) - \frac{\tau}{2}}{\frac{\tau}{2}} }, \\ 
+    % & Q_{\theta}(s,a) \leftarrow \frac{Q_{\theta}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\theta}(s,\cdot)}{\tau}},\\
+    &\mathcal{L}_\text{actor}(\phi) = -\expectation{s,a \sim \mathcal{D}}{ \AdaBracket{ \exp_2\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\pi_{\omega}(a|s)}} -  Q_{\theta}(s,a)  \logq{\frac{1}{\pi_{\omega}(a|s)}} } \ln\pi_{\phi}(a|s) }, \\
     & \mathcal{L}_\text{baseline}(\zeta) = \expectation{s\sim\mathcal{D}, a\sim \pi_{\phi}(s)}{\AdaBracket{ v_{\zeta}(s) - Q_{\theta}(s,a) - \tau\ln_2\pi_{\phi}(a|s)}^2},\\

### 2023-04-18 paper/main.tex
updated the paper with KL divergence upper bound

@@
     \begin{split}
-        q \in  \mathbb{R}, \quad \logq{x} = \frac{x^{q-1} - 1}{q-1}, \quad \expq{x} = \AdaRectBracket{1 + (q - 1)x}^{\frac{1}{q-1}}_{+},\quad \tsallis{\pi} = -\frac{1}{2}\AdaAngleProduct{\pi}{\logq{\pi}}.
+        q \in  \mathbb{R}, \quad \logq{x} = \frac{x^{q-1} - 1}{q-1}, \quad \expq{x} = \AdaRectBracket{1 + (q - 1)x}^{\frac{1}{q-1}}_{+},\quad \tsallis{\pi} = -\AdaAngleProduct{\pi}{\logq{\pi}}.
     \end{split}
@@
 \begin{align*}
-    \pi_{k+1}(a|s) = &\AdaRectBracket{\frac{Q_k(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_k(s, \cdot)}{\tau}}}_{+} = \exp_2 \AdaBracket{\frac{Q_k(s,a)}{\tau q} - \psi\AdaBracket{\frac{Q_k(s, \cdot)}{\tau q}}}, \\
-    &\qquad \qquad \psi\AdaBracket{\frac{Q_{\pi}(s,\cdot)}{\tau}} \doteq \frac{\sum_{a\in K(s)} \frac{Q(s,a)}{\tau} - 1 }{|K{(s)}|},
+    &\pi_{k+1}(a|s) = \AdaRectBracket{\frac{Q_k(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_k(s, \cdot)}{\tau}}}_{+} = \exp_2 \AdaBracket{\frac{Q_k(s,a)}{\tau } - \tilde{\psi}\AdaBracket{\frac{Q_k(s, \cdot)}{\tau }}}, \\
+    &{\psi}\AdaBracket{\frac{Q_{k}(s,\cdot)}{\tau}} \doteq \frac{\sum_{a\in K(s)} \frac{Q(s,a)}{\tau} - 1 }{|K{(s)}|}, \,\, \tilde{\psi}\AdaBracket{\frac{Q_{k}(s,\cdot)}{\tau}} := \psi\AdaBracket{\frac{Q_{k}(s,\cdot)}{\tau}} + \frac{1}{q-1}
 \end{align*}
@@
     \centering
-    \begin{subfigure}[b]{0.475\textwidth}
-        \includegraphics[width=\textwidth]{img/scaled_Gaussian_sparsemax.png}
+    \begin{subfigure}[b]{0.245\textwidth}
+        \includegraphics[width=\textwidth]{img/q2_Gaussian_sparsemax.png}
     \end{subfigure}
-    \begin{subfigure}[b]{0.475\textwidth}
-        \includegraphics[width=\textwidth]{img/scaled_Boltzmann_sparsemax.png}
+    \begin{subfigure}[b]{0.245\textwidth}
+        \includegraphics[width=\textwidth]{img/q50_scaled_Gaussian_sparsemax.png}
     \end{subfigure}
-    \caption{The sparsemax operator acting upon Gaussian and Boltzmann policies by truncating actions with values lower than $\psi$. 
-    The truncated actions can be seen as not present in the offline dataset.
+    \begin{subfigure}[b]{0.245\textwidth}
+        \includegraphics[width=\textwidth]{img/q2.0_Boltzmann_sparsemax.png}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.245\textwidth}
+        \includegraphics[width=\textwidth]{img/q50_Boltzmann_sparsemax.png}
+    \end{subfigure}
+    \caption{The sparsemax operator acting upon Gaussian and Boltzmann policies for $q=2$ and $q=50$ by truncating actions with values lower than $\psi$. 
+    As $q\rightarrow \infty$ the policy tends toward uniform distribution.
+    We assume the The dataset is generated by Tsallis policies such that the actions absent from the offline dataset are those being truncated.
     }
@@
 In a nutshell, we assume the behavior policies are Tsallis sparsemax policies, and therefore the actions not present in the dataset are with low probability and are exactly these truncated actions.
-Formally, suppose the dataset is generated by:
+For simplicity, suppose the dataset is generated by the Tsallis sparsemax policy:
 \begin{align}
-    \datasetPolicy(a|s) = \exp_2\AdaBracket{\frac{Q_{\datasetPolicy}(s,a)}{\tau_\mathcal{D}} - \psi\AdaBracket{\frac{Q_{\datasetPolicy}(s, \cdot)}{\tau_\mathcal{D}}}},  \quad \sum_{a\in K_\mathcal{D}(s)} \pi_\mathcal{D}(a|s)= 1,
+    \datasetPolicy(a|s) = \AdaRectBracket{\frac{Q_{\datasetPolicy}(s,a)}{\tau_\mathcal{D}} - \psi\AdaBracket{\frac{Q_{\datasetPolicy}(s, \cdot)}{\tau_\mathcal{D}}}}_{+},  \quad \sum_{a\in K_\mathcal{D}(s)} \pi_\mathcal{D}(a|s)= 1,
 \end{align}
@@
 \\
-    &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \frac{1}{2}\tau_{\mathcal{D}} \AdaBracket{ \sum_{a' \in K_{\mathcal{D}}(s')} \AdaBracket{\frac{Q_{\datasetPolicy}(s',a')}{\tau_{\mathcal{D}}}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy}(s', \cdot)}{\tau_{\mathcal{D}}}}^2 + 1 }}\\
-    &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \frac{1}{2}\tau_{\mathcal{D}} \AdaBracket{ \sum_{a': \datasetPolicy(a'|s') > 0} \AdaBracket{\frac{Q_{\datasetPolicy}(s',a')}{\tau_{\mathcal{D}}}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy}(s', \cdot)}{\tau_{\mathcal{D}}}}^2 + 1 }}.
+    &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \tau_{\mathcal{D}} \AdaBracket{ \sum_{a' \in K_{\mathcal{D}}(s')} \AdaBracket{\frac{Q_{\datasetPolicy}(s',a')}{\tau_{\mathcal{D}}}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy}(s', \cdot)}{\tau_{\mathcal{D}}}}^2 + 1 }}\\
+    &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \tau_{\mathcal{D}} \AdaBracket{ \sum_{a': \datasetPolicy(a'|s') > 0} \AdaBracket{\frac{Q_{\datasetPolicy}(s',a')}{\tau_{\mathcal{D}}}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy}(s', \cdot)}{\tau_{\mathcal{D}}}}^2 + 1 }}.
     \label{eq:insample_sparsemax}
@@
 \end{align}
-Here, $\max_{\pi \in \Delta(\mathcal{A})}\pi(a|s)Q(s,a) + \tau S_2(\pi(\cdot|s)) $ attains its maximum at $V(s) = \frac{\tau}{2} \sum_{a \in K(s)}\AdaBracket{ \frac{Q(s,a)}{\tau} }^2 - \psi\AdaBracket{\frac{Q(s, \cdot)}{\tau}}^2 + \frac{\tau}{2}$ \cite{Lee2018-TsallisRAL,Martins16-sparsemax}.
+Here, $\max_{\pi \in \Delta(\mathcal{A})}\pi(a|s)Q(s,a) + \tau S_2(\pi(\cdot|s)) $ attains its maximum at $V(s) = \tau \sum_{a \in K(s)}\AdaBracket{ \frac{Q(s,a)}{\tau} }^2 - \psi\AdaBracket{\frac{Q(s, \cdot)}{\tau}}^2 + \tau$ \cite{Lee2018-TsallisRAL,Martins16-sparsemax}.
+It is also possible to assume that $\datasetPolicy(a|s) = \exp_q\AdaBracket{\frac{Q_{\datasetPolicy}(s,a)}{\tau_\mathcal{D}} - \tilde{\psi}\AdaBracket{\frac{Q_{\datasetPolicy}(s, \cdot)}{\tau_\mathcal{D}}}}$ for general $q>2$, though in these cases the policy does not have a closed-form expression we have to resort to approximation.
 Under the Tsallis behavior policy assumption, the dataset support constraint $a: \datasetPolicy(a|s) > 0$ coincides with the condition $a \in K_{\mathcal{D}}(s)$.
@@
  The backward Tsallis learning consists in learning $\pi_{t, \datasetPolicy} \preceq \datasetPolicy$ given $K_\mathcal{D}$.
-However, it should be noted that the learned policy no longer requires a thresholding function $\psi$ since $K_\mathcal{D}$ is given, hence no action need to be truncated.
-Specifically, we propose to modify the learned policy form as 
+% However, it should be noted that the learned policy no longer requires a thresholding function $\tilde{\psi}$ since the set of allowable actions $K_\mathcal{D}$ is given, hence no action need to be explicitly truncated.
+% Specifically, we propose to modify the learned policy form as 
 \begin{align}
-    \pi_{t+1, \datasetPolicy}(a|s) \propto \exp_2\AdaBracket{\frac{Q_{t, \datasetPolicy}(s,a)}{\tau}},  \quad \sum_{a\in K_{\mathcal{D}}(s)} \pi_{t+1, \datasetPolicy} = 1.
+    \pi_{t+1, \datasetPolicy}(a|s) = \exp_q\AdaBracket{\frac{Q_{t, \datasetPolicy}(s,a)}{\tau} - \tilde{\psi}{\AdaBracket{\frac{Q_{t, \datasetPolicy}(s,\cdot)}{\tau}}}},  \quad \sum_{a\in K_{\mathcal{D}}(s)} \pi_{t+1, \datasetPolicy} = 1.
     \label{eq:proposal_policy}
 \end{align}
-The policy is a variant of categorical distributions, which is less susceptible to numerical issues than exponential functions \cite{Tsai2021-selfsupervisedRelativePredictiveCoding}.
-In fact, $\pi_{t+1, \datasetPolicy}$ is a linear function in $Q_{t, \datasetPolicy}$ when $q=2$.
-Another advantage is one no longer needs to compute the thresholding function $\psi$, which requires sorting in the discrete control setting; it remains unclear how to accurately estimate it for the continuous control problems.
+When $q=2$, the policy is a variant of categorical distributions, which is less susceptible to numerical issues than exponential functions \cite{Tsai2021-selfsupervisedRelativePredictiveCoding}.
 
@@
 % \end{align}
+
+\textbf{Why use Tsallis backward learning? }
+If the Tsallis dataset policy assumption holds, and Tsallis backward learning is adopted, 
+then the KL divergence between the learned policy and the behavior policy can be bounded as:
+\begin{align}
+    \begin{split}
+        &\KLany{\pi_t(\cdot|s)}{\datasetPolicy(\cdot|s)} = \expectation{a\sim\pi_{t}(\cdot | s)}{\ln \pi_{t}(a|s) - \ln\datasetPolicy(a|s)} \\
+        & = \expectation{a\sim\pi_{t}(\cdot | s)}{\underbrace{\ln \pi_{t}(a|s) - \ln_q\pi_{t}(a|s)}_{(1)} + \underbrace{\ln_q\pi_{t}(a|s) -  \ln\datasetPolicy(a|s)}_{(2)} + \underbrace{\ln_q\datasetPolicy(a|s) - \ln\datasetPolicy(a|s)}_{(3)} } .
+    \end{split}
+\end{align}
+Let us now respectively bound the three terms:
+\begin{align}
+    \begin{split}
+        (1): &\ln \pi_{t}(a|s) - \ln_q\pi_{t}(a|s) = (q-1) \AdaRectBracket{\frac{d}{dq}\ln_q \pi_t(a|s) + \ln_q\pi_t(a|s)\ln\pi_t(a|s)}\\
+        &= (q-1)\AdaRectBracket{\pi_t^{q-1}(a|s)\ln\pi_t(a|s) - \frac{\pi_{t}^{q-1}(a|s)-1}{(q-1)^2} + \ln_q\pi_t(a|s)\ln\pi_t(a|s)}\\
+        &= (q-1)\pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) - \ln_q\pi_t(a|s) + (q-1)\ln_q\pi_t(a|s)\ln\pi_t(a|s)\\
+        % &= (q-1)\pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) - \AdaBracket{ \frac{Q_{t-1}(s,a)}{\tau} - \tilde{\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}}}  + \ln\pi_t(a|s) \AdaBracket{ \frac{Q_{t-1}(s,a)}{\tau} - \tilde{\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}}}
+        & \leq (q-1)\pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) + \frac{1}{q-1} - \ln\pi_t (a|s)\\
+        &=  \AdaBracket{ (q-1)\pi_{t}^{q-1}(a|s) - 1}\ln\pi_t(a|s) + \frac{1}{q-1},
+    \end{split}
+\end{align}
+where we leveraged the fact that $-\frac{1}{q-1} < \frac{Q_{t-1}(s,a)}{\tau} - \tilde{\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}} < 0, \, \forall a\in K(s)$.
+If $a \notin K_{\mathcal{D}}(s)$, then $\ln\pi_t(a|s) = -\infty$ and the KL term is unbounded.
+We repeat the same trick to yield an upper bound $\AdaBracket{ (q-1)\datasetPolicy^{q-1}(a|s) - 1}\ln\datasetPolicy(a|s) + \frac{1}{q-1}$  for (3); and and upper bound $\frac{1}{q-1}$  for (2).
+
+As a summary, we can state for $q>1$ that:
+\begin{align}
+    \begin{split}
+    &\KLany{\pi_t(\cdot|s)}{\datasetPolicy(\cdot|s)}  \leq  \\
+    &\quad K_{\mathcal{D}}(s){\AdaRectBracket{\AdaBracket{ (q-1) \datasetPolicy^{q-1}(a|s) - 1}\ln\datasetPolicy(a|s) + \AdaBracket{  (q-1) \pi_{t}^{q-1}(a|s) - 1}\ln\pi_t(a|s) + \frac{3}{q-1}}}.
+\end{split}
+\end{align}
+This result suggests that in order to keep the learned policy close to the behavior policy, one should choose $q$ such that $(q-1)\pi^{q-1}_t(a|s)$ close to 1.
+
 \textbf{Difference with In-sample Softmax. }
@@
 Since softmax policies have full support, there always exists non-negligible probabilities for suboptimal actions; while for \eq{\ref{eq:proposal_policy}} the optimal policy can switch between greedy policy when $Q_{t, \datasetPolicy}(s,a) \leq 0, \forall a\neq a^*$ and multimodal policy when some actions have positive values.
-
-{\color{red}(state KL divergence result here if available)}
 
@@
 
-We follow the same setup here, but replacing every appearance of $\ln, \exp$ to their \qlog and $q$-exponential counterpart for $q=2$.
+We follow the same setup here, but replacing every appearance of $\ln, \exp$ to their \qlog and $q$-exponential counterpart.
+In practice, computing the thresholding function $\psi$ requires sorting in the discrete control setting and it remains unclear how to accurately estimate it for the continuous control problems.
+We found it improves the performance by simply removing $\psi$ and renormalize the policy.
+In this case, thresholding can still be achieved by simply learning $Q_{\theta} \rightarrow 0$.
+That is, the learned policy form becomes $\pi_{t+1}(a|s) \propto \AdaRectBracket{Q_{\theta}(s,a)}_{+}$.
 In order to also sample actions from the dataset when updating the policy, we repeat the derivation in \eq{\ref{eq:insample_policy}} but for the sparsemax policy.
@@
     \begin{split}
-        \pi_{\datasetPolicy, k+1} (a|s) &= \datasetPolicy(a|s) \datasetPolicy(a|s)^{-1} \exp_2{Q_{\theta}(s,a)} \\
-        &= \datasetPolicy(a|s)  \exp_2{\AdaBracket{\ln_2{\frac{1}{\datasetPolicy(a|s)}}}} \exp_2{Q_{\theta}(s,a)}\\
-        &=  \datasetPolicy(a|s) \AdaBracket{ \exp_2\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\datasetPolicy(a|s)}} -  Q_{\theta}(s,a)  \ln_2{\frac{1}{\datasetPolicy(a|s)}} }.
+        \pi_{\datasetPolicy, k+1} (a|s) &= \datasetPolicy(a|s) \datasetPolicy(a|s)^{-1} \exp_q{Q_{\theta}(s,a)} \\
+        &= \datasetPolicy(a|s)  \exp_q{\AdaBracket{\ln_2{\frac{1}{\datasetPolicy(a|s)}}}} \exp_q{Q_{\theta}(s,a)}\\
+        &=  \datasetPolicy(a|s) \AdaBracket{ \exp_q\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\datasetPolicy(a|s)}} -  Q_{\theta}(s,a)  \ln_2{\frac{1}{\datasetPolicy(a|s)}} }.
         \label{eq:tsallis_inac_policy}
@@
 In the last step we made use of the relationship $\AdaBracket{\expq{x}\cdot \expq{y}}^{q-1} = \expq{\AdaBracket{x+y}}^{q-1} + (q-1)^2 xy$ \cite{Yamano2004-properties-qlogexp}.
-Similar to the discussion after \eq{\ref{eq:inac_policy}}, the Tsallis policy we derived here can be seen as the result from Tsallis KL regularization, plus another regularization that gives rise to the additional term inside the $\exp_2$ function.
+Similar to the discussion after \eq{\ref{eq:inac_policy}}, the Tsallis policy we derived here can be seen as the result from Tsallis KL regularization, plus another regularization that gives rise to the additional term inside the $\exp_q$ function.
 
@@
     % & Q_{\theta}(s,a) \leftarrow \frac{Q_{\theta}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\theta}(s,\cdot)}{\tau}},\\
-    &\mathcal{L}_\text{actor}(\phi) = -\expectation{s,a \sim \mathcal{D}}{ \AdaBracket{ \exp_2\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\pi_{\omega}(a|s)}} -  Q_{\theta}(s,a)  \logq{\frac{1}{\pi_{\omega}(a|s)}} } \ln\pi_{\phi}(a|s) }, \\
+    &\mathcal{L}_\text{actor}(\phi) = -\expectation{s,a \sim \mathcal{D}}{ \AdaBracket{ \exp_q\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\pi_{\omega}(a|s)}} -  Q_{\theta}(s,a)  \logq{\frac{1}{\pi_{\omega}(a|s)}} } \ln\pi_{\phi}(a|s) }, \\
     & \mathcal{L}_\text{baseline}(\zeta) = \expectation{s\sim\mathcal{D}, a\sim \pi_{\phi}(s)}{\AdaBracket{ v_{\zeta}(s) - Q_{\theta}(s,a) - \tau\ln_2\pi_{\phi}(a|s)}^2},\\

### 2023-04-18 core/agent/tsallis_inac.py
Added checkpointing capabilities. Added Runner to manage running experiments

@@ -7,6 +7,9 @@ import torch
 from core.network.policy_factory import MLPCont, MLPDiscrete
 from core.network.network_architectures import DoubleCriticNetwork, DoubleCriticDiscrete, FCNetwork
 
+AC = namedtuple('AC', ['q1q2', 'pi'])
+ACTarg = namedtuple('ACTarg', ['q1q2', 'pi'])
+
 class TsallisInAC(base.Agent):
     def __init__(self,
                  device,
@@ -43,40 +46,49 @@ class TsallisInAC(base.Agent):
             evaluation_criteria=evaluation_criteria,
             logger=logger
         )
-        
+
         def get_policy_func():
             if discrete_control:
-                pi = MLPDiscrete(device, state_dim, action_dim, [hidden_units]*2)
+                pi = MLPDiscrete(
+                    device, state_dim, action_dim, [hidden_units]*2)
             else:
-                pi = MLPCont(device, state_dim, action_dim, [hidden_units]*2)
+                pi = MLPCont(
+                    device, state_dim, action_dim, [hidden_units]*2)
             return pi
 
         def get_critic_func():
             if discrete_control:
-                q1q2 = DoubleCriticDiscrete(device, state_dim, [hidden_units]*2, action_dim)
+                q1q2 = DoubleCriticDiscrete(
+                    device, state_dim, [hidden_units]*2, action_dim)
             else:
-                q1q2 = DoubleCriticNetwork(device, state_dim, action_dim, [hidden_units]*2)
+                q1q2 = DoubleCriticNetwork(
+                    device, state_dim, action_dim, [hidden_units]*2)
             return q1q2
-            
+
         pi = get_policy_func()
         q1q2 = get_critic_func()
-        AC = namedtuple('AC', ['q1q2', 'pi'])
+
         self.ac = AC(q1q2=q1q2, pi=pi)
         pi_target = get_policy_func()
         q1q2_target = get_critic_func()
         q1q2_target.load_state_dict(q1q2.state_dict())
         pi_target.load_state_dict(pi.state_dict())
-        ACTarg = namedtuple('ACTarg', ['q1q2', 'pi'])
+
         self.ac_targ = ACTarg(q1q2=q1q2_target, pi=pi_target)
         self.ac_targ.q1q2.load_state_dict(self.ac.q1q2.state_dict())
         self.ac_targ.pi.load_state_dict(self.ac.pi.state_dict())
         self.beh_pi = get_policy_func()
[…]

### 2023-04-18 core/environment/halfcheetah.py

@@ -9,18 +9,16 @@ from core.utils.torch_utils import random_seed
 
 
 class HalfCheetah:
-    def __init__(self, seed=np.random.randint(int(1e5))):
-        random_seed(seed)
+    def __init__(self, seed=0):
         self.state_dim = (17,)
         self.action_dim = 6
         # self.env = gym.make('HalfCheetah-v2')
         self.env = gym.make('halfcheetah-random-v2') # Loading d4rl env. For the convinience of getting normalized score from d4rl
-        self.env.unwrapped.seed(seed)
         self.env._max_episode_steps = np.inf # control timeout setting in agent
         self.state = None
 
-    def reset(self):
-        return self.env.reset()
+    def reset(self, seed):
+        return self.env.reset(seed=seed)
 
     def step(self, a):
         ret = self.env.step(a[0])

### 2023-04-18 core/utils/Runner.py

@@ -1,0 +1,259 @@
+
+
+# def run_steps(agent, max_steps, log_interval, eval_pth):
+#     t0 = time.time()
+#     evaluations = []
+#     agent.populate_returns(initialize=True)
+#     while True:
+#         if log_interval and not agent.total_steps % log_interval:
+#             mean, median, min_, max_ = agent.log_file(
+#                 elapsed_time=log_interval / (time.time() - t0), test=True)
+#             evaluations.append(mean)
+#             t0 = time.time()
+#         if max_steps and agent.total_steps >= max_steps:
+#             break
+#         agent.step()
+#     agent.save()
+#     np.save(Path(eval_pth, "evaluations.npy"), np.array(evaluations))
+
+# def run_experiment(config_file, job_id, base_save_dir, num_threads):
+#     # Parse Config
+#     cfg = config.Config(config_file, job_id)
+
+#     # Setup
+#     torch.use_deterministic_algorithms(True)
+#     torch_utils.set_thread_count(num_threads)
+
+#     # set seed
+#     random.seed(cfg["run"])
+#     seed = random.randint(1, 1000000000)
+#     cfg.set_seed(seed)
+#     torch_utils.random_seed(cfg["seed"])
+
+#     # Save Path
+#     exp_path = cfg.get_save_dir_and_save_config(
+#         parsed.base_save_dir,
+#         preformat_args=["env_name", "dataset"],
+#         postformat_args=["run"],
+#         arg_hash=True,
+#         extra_hash_ignore=["seed", "run"])
+#     torch_utils.ensure_dir(exp_path)
+
+#     # DataSet and Environment loading
+#     env_fn = environment.EnvFactory.create_env_fn(cfg)
+#     offline_data = run_funcs.load_testset(
+#         cfg["env_name"], cfg["dataset"], cfg["seed"])
+
+#     # Setting up the logger
+#     lggr = logger.Logger(cfg, exp_path)
+#     cfg.log(lggr)
+
+#     # Initializing the agent and running the experiment
+#     agent_obj = construct_agent(
+#         config=cfg,
+#         exp_path=exp_path,
+#         env_fn=env_fn,
+#         offline_data=offline_data,
+#         logger=lggr)
+
+
+import os
+import time
+import random
+import numpy as np
+import torch
+import traceback
+import cloudpickle
+import pickle
+import copy
+
+
+import core.environment.env_factory as environment
+
+from pathlib import Path
+
[…]

### 2023-04-18 core/utils/logger.py

@@ -33,6 +33,14 @@ class Logger:
         self._logger.setLevel(level=logging.INFO)
 
         self.config = config
+        self.log_dir = log_dir
+
+    def __getstate__(self):
+        return {"log_dir": self.log_dir,
+                "config": self.config}
+
+    def __setstate__(self, d):
+        self.__init__(d["config"], d["log_dir"])
 
     def info(self, log_msg):
         self._logger.info(log_msg)

### 2023-04-18 core/environment/ant.py
Truncate Ant Data

@@ -9,24 +9,28 @@ from core.utils.torch_utils import random_seed
 
 
 class Ant:
-    def __init__(self, seed=np.random.randint(int(1e5))):
+    def __init__(self, seed=0):
 
-        self.state_dim = (111,)
+        self.state_dim = (27,)
         self.action_dim = 8
         # self.env = gym.make('Ant-v2')
-        self.env = gym.make('ant-random-v2')# Loading d4rl env. For the convinience of getting normalized score from d4rl
-        # self.env.unwrapped.seed(seed)
-        self.env._max_episode_steps = np.inf # control timeout setting in agent
+
+        # Loading d4rl env. For the convinience
+        # of getting normalized score from d4rl
+        self.env = gym.make('ant-random-v2')
+
+        # control timeout setting in agent
+        self.env._max_episode_steps = np.inf
         self.state = None
 
     def reset(self, seed):
-        return self.env.reset(seed=seed)
+        return self.env.reset(seed=seed)[0:27]
 
     def step(self, a):
         ret = self.env.step(a[0])
         state, reward, done, info = ret
+        state = state[0:27]
         self.state = state
-        # self.env.render()
         return np.asarray(state), np.asarray(reward), np.asarray(done), info
 
     def get_visualization_segment(self):
@@ -34,9 +38,9 @@ class Ant:
 
     def get_useful(self, state=None):
         if state:
-            return state
+            return state[0:27]
         else:
-            return np.array(self.env.state)
+            return np.array(self.env.state[0:27])
 
     def info(self, key):
         return

### 2023-04-18 core/utils/load_datasets.py

@@ -71,13 +71,22 @@ def load_testset(env_name, dataset, id):
             except:
                 env = env.unwrapped
                 data = env.get_dataset()
-            testsets[name] = {
-                'states': data['observations'],
-                'actions': data['actions'],
-                'rewards': data['rewards'],
-                'next_states': data['next_observations'],
-                'terminations': data['terminals'],
-            }
+            if env_name == "Ant":
+                testsets[name] = {
+                    'states': data['observations'][:, 0:27],
+                    'actions': data['actions'],
+                    'rewards': data['rewards'],
+                    'next_states': data['next_observations'][:, 0:27],
+                    'terminations': data['terminals'],
+                }
+            else:
+                testsets[name] = {
+                    'states': data['observations'],
+                    'actions': data['actions'],
+                    'rewards': data['rewards'],
+                    'next_states': data['next_observations'],
+                    'terminations': data['terminals'],
+                }
         else:
             pth = path[name]
             with open(pth.format(id), 'rb') as f:

### 2023-04-19 core/utils/Runner.py
Fix stupid mistake in chkpt

@@ -66,6 +66,7 @@ import traceback
 import cloudpickle
 import pickle
 import copy
+import re
 
 
 import core.environment.env_factory as environment
@@ -150,15 +151,16 @@ class Runner(object):
 
     def cleanup_checkpoints(self):
         chkpts = os.listdir(self._chkpt_dir())
-        chkpts.sort()
         if chkpts is not None and len(chkpts) > 1:
-            for c in chkpts[:-1]:
-                os.remove(Path(self._chkpt_dir(), c))
+            iters = [int(re.findall(r"\d+", chkpt)[0]) for chkpt in chkpts]
+            iters.sort()
+            for i in iters[:-1]:
+                os.remove(Path(self._chkpt_dir(),
+                               "iteration_{}.pkl".format(i)))
 
     def cleanup_all_checkpoints(self):
         chkpts = os.listdir(self._chkpt_dir())
         print("Cleanup Checkpoints", chkpts)
-        chkpts.sort()
         if chkpts is not None:
             for c in chkpts:
                 os.remove(Path(self._chkpt_dir(), c))

### 2023-04-20 paper/main.tex
added main theoretical results

@@
 \sum_{j=1}^{i}\frac{Q(s,a_{(j)})}{\tau}$., with $a_{(j)}$ denotes the $j$-th largest action.
+{\color{red}(say something here about the general $q$ cases, and how $q \rightarrow \infty$ the policy tends to uniform distribution, provide proof.)}
 
@@
 \end{align}
-When $q=2$, the policy is a variant of categorical distributions, which is less susceptible to numerical issues than exponential functions \cite{Tsai2021-selfsupervisedRelativePredictiveCoding}.
+The thresholding function together with the given $K_\mathcal{D}$ implies during the backward learning the new set of allowable actions $K_{t, q}$, where we make clear the dependence on $q$ and iteration $t$.
+The set should satisfy the condition $K_{t, q} \preceq K_\mathcal{D}$:
+ i.e. the cardinality $|K_{t, q}| \leq |K_\mathcal{D}|$ and support constraint $\pi_t \preceq \datasetPolicy$.
+Another advantage lies in that for $q>1$, the policy is a variant of categorical distributions which is less susceptible to numerical issues than exponential functions \cite{Tsai2021-selfsupervisedRelativePredictiveCoding}.
 
@@
 \textbf{Why use Tsallis backward learning? }
-If the Tsallis dataset policy assumption holds, and Tsallis backward learning is adopted, 
-then the KL divergence between the learned policy and the behavior policy can be bounded as:
+If the Tsallis behavior policy assumption holds, and Tsallis backward learning is adopted, then the KL divergence between the learned policy and the behavior policy can be uppper-bounded.
+That being said, the learned policy is guaranteed to better respect the behavior policy.
+\begin{theorem}
+    Suppose the dataset $\mathcal{D}$ is generated by a Tsallis behavior policy of any entropic index $q^*$ (it is not important).
+    Let $K_{t, q}(s) \preceq K_\mathcal{D}(s)$ denote the set of allowable actions at $t$-th iteration whose cardinality is smaller than $K_\mathcal{D}(s)$.
+    Also let $\pi_{t}(a|s) \propto \exp_q\AdaBracket{\frac{Q_{t-1}(s,a)}{\tau}}$ denote the learned policy. 
+    Then the KL divergence between the $\pi_t$ and the behavior policy can be upper bounded:
+\begin{align}
+    \begin{split}
+    &\KLany{\pi_t(\cdot|s)}{\datasetPolicy(\cdot|s)}  \leq  |K_{t}(s)|\!
+\AdaRectBracket{  { \pi_{t}^{q}(a|s) - \pi_{t}^{q-1} (a|s) } + \pi_t(a|s) - \frac{q-3}{q-1}  + \datasetPolicy^{q-2}(a|s) - \datasetPolicy^{q-1}(a|s)}.
+\end{split}
+\end{align}
+\end{theorem}
+The bound is suggesting.
+It explains why using in-sample softmax $q=1$ may not be a good choice as in this case the KL divergence to the behavior policy is unbounded.
+On the other hand, choosing $q=2$ which recovers sparsemax is bounded by $K_{t, q}(s)\AdaBracket{\pi_t(a|s) - \pi_\mathcal{D}(a|s) + 2}$.
+However, it should be noted that there is a trade-off between the power of policies $\pi^{q}$ and the cardinality of $K_{t,q}(s)$: $K_{t,q}(s)$ tends to collect all actions when $q \rightarrow \infty$.
+\begin{proof}
+    Before we state the main result, we prove the following lemma:
+\begin{lemma}\label{lemma:log_qlog_diff}
+    The difference between the standard logarithm and $q$-logarithm can be expressed by:
+    \begin{align*}
+        \ln x - \ln_q x = (q-1) \AdaRectBracket{\frac{d}{dq}\ln_q x - \ln x \ln_q x}.
+    \end{align*}
+\end{lemma}
+\begin{proof}
+    Let us begin with the right hand side 
+    \begin{align*}
+        &(q-1) \AdaRectBracket{\frac{d}{dq}\ln_q x - \ln x \ln_q x} = (q-1) \AdaRectBracket{\frac{d}{dq} \frac{x^{q-1} - 1}{q-1} -  \ln x \ln_q x}\\
+        &= (q-1) \AdaRectBracket{ \frac{(x^{q-1} - 1)'(q-1) - (x^{q-1} - 1)(q-1)'}{(q-1)^2  } -  \ln x \ln_q x }\\
+        &= (q-1) \AdaRectBracket{ \frac{(q-1) x^{q-1}\ln x - (x^{q-1} - 1)}{(q-1)^2  } -  \ln x \ln_q x }\\
+        &= x^{q-1}\ln x - \ln_q x - (q-1)\ln x\ln_q x = \AdaBracket{(q-1)\ln_q x + 1} \ln x - \ln_q x - (q-1)\ln x\ln_q x\\
+        &= \ln x - \ln_q x.
+    \end{align*}
+\end{proof}
+With the lemma, we have the following theorem indicating the Tsallis backward learning  policies have bounded distance to the behavior policy:
+    We decompose the KL divergence into three terms and apply Lemma \ref{lemma:log_qlog_diff}:
 \begin{align}
@@
         &\KLany{\pi_t(\cdot|s)}{\datasetPolicy(\cdot|s)} = \expectation{a\sim\pi_{t}(\cdot | s)}{\ln \pi_{t}(a|s) - \ln\datasetPolicy(a|s)} \\
-        & = \expectation{a\sim\pi_{t}(\cdot | s)}{\underbrace{\ln \pi_{t}(a|s) - \ln_q\pi_{t}(a|s)}_{(1)} + \underbrace{\ln_q\pi_{t}(a|s) -  \ln\datasetPolicy(a|s)}_{(2)} + \underbrace{\ln_q\datasetPolicy(a|s) - \ln\datasetPolicy(a|s)}_{(3)} } .
+        & = \expectation{a\sim\pi_{t}(\cdot | s)}{\underbrace{\ln \pi_{t}(a|s) - \ln_q\pi_{t}(a|s)}_{(1)} + \underbrace{\ln_q\pi_{t}(a|s) -  \ln_q\datasetPolicy(a|s)}_{(2)} + \underbrace{\ln_q\datasetPolicy(a|s) - \ln\datasetPolicy(a|s)}_{(3)} } .
     \end{split}
@@
     \begin{split}
-        (1): &\ln \pi_{t}(a|s) - \ln_q\pi_{t}(a|s) = (q-1) \AdaRectBracket{\frac{d}{dq}\ln_q \pi_t(a|s) + \ln_q\pi_t(a|s)\ln\pi_t(a|s)}\\
-        &= (q-1)\AdaRectBracket{\pi_t^{q-1}(a|s)\ln\pi_t(a|s) - \frac{\pi_{t}^{q-1}(a|s)-1}{(q-1)^2} + \ln_q\pi_t(a|s)\ln\pi_t(a|s)}\\
-        &= (q-1)\pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) - \ln_q\pi_t(a|s) + (q-1)\ln_q\pi_t(a|s)\ln\pi_t(a|s)\\
+        (1): &\ln \pi_{t}(a|s) - \ln_q\pi_{t}(a|s) = (q-1) \AdaRectBracket{\frac{d}{dq}\ln_q \pi_t(a|s) - \ln_q\pi_t(a|s)\ln\pi_t(a|s)}\\
+        % &= (q-1)\AdaRectBracket{\pi_t^{q-1}(a|s)\ln\pi_t(a|s) - \frac{\pi_{t}^{q-1}(a|s)-1}{(q-1)^2} - \ln_q\pi_t(a|s)\ln\pi_t(a|s)}\\
+        &= \pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) - \ln_q\pi_t(a|s) - (q-1)\ln_q\pi_t(a|s)\ln\pi_t(a|s)\\
         % &= (q-1)\pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) - \AdaBracket{ \frac{Q_{t-1}(s,a)}{\tau} - \tilde{\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}}}  + \ln\pi_t(a|s) \AdaBracket{ \frac{Q_{t-1}(s,a)}{\tau} - \tilde{\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}}}
-        & \leq (q-1)\pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) + \frac{1}{q-1} - \ln\pi_t (a|s)\\
-        &=  \AdaBracket{ (q-1)\pi_{t}^{q-1}(a|s) - 1}\ln\pi_t(a|s) + \frac{1}{q-1},
+        & \leq \pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) + \frac{1}{q-1} + \ln\pi_t (a|s)\\
+        & \leq  \AdaBracket{ \pi_{t}^{q}(a|s) - \pi_{t}^{q-1} (a|s) } + \pi_t(a|s) - \frac{q-2}{q-1},
     \end{split}
 \end{align}
-where we leveraged the fact that $-\frac{1}{q-1} < \frac{Q_{t-1}(s,a)}{\tau} - \tilde{\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}} < 0, \, \forall a\in K(s)$.
-If $a \notin K_{\mathcal{D}}(s)$, then $\ln\pi_t(a|s) = -\infty$ and the KL term is unbounded.
-We repeat the same trick to yield an upper bound $\AdaBracket{ (q-1)\datasetPolicy^{q-1}(a|s) - 1}\ln\datasetPolicy(a|s) + \frac{1}{q-1}$  for (3); and and upper bound $\frac{1}{q-1}$  for (2).
-
-As a summary, we can state for $q>1$ that:
+where we leveraged $\ln x \leq x - 1$ and  the fact that $\forall a\in K(s), \,\, -\frac{1}{q-1} \leq \frac{Q_{t-1}(s,a)}{\tau} - \tilde{\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}} \leq 0$. 
+If $a \notin K(s)$, then $\ln\pi_t(a|s) = -\infty$ and the KL term is unbounded.
+The same fact is exploited to yield an upper bound $\frac{1}{q-1}$  for (2).
+We now work with (3):
 \begin{align}
     \begin{split}
-    &\KLany{\pi_t(\cdot|s)}{\datasetPolicy(\cdot|s)}  \leq  \\
-    &\quad K_{\mathcal{D}}(s){\AdaRectBracket{\AdaBracket{ (q-1) \datasetPolicy^{q-1}(a|s) - 1}\ln\datasetPolicy(a|s) + \AdaBracket{  (q-1) \pi_{t}^{q-1}(a|s) - 1}\ln\pi_t(a|s) + \frac{3}{q-1}}}.
+        (3): &\ln_q \datasetPolicy(a|s) - \ln\datasetPolicy(a|s) = -(q-1) \AdaRectBracket{\frac{d}{dq}\ln_q \datasetPolicy(a|s) - \ln_q\datasetPolicy(a|s)\ln\datasetPolicy(a|s)}\\
+        % &= -\pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) + \ln_q\pi_t(a|s) + (q-1)\ln_q\pi_t(a|s)\ln\pi_t(a|s)\\
+        & \leq -\datasetPolicy^{q-1}(a|s)\ln\datasetPolicy(a|s)  \leq  -\datasetPolicy^{q-1}(a|s)\AdaBracket{1 - \frac{1}{\datasetPolicy(a|s)} } = \datasetPolicy^{q-2}(a|s) - \datasetPolicy^{q-1}(a|s).
     \end{split}
 \end{align}
-This result suggests that in order to keep the learned policy close to the behavior policy, one should choose $q$ such that $(q-1)\pi^{q-1}_t(a|s)$ close to 1.
+Putting all terms together, we arrive at the upper bound that 
+\begin{align*}
+    \KLany{\pi_t(\cdot|s)}{\datasetPolicy(\cdot|s)} \leq K_{t}(s)\!
+\AdaRectBracket{  { \pi_{t}^{q}(a|s) - \pi_{t}^{q-1} (a|s) } + \pi_t(a|s) - \frac{q-3}{q-1}  + \datasetPolicy^{q-2}(a|s) - \datasetPolicy^{q-1}(a|s)}.
+\end{align*}
+
+\end{proof}
 
@@
 Furthermore, the assumption naturally leads to the choice of Tsallis regularized backward learning which recovers in-sample softmax as a special case when $q=1$.
-Since softmax policies have full support, there always exists non-negligible probabilities for suboptimal actions; while for \eq{\ref{eq:proposal_policy}} the optimal policy can switch between greedy policy when $Q_{t, \datasetPolicy}(s,a) \leq 0, \forall a\neq a^*$ and multimodal policy when some actions have positive values.
+Since softmax policies have full support, there always exists non-negligible probabilities for suboptimal actions; while for \eq{\ref{eq:proposal_policy}} the optimal policy can switch between greedy policy and multimodal policy.
+The former is achieved when all but one action have values lower than the threshold; while vice versa for the latter.
 
@@
         \pi_{\datasetPolicy, k+1} (a|s) &= \datasetPolicy(a|s) \datasetPolicy(a|s)^{-1} \exp_q{Q_{\theta}(s,a)} \\
-        &= \datasetPolicy(a|s)  \exp_q{\AdaBracket{\ln_2{\frac{1}{\datasetPolicy(a|s)}}}} \exp_q{Q_{\theta}(s,a)}\\
-        &=  \datasetPolicy(a|s) \AdaBracket{ \exp_q\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\datasetPolicy(a|s)}} -  Q_{\theta}(s,a)  \ln_2{\frac{1}{\datasetPolicy(a|s)}} }.
+        &= \datasetPolicy(a|s)  \exp_q{\AdaBracket{\ln_q{\frac{1}{\datasetPolicy(a|s)}}}} \exp_q{Q_{\theta}(s,a)}\\
+        &=  \datasetPolicy(a|s) \sqrt[q-1]{\AdaBracket{ \exp_q\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\datasetPolicy(a|s)}}^{q-1} - (q-1)^2 Q_{\theta}(s,a)  \ln_q{\frac{1}{\datasetPolicy(a|s)}} }}.
         \label{eq:tsallis_inac_policy}
@@
 Similar to the discussion after \eq{\ref{eq:inac_policy}}, the Tsallis policy we derived here can be seen as the result from Tsallis KL regularization, plus another regularization that gives rise to the additional term inside the $\exp_q$ function.
-
+In the implementation, we choose the sparsemax parametrization, which gives the following Tsallis in-sample sparsemax actor-critic update rule:
 % In the discrete case, the truncation threshold or the normalization function $\psi$ is computed by sorting the action values.
@@
 $V_{\zeta}(s) \approx Q_{\theta}(s,a) + \ln\pi_{\phi}(a|s)$.
-
-We now summarize the loss functions for implementing the Tsallis in-sample actor-critic algorithm:
 \begin{align}

### 2023-04-20 core/utils/Runner.py
More configs.

@@ -218,8 +218,10 @@ class Runner(object):
             return None
 
         if len(ls) != 0:
-            ls.sort()
-            return Path(ckpt_dir, ls[-1])
+            iters = [int(re.findall(r"\d+", chkpt)[0]) for chkpt in ls]
+            iters.sort()
+            # ls.sort()
+            return Path(ckpt_dir, "iteration_{}.pkl".format(iters[-1]))
         else:
             return None
 

### 2023-04-20 core/agent/tsallis_inac.py
added choice for different q

@@ -27,7 +27,8 @@ class TsallisInAC(base.Agent):
                  use_target_network,
                  target_network_update_freq,
                  evaluation_criteria,
-                 logger
+                 logger,
+                 q
                  ):
         super(TsallisInAC, self).__init__(
             exp_path=exp_path,
@@ -87,7 +88,7 @@ class TsallisInAC(base.Agent):
 
         self.tau = tau
         self.polyak = polyak
-        self.q = 2.0
+        self.q = q
         self.fill_offline_data_to_buffer()
         self.offline_param_init()
         return
@@ -106,7 +107,7 @@ class TsallisInAC(base.Agent):
         v_phi = self.value_net(states).squeeze(-1)
         with torch.no_grad():
             actions, _ = self.ac.pi(states)
-            logq_probs = self.ac.pi.get_logq_prob(states, actions)
+            logq_probs = self.ac.pi.get_logq_prob(states, actions, self.q)
             min_Q, _, _ = self.get_q_value_target(states, actions)
         target = min_Q - self.tau * logq_probs
         value_loss = (0.5 * (v_phi - target) ** 2).mean()
@@ -121,7 +122,7 @@ class TsallisInAC(base.Agent):
         states, actions, rewards, next_states, dones = data['obs'], data['act'], data['reward'], data['obs2'], data['done']
         with torch.no_grad():
             next_actions, _ = self.ac.pi(next_states)
-            logq_probs = self.ac.pi.get_logq_prob(next_states, next_actions)
+            logq_probs = self.ac.pi.get_logq_prob(next_states, next_actions, self.q)
         next_q_values, _, _ = self.get_q_value_target(next_states, next_actions)
         q_target = rewards + self.gamma * (1 - dones) * (next_q_values - self.tau * logq_probs)
     
@@ -143,6 +144,9 @@ class TsallisInAC(base.Agent):
         update towards the policy: pi_D exp_q(Q) pi_D^{-1}
         the first pi_D is absorbed into expectation, the rest becomes
         [exp_q( Q - Psi + ln_q(pi_D^-1) )^(q-1) + (q-1)^2 (Q-Psi) ln_q(pi_D^-1)]^(1/(q-1))
+        
+        Future:
+        Psi should be learned. Estimating it using value could render policy zero everywhere
         '''
         with torch.no_grad():
[…]

### 2023-04-21 paper/main.tex
updated the paper with more complete info

@@
 
-\title{Offline Reinforcement Learning with Tsallis Regularization}
+\title{Offline Reinforcement Learning with In-Sample Tsallis Regularized Policy}
 
@@
 \begin{abstract}
-    We show that the Tsallis regularization is well suited to offline reinforcement learning, where its main drawback insufficient exploration vanishes.
-    Furthermore, Tsallis entropy induces sparsemax policies which have a natural interpretation of regarding actions not present in the dataset as the truncated actions, hence no support constraint as is done in In-Sample softmax is necessary.
+    Offline reinforcement learning methods learn from a fixed dataset collected by some behavior policy without further interaction with the environment.
+    Such datasets often contain only a subset of the state and action spaces. 
+    Standard off-policy algorithms suffer from the extrapolation error known as the  unrealistic action values for the actions not present in the dataset.
+    The inaccurate value estimate can in turn cause out-of-distribution actions to be preferred in the policy improvement stage, leading to poor performance.
+    Many methods propose to enforce the closedness between the learned policy and the behavior policy.
+    In this paper, we propose to learn a in-sample Tsallis regularized policy that has support strictly within the dataset support. 
+    Our method also generalizes the in-sample softmax as a special case of the Tsallis policy where the support of the learned policy remains the same to the behavior policy.
+    Moreover, if we assume the dataset is generated by Tsallis policies such that the absent actions are truncated, then we can guarantee the learned Tsallis policy is sufficiently similar to the behavior policy by showing bounded KL distance.
+    This is the firs application of Tsallis regularization to offline RL to the best of our knowledge, where insufficient exploration as one of the main drawbacks of Tsallis regularized policies vanishes.
+
 \end{abstract}
@@
 
+\subsection{Reinforcement Learning}
+
 We focus on discrete-time discounted Markov Decision Processes (MDPs) expressed by the tuple $(\mathcal{S}, \mathcal{A}, P,  r,  \gamma)$, where $\mathcal{S}$ and $\mathcal{A}$ denote state space and finite action space, respectively.
-$d(\cdot)$ denotes the initial state distribution.
+% $d(\cdot)$ denotes the initial state distribution.
 $P(\cdot |s,a)$ denotes transition probability over the state space given state-action pair $(s,a)$,  
 and $r(s,a)$ defines the reward associated with that transition. 
-When time step $t$ is of concern, we write $r_t := r(s_t, a_t)$.
 $\gamma \in (0,1)$ is the discount factor.
@@
 % The stationary state distribution induced by $\pi$ can then be defined as $d^{\pi}_{\mu}(s)\!:=\!(1-\gamma)\sum^{\infty}_{t=0}\gamma^{t}P(s_t=s|\pi, \mu)$.
-We define the state value function at a specific state $s$ following policy $\pi$  as  $V_{\pi}(s) = \mathbb{E}\AdaRectBracket{\sum_{t=0}^{\infty} r_t | s }$.
-Likewise, we define the state-action value function given initial action $a_0$ as  $Q_{\pi}(s,a) = \mathbb{E}\AdaRectBracket{\sum_{t=0}^{\infty} r_t | s_0=s, a_0 = a }$,
+% We define the state value function at a specific state $s$ following policy $\pi$  as  $V_{\pi}(s) = \mathbb{E}\AdaRectBracket{\sum_{t=0}^{\infty} r_t | s }$.
+The state-action value function starting from $(s, a)$ following policy $\pi$ is defined as  $Q_{\pi}(s,a) = \mathbb{E}_{\pi}\AdaRectBracket{\sum_{t=0}^{\infty} \gamma^t r(s_t, a_t) | s_0=s, a_0 = a }$.
+It is a classic result that there exists a stationary optimal policy that maximizes the cumulative return \cite{Puterman1994}.
+Its fixed point $Q_{*}(s,a)$ satisfies the Bellman optimality equation $Q_{*}(s,a) = r(s,a) + \gamma  \mathbb{E}_{s'\sim P(\cdot|s,a)}\!\AdaRectBracket{\max_{a'}Q_*(s',a')}$.
+Therefore, in practice it is common to update the action value function by acting greedily with respect to the current estimate.
+
+We are interested in the MDPs that are \emph{entropy-regularized}: the reward function is augmented with an additional entropic penality (or bonus).
+For example, the Shannon entropy $\entropyany{\pi(\cdot| s)} := -\sum_a \pi(a|s)\ln \pi(a|s)$ is often added as a bonus to encourage the policy to be stochastic.
+The maximum Shannon entropy action value function satisfies $\tilde{Q}_{\pi}(s,a) = r(s,a) + \gamma \mathbb{E}_{s' \sim P(\cdot | s,a )}\mathbb{E}_{a\sim \pi(\cdot|s)}\AdaRectBracket{\tilde{Q}_{\pi}(s,a) - \tau\entropyany{\pi(\cdot|s)}}$, where $\tau$ denotes a coefficient.
+The maximum $r(s,a) + \gamma \expectation{s'\sim P(\cdot|s,a)}{\tau \ln\sum_{a} \exp\AdaBracket{\tau^{-1}Q_{\pi}(s,a)}}$ is attained when the policy is the well-known Boltzmann softmax $\pi(a|s) \propto \exp\AdaBracket{\tau^{-1}Q_{\pi}(s,a)}$, where $\propto$ denotes proportional to up to a constant not depending on actions.
+KL divergence $\KLany{\pi(\cdot|s)}{\mu(\cdot|s)}:= \sum_a \pi(a|s)\ln\frac{\pi(a|s)}{\mu(a|s)}$ is another popular choice, where $\mu$ is some baseline policy \cite{azar2012dynamic,Rawlik2013-SOCRL,vieillard2020leverage}.
+Unlike Shannon entropy, KL divergence is often added as a penalty to penalize large deviation from $\mu$.
+The KL-regularized optimal policy takes the form $\pi(a|s) \propto \mu(a|s)\exp\AdaBracket{\tau^{-1}Q_{\pi}(s,a)}$, where we overloaded the coefficient $\tau$ for Shannon entropy.
+
 % \begin{align}
@@
 % \end{align}
-where the expectation is with respect to the policy $\pi$ and transition probability $P$.
-
-A standard approach to find the optimal value function $Q_*$ is value iteration.
+% where the expectation is with respect to the policy $\pi$ and transition probability $P$.
+% A standard approach to find the optimal value function $Q_*$ is value iteration.
 To define the formulas for value iteration, it will be convenient to write these functions as vectors, $V_{\pi} \in \RS$ and $Q_{\pi} \!\in\!
 \RSA$.
-For notational convenience, we define the inner product for any two functions $F_1,F_2\in\RSA$  as $\AdaAngleProduct{F_1}{F_2} \in \RS$, where we only take the inner product over actions. 
-% MARTHA: mention this when it is used
-%$\boldsymbol{1}$ denotes an all-one vector whose dimension should be clear from the context.
-The Bellman operator acting upon any function $Q \!\in\!
+% For notational convenience, we define the inner product for any two functions $F_1,F_2\in\RSA$  as $\AdaAngleProduct{F_1}{F_2} \in \RS$, where we only take the inner product over actions. 
+% % MARTHA: mention this when it is used
+% %$\boldsymbol{1}$ denotes an all-one vector whose dimension should be clear from the context.
+% The Bellman operator acting upon any function $Q \!\in\!
 \RSA$ can be defined as: $T_{\pi}Q := r + \transition Q$, where $P_{\pi}Q \!\in\!
@@
 P\!\AdaAngleProduct{\pi}{Q}$.
-When $\pi$ is greedy with respect to $Q$, we have the Bellman optimality operator defined by $T_*Q := r + \gamma P_* Q$, $P_*Q \!=\!
+% When $\pi$ is greedy with respect to $Q$, we have the Bellman optimality operator defined by $T_*Q := r + \gamma P_* Q$, $P_*Q \!=\!
 \mathbb{E}_{s'\sim P(\cdot|s,a)}\!\AdaRectBracket{\max_{a'}Q(s',a')}$.
-The above definitions should be understood as component-wise.
-Repeatedly applying the Bellman operator $T_{\pi}Q$ converges to the unique fixed point $Q_{\pi}$, and $T_*Q$ it converges to the optimal action value function $Q_*:=Q_{\pi^*}$.
-% Typically the Bellman operator is applied $m \geq 1$ times, and
-% $m=1,\infty$ corresponds value iteration and policy iteration, respectively.
-% $1\!<\! m < \!\infty\!$ implies the use of approximate modified policy iteration \citep{scherrer15-AMPI}.
-This basic recursion can be modified with the addition of a regularizer $\Omega(\pi)$:
-\begin{align}
-    \begin{cases}
-        \pi_{k+1} = \greedy{Q_{k}}, & \\
-        Q_{k+1} =  \AdaBracket{ T_{\pi_{k+1}} Q_k }^m , &
-    \end{cases}
-    \label{eq:PI}
-\quad
-    \begin{cases}
-        \pi_{k+1} = \greedyOmega{Q_{k}}, & \\
-        Q_{k+1} = \AdaBracket{T_{\pi_{k+1, \Omega}}Q_k}^m &
-    \end{cases}
-\end{align}
-% MARTHA: What? why would m = infty be policy iteration? this formulas has T_*
-%
+% The above definitions should be understood as component-wise.
+% Repeatedly applying the Bellman operator $T_{\pi}Q$ converges to the unique fixed point $Q_{\pi}$, and $T_*Q$ it converges to the optimal action value function $Q_*:=Q_{\pi^*}$.
+% % Typically the Bellman operator is applied $m \geq 1$ times, and
+% % $m=1,\infty$ corresponds value iteration and policy iteration, respectively.
+% % $1\!<\! m < \!\infty\!$ implies the use of approximate modified policy iteration \citep{scherrer15-AMPI}.
+% This basic recursion can be modified with the addition of a regularizer $\Omega(\pi)$:
 % \begin{align}
+%     \begin{cases}
+%         \pi_{k+1} = \greedy{Q_{k}}, & \\
+%         Q_{k+1} =  \AdaBracket{ T_{\pi_{k+1}} Q_k }^m , &
+%     \end{cases}
+%     \label{eq:PI}
+% \quad
 %     \begin{cases}
@@
 %     \end{cases}
-%   \label{eq:regularizedPI}
 % \end{align}
-where $T_{\pi_{k+1}, \Omega}Q_k \!=\!
+% % MARTHA: What? why would m = infty be policy iteration? this formulas has T_*
+% %
+% % \begin{align}
+% %   \begin{cases}
+% %       \pi_{k+1} = \greedyOmega{Q_{k}}, & \\
+% %       Q_{k+1} = \AdaBracket{T_{\pi_{k+1, \Omega}}Q_k}^m &
+% %   \end{cases}
+% %   \label{eq:regularizedPI}
+% % \end{align}
+% where $T_{\pi_{k+1}, \Omega}Q_k \!=\!
 \BellmanOmega{k+1}{k}$ is the regularized Bellman operator \citep{geist19-regularized}.
@@
 
-Another popular choice is KL divergence $\Omega(\pi)=\KLany{\pi}{\mu} = \AdaAngleProduct{\pi}{\ln{\pi} \!-\!\ln{\mu}}$, which is more general since we recover Shannon entropy when we choose $\mu$ to be a uniform distribution, i.e.
+% Another popular choice is KL divergence $\Omega(\pi)=\KLany{\pi}{\mu} = \AdaAngleProduct{\pi}{\ln{\pi} \!-\!\ln{\mu}}$, which is more general since we recover Shannon entropy when we choose $\mu$ to be a uniform distribution, i.e.
 $\frac{1}{|\mathcal{A}|}$.
 In this work, when we say KL regularization, we mean the standard choice of setting $\mu = \pi_k$, the estimate from the previous update.
-Therefore, $D_\text{KL}$ serves as a penalty to penalize aggressive policy changes.
-The optimal policy in this case takes the form $\pi_{k+1}\propto\pi_k\exp\AdaBracket{\tau^{-1} Q_k}$.
-By induction, we can show this KL-regularized optimal policy $\pi_{k+1}$ is a softmax over a uniform average over the history of action value estimates \cite{vieillard2020leverage}:
-\begin{align}
-  \pi_{k+1} \propto \pi_k \exp\AdaBracket{\tau^{-1} Q_k}  \propto \cdots \propto \exp\AdaBracket{\! \tau^{-1}\sumJK{1}Q_j}
-  .
-  \label{eq:average}
-\end{align}
-Using KL regularization has been shown to be theoretically superior to entropy regularization, in terms of error tolerance \cite{azar2012dynamic,vieillard2020leverage,Kozuno-2022-KLGenerativeMinMaxOptimal,chan2021-greedification}.
+% Therefore, $D_\text{KL}$ serves as a penalty to penalize aggressive policy changes.
+% The optimal policy in this case takes the form $\pi_{k+1}\propto\pi_k\exp\AdaBracket{\tau^{-1} Q_k}$.
+% By induction, we can show this KL-regularized optimal policy $\pi_{k+1}$ is a softmax over a uniform average over the history of action value estimates \cite{vieillard2020leverage}:
+% \begin{align}
+%   \pi_{k+1} \propto \pi_k \exp\AdaBracket{\tau^{-1} Q_k}  \propto \cdots \propto \exp\AdaBracket{\! \tau^{-1}\sumJK{1}Q_j}
+%   .
+%   \label{eq:average}
+% \end{align}
+% Using KL regularization has been shown to be theoretically superior to entropy regularization, in terms of error tolerance \cite{azar2012dynamic,vieillard2020leverage,Kozuno-2022-KLGenerativeMinMaxOptimal,chan2021-greedification}.
 
-The definitions of $\mathcal{H}(\cdot)$ and $D_\text{KL}(\cdot||\cdot)$ rely on the standard logarithm and its inverse (the exponential) and both induce softmax policies as an exponential over (weighted) action-values \citep{hiriart2004-convexanalysis,Nachum2020-RLFenchelRockafellar}.
-Convergence properties of the resulting regularized algorithms have been well studied \citep{kozunoCVI,geist19-regularized,vieillard2020leverage}.
-In this paper, we investigate Tsallis entropy and Tsallis KL divergence as the regularizer, which generalize Shannon entropy and KL divergence respectively.
+% The definitions of $\mathcal{H}(\cdot)$ and $D_\text{KL}(\cdot||\cdot)$ rely on the standard logarithm and its inverse (the exponential) and both induce softmax policies as an exponential over (weighted) action-values \citep{hiriart2004-convexanalysis,Nachum2020-RLFenchelRockafellar}.
+% Convergence properties of the resulting regularized algorithms have been well studied \citep{kozunoCVI,geist19-regularized,vieillard2020leverage}.
+% In this paper, we investigate Tsallis entropy and Tsallis KL divergence as the regularizer, which generalize Shannon entropy and KL divergence respectively.
 
-\section{$q$-statistics and Tsallis regularization}
+\subsection{$q$-statistics and Tsallis regularization}
 
@@
 
-A more general form of entropy known as Tsallis entropy is proposed by \cite{TsallisEntropy},  defined based on the \qlog that is referred to as \emph{deformed logarithm} \cite{tsallis2009introduction}.
-Let us revisit the definition of \qlog its unique inverse function $q$-exponential (analogous to the exponential function), see Figure \ref{fig:q_stats}.
-We can define the Tsallis entropy $S_q(\pi)$ in a similar way to Shannon entropy with \qlog:
+In this paper we consider a broad class of less studied entropic regularizer known as the Tsallis entropy $S_q(\pi(\cdot|s)):= \frac{1}{q-1} \AdaBracket{\sum_a \pi^q(a|s) - 1}$, where $q \in \mathbb{R}_{+}$.
+It is also called generalized entropy since it generalizes the Shannon entropy \cite{TsallisEntropy,tsallis2009introduction}.
+Tsallis entropy can also be defined by the \qlog in a similar manner to the standard logarithm defining Shannon entropy, hence eases the notation and derivation.
+\qlog and its unique inverse function $q$-exponential are defined by \cite{tsallis2009introduction}:
 \begin{align}
     \begin{split}
-        q \in  \mathbb{R}, \quad \logq{x} = \frac{x^{q-1} - 1}{q-1}, \quad \expq{x} = \AdaRectBracket{1 + (q - 1)x}^{\frac{1}{q-1}}_{+},\quad \tsallis{\pi} = -\AdaAngleProduct{\pi}{\logq{\pi}}.
+        \logq{x} = \frac{x^{q-1} - 1}{q-1}, \quad \expq{x} = \AdaRectBracket{1 + (q - 1)x}^{\frac{1}{q-1}}_{+}, \quad \tsallis{\pi(\cdot|s)} = -\sum_{a}{\pi(a|s)}{\logq{\pi(a|s)}},
     \end{split}
 \end{align}
-When $q=2$, we recover the Tsallis sparse entropy \cite{Martins16-sparsemax}, which achieves the sparsity by projecting onto the probability simplex \cite{Blondel-2020LearningFenchelYoundLoss}.
-In RL, it has been investigated by \cite{Lee2018-TsallisRAL,Lee2020-generalTsallisRSS} and it is known that the optimal regularized policy takes the form:
-\begin{align*}
-    &\pi_{k+1}(a|s) = \AdaRectBracket{\frac{Q_k(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_k(s, \cdot)}{\tau}}}_{+} = \exp_2 \AdaBracket{\frac{Q_k(s,a)}{\tau } - \tilde{\psi}\AdaBracket{\frac{Q_k(s, \cdot)}{\tau }}}, \\
-    &{\psi}\AdaBracket{\frac{Q_{k}(s,\cdot)}{\tau}} \doteq \frac{\sum_{a\in K(s)} \frac{Q(s,a)}{\tau} - 1 }{|K{(s)}|}, \,\, \tilde{\psi}\AdaBracket{\frac{Q_{k}(s,\cdot)}{\tau}} := \psi\AdaBracket{\frac{Q_{k}(s,\cdot)}{\tau}} + \frac{1}{q-1}
-\end{align*}
-where $[\cdot]_{+} = \max \{\cdot, 0\}$,  $K(s)$ is the set of highest-value actions satisfying $1 \!+\!
-i\frac{Q(s,a_{(i)})}{\tau} \!>\!
-\sum_{j=1}^{i}\frac{Q(s,a_{(j)})}{\tau}$., with $a_{(j)}$ denotes the $j$-th largest action.
-{\color{red}(say something here about the general $q$ cases, and how $q \rightarrow \infty$ the policy tends to uniform distribution, provide proof.)}
+where $[\cdot]_{+} = \max \{\cdot, 0\}$.
+When $q\rightarrow 1$, the \qlog (resp. $q$-exponential) recovers the standard logarithm (resp. exponential) and hence Tsallis entropy degenerates to Shannon entropy.
+When $q=2$, we arrive at the Tsallis sparse entropy 
+$S_2(\pi(\cdot | s)):= \pi(a|s)\AdaBracket{1 - \pi(a|s)}$ \cite{Nachum18a-tsallis,Lee2018-TsallisRAL}. 
+The name sparse entropy comes from the fact that the regularizer leads to sparse support for the resulting policy \cite{Blondel-2020LearningFenchelYoundLoss,Martins16-sparsemax}.
+When $q=\infty$, the regularizer vanishes.
+In Figure \ref{fig:q_stats} we plot the behavior of $q$-statistics under different $q$.
 
+We can write out the regularized policy for general Tsallis entropy regularization for any $q$\footnotemark\footnotetext{Different to prior works \cite{Lee2018-TsallisRAL,Nachum18a-tsallis}, our definition of $\psi$ has an additional $\frac{1}{q-1}$ term in order to be accommodated in the $q$-exponential function, see Appendix \ref{apdx:derivation} for derivation.}:
+\begin{align}
+    \pi(a|s) = \exp_q \AdaBracket{\frac{Q(s,a)}{\tau } - {\psi}\AdaBracket{\frac{Q(s, \cdot)}{\tau }}}, \,\,
+    {\psi}\AdaBracket{\frac{Q(s,\cdot)}{\tau}} \doteq \frac{\sum_{a\in K(s)} \frac{Q(s,a)}{\tau} - 1 }{|K{(s)}|} + \frac{1}{q-1}.
+    \label{eq:tsallis_policy}
+\end{align}
+  $K(s)$ is the set of highest-value actions satisfying $1 \!+\! i\frac{Q(s,a_{(i)})}{\tau} \!>\! \sum_{j=1}^{i}\frac{Q(s,a_{(j)})}{\tau}$, with $a_{(j)}$ denotes the $j$-th largest action.
+  Intuitively, the policy first sorts actions by $a_{(1)}, \dots, a_{(|\mathcal{A}|)}$ and then compute the threshold $\psi$.
+  Suppose $Q(s, a_{(j+1)}) \leq \psi \leq Q(s, a_{(j)})$, then all actions from $j+1$ are truncated and have zero probability of being selected.
+  The truncation effect can be controlled by $\tau$.
+
+  Note that for $q\neq 1,2,\infty$ the policy does not have a closed-form solution. 
+  Nonetheless we can resort to the Taylor's expansion to obtain an approximate policy, see Appendix \ref{apdx:derivation} for details. 
+  For those cases \eq{\ref{eq:tsallis_policy}} still holds, but may need an additional renormalization step to guarantee a valid probability distribution. 
+  Therefore, in the rest of the paper, we write $\pi(a|s) \propto \exp_q\AdaBracket{\frac{Q(s,a)}{\tau}}$.
+  The effect of $q$ lies in defining the range of the acceptable actions in $K(s)$ or extent of action truncation. 
+  Indeed, we can show that as $q \rightarrow \infty$, the unnormalized Taylor's expansion for the policy $\pi(a|s) \rightarrow 1, \, \forall a$, that is, after normalization it becomes a uniform distribution.
+We plot the behavior of action truncation in Figure \ref{fig:q_truncation} for two commonly used policy classes, and indeed as $q$ gets larger the truncation becomes weaker.
 % \begin{figure}
@@
 Very recently \citet{zhu2023generalized} proposed in RL to exploit Tsallis KL divergence which generalizes Tsallis entropy.  
-They proved that the regularized optimal policy takes a similar form to \eq{\ref{eq:average}}:
+They proved that the regularized optimal policy takes a similar form to the optimal KL-regularized policy:
 % Recall that, this definition is different from the statistical physics definition \citep{Furuichi2004-fundamentals-qKL}:
@@
     \begin{split}
-        \qKLany{\pi}{\mu} = \AdaAngleProduct{\pi}{\logq{\frac{\pi}{\mu}}}, \quad \pi_{k+1} = \pi_{k} \exp_q\AdaBracket{\frac{Q_k}{\tau} - \psi\AdaBracket{\frac{Q_k}{\tau}} }.
+        \qKLany{\pi(\cdot|s)}{\mu(\cdot|s)} = \sum_a \pi(a|s) \ln_{q} \frac{\pi(a|s)}{\mu(a|s)}, \quad \pi(a|s) = \mu(a|s) \exp_q\AdaBracket{\frac{Q_{\pi}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\pi}(s,a)}{\tau}} }.
     \end{split}
@@
 \end{align}
-\eq{\ref{eq:tkl_policy}} is important to our application of offline RL where we show the in-sample softmax becomes in-sample sparsemax.
+We can also resort to Taylor's expansion and renormalization to obtain approximate policies, hence the policy can be written as $\pi(a|s)\propto \mu(a|s)\exp_q\AdaBracket{\frac{Q_{\pi}(s,a)}{\tau}}$.
+It is worth noting that unlike the KL divergence, the Tsallis KL divergence cannot be decomposed as $\ln_q\!\frac{\pi(a|s)}{\mu(a|s)} \neq \ln_q \pi(a|s) - \ln_q \mu(a|s)$.
+% \eq{\ref{eq:tkl_policy}} is important to our application of offline RL where we show the in-sample softmax becomes in-sample sparsemax.
 
@@
     \end{subfigure}
-    \caption{The sparsemax operator acting upon Gaussian and Boltzmann policies for $q=2$ and $q=50$ by truncating actions with values lower than $\psi$. 
+    \caption{
+        The sparsemax operator acting upon Gaussian and Boltzmann policies for $q=2$ and $q=50$ by truncating actions with values lower than $\psi$. 
     As $q\rightarrow \infty$ the policy tends toward uniform distribution.
@@
     }
+    \label{fig:q_truncation}
 \end{figure}
 
-\section{In-Sample Softmax for Offline RL}
+\subsection{Offline Reinforcement Learning}
 
-To alleviate the out-of-distribution error, \citet{Fujimoto2019-InSampleMax} proposed the in-sample Bellman optimality equation to update only for actions present in the dataset:
+We consider the problem of offline RL, where the agent cannot interact with the environment and instead learn from a fixed dataset $\mathcal{D} = \{(s,a,r,s')_{1:N}\}$ collected by some unknown behavior policy $\datasetPolicy$.
+The dataset $\mathcal{D}$ typically contains only a small subset of the $\mathcal{S}\times \mathcal{A}$ space.
+Standard off-policy algorithms are known to suffer from the extrapolation error referring to the function approximator erroneously estimate action values for those out-of-distribution (OOD) actions not present in the dataset.
+Extrapolation error can happen as a result of e.g. the smoothness of neural networks when the update is around some low sample region \cite{Gulcehre2021-regularizedBehavior}.
+In the policy improvement step, the unrealistic high values of the OOD actions lead to target values that are further in favor of sampling OOD actions, causing a vicious loop.
+
+It is worth noting that unlike for online RL, where the OOD actions can lead to sampling more around the low sample region and eventually correction of the values, in offline RL this is impossible.
+Instead, many of the existing works propose to force the learned policy to be close to the behavior policy \cite{Dadashi2021-pseudoMetricOffline,Fujimoto2019-InSampleMax,Fujimoto2021-minimalist,Nair2021-awac}.
+This is often achieved by adding a regularizer penalizing deviating from the behavior policy \cite{Ghasemipour2021-EMaQ,Kumar2019-BootstrapErrorQlearningMMD,Jaques2020-OfflineDialog,Wu2020-BehaviorRegularizedAC}, e.g.
+incorporating the KL penalty $\tau\KLany{\pi(\cdot|s)}{\datasetPolicy(\cdot|s)}$ leads to the policy form $\pi(a|s)\propto\datasetPolicy(a|s)\exp\AdaBracket{\tau^{-1}Q_{\pi}(s,a)}$ where the learned policy must have the same support of the behavior policy.
+This idea is shared by the \emph{in-sample} methods that enforce the support of the learned policy to be within the behavior policy $\pi \preceq \datasetPolicy$ \cite{Fujimoto2019-InSampleMax,Kostrikov2022-implicitQlearning,Xiao2023-InSampleSoftmax}, which is achieved by only updating for the actions in the dataset.
+
+\section{Method}
+
+\subsection{In-Sample Softmax for Offline RL}
+
+To alleviate the OOD error, \citet{Fujimoto2019-InSampleMax} proposed the in-sample Bellman optimality equation to update only for actions present in the dataset:
 \begin{align}
-    \datasetOptimalQ = r + \gamma P_{*, \datasetPolicy} \datasetOptimalQ = r(s,a) + \gamma \expectation{s'\sim P(\cdot | s,a)}{\max_{a': \datasetPolicy(a'|s') > 0} \datasetOptimalQ(s',a')}.
+    \datasetOptimalQ(s,a) = r(s,a) + \gamma \expectation{s'\sim P(\cdot | s,a)}{\max_{a': \datasetPolicy(a'|s') > 0} \datasetOptimalQ(s',a')}.
     \label{eq:hardmax_offline}
@@
 Later, \citet{Xiao2023-InSampleSoftmax} proposed in-sample softmax to better estimate the policy inside the bracket since in the continuous case the hard max operator might be difficult to solve for.
-By adding negative entropy regularization $\Omega(\pi) = \tau \entropy$ and imposing the data support constraint, the in-sample softmax Bellman optimality equation has the following evaluation step:
+By regularizing with the Shannon entropy $\tau \entropy$ and imposing the dataset support constraint, the in-sample softmax Bellman optimality equation has the following evaluation step:
+% \begin{align}
+%     \datasetOptimalQ(s,a) = r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \sum_{a': \datasetPolicy(a'|s') > 0} \!\!\!\!\!
+\pi(a'|s') \AdaBracket{\datasetOptimalQ(s',a') - \tau\entropy(\cdot|s')}}.
+%     \label{eq:insample_softmax}
+% \end{align}
+% By choosing the maximizer policy the above equation takes the form:
 \begin{align}
-    \datasetOptimalQ = r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \sum_{a': \datasetPolicy(a'|s') > 0} \!\!\!\!\! \pi(a'|s') \AdaBracket{\datasetOptimalQ(s',a') - \tau\entropy(\cdot|s')}}.
-    \label{eq:insample_softmax}
-\end{align}
-By choosing the maximizer policy the above equation takes the form:
-\begin{align}
-    \datasetOptimalQ = r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{\tau \ln \!\!\!\!\!\sum_{a': \datasetPolicy(a'|s') > 0} \!\!\!\!\! \exp\AdaBracket{\tau^{-1} \datasetOptimalQ(s',a')}}.
+    \datasetOptimalQ (s,a) = r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{\tau \ln \!\!\!\!\!\sum_{a': \datasetPolicy(a'|s') > 0} \!\!\!\!\!
+\exp\AdaBracket{\tau^{-1} \datasetOptimalQ(s',a')}}.
 \end{align}
@@
 \end{align}
-The expectation is approximated by Monte-Carlo sampling actions from the dataset.
+The expectation can be approximated by Monte-Carlo sampling actions from the dataset.
 Since the term $\exp\AdaBracket{\tau^{-1}\datasetOptimalQ (s',a')}$ appears also in the regularized softmax policy, in-sample softmax updates the policy towards 
@@
 
-We interpret \eq{\ref{eq:inac_policy}} from another perspective.
-the in-sample softmax policy can be decomposed into two terms:
-the first term $\datasetPolicy(a|s) \exp\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,a)}{\tau} }$ acts as a KL-regularized policy with respect to the behavior policy (see \eq{\ref{eq:average}}); 
-the second term $\exp\AdaBracket{-\ln\hat{\pi}_{\mathcal{D}}(a|s)}$ can be seen as induced by another regularization $\Omega(\pi) = -\AdaAngleProduct{\pi}{\ln{\hat{\pi}_{\mathcal{D}}}}$, the cross entropy between the in-sample softmax policy and the behavior policy.
+The benefits of the in-sample softmax can be interpreted as (1) softmax is easier to compute than hard max in the continuous action setting; 
+(2) the in-sample softmax policy \eq{\ref{eq:inac_policy}} can be seen as a KL-regularized policy and hence the dataset support constraint is satisfied.
+In fact, the support of $ \pi_{\datasetPolicy, k+1}$ should be exactly the same as $\datasetPolicy$ since $\exp(\frac{\cdot}{\tau}) > 0$ as long as $\tau \neq \infty$.
+We now explain why \eq{\ref{eq:inac_policy}} can be regarded as a KL-regularized policy: it can be decomposed into two terms:
+the first term $\datasetPolicy(a|s) \exp\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,a)}{\tau} }$ acts as a KL-regularized policy with respect to the behavior policy; 
+the second term $\exp\AdaBracket{-\ln\hat{\pi}_{\mathcal{D}}(a|s)}$ can be seen as induced by another regularization $-\sum_{a}{\pi(a|s)}{\ln{\hat{\pi}_{\mathcal{D}}(a|s)}}$, which is the cross entropy between the in-sample softmax policy and the (learned) behavior policy.
 
-\section{Forward-Backward Tsallis Learning for Offline RL}
+However, the fact that the softmax policies always have full support indicates there is a persistent gap to \eq{\ref{eq:hardmax_offline}}.
+Since we want to improve upon the behavior policy, it is expected that action candidates should gradually narrow down to the maximizer.
+Furthermore, since \eq{\ref{eq:inac_policy}} is only implicitly regularizing the policy and no explicit KL regularization for the value estimation \cite{Wu2020-BehaviorRegularizedAC}, there is no guarantee that the learned policy should be close to the behavior policy.
 
-Summary:
-\begin{itemize}
-    \item Forward: Tsallis sparsemax policies generate the dataset.
-This assumption is not very strong since Tsallis entropy generalizes Shannon entropy and the behavior policy class is sufficiently rich.
-    \item Backward: learn the optimal policy with the modified Tsallis entropy regularization. The learned policy differs from the conventional sparsemax policies in that no thresholding is required.
-\end{itemize}
+\subsection{In-Sample Tsallis Regularization}
 
-The key to both in-sample max \cite{Fujimoto2019-InSampleMax} and in-sample softmax \cite{Xiao2023-InSampleSoftmax} is the dataset support constraint.
-In this paper, we propose to exploit the properties of sparsemax policies to naturally satisfy this constraint.
-In a nutshell, we assume the behavior policies are Tsallis sparsemax policies, and therefore the actions not present in the dataset are with low probability and are exactly these truncated actions.
-For simplicity, suppose the dataset is generated by the Tsallis sparsemax policy:
+We propose to replace the Shannon entropy in in-sample softmax to the Tsallis entropy.
+The seemingly simple replacement, however, leads to drastically different behavior of the policy.
+Indeed, since we can control the truncation effect of the Tsallis regularized policies, we have the support of $\pi^{\text{Tsallis}}_{\datasetPolicy, k+1} \preceq \datasetPolicy$, i.e.
+the support of the learned policy is either equal or within the dataset support.
+
+Furthermore, Tsallis policies provide an means to formulate the common assumption that actions not present in the dataset are with low probability of being selected.
+If we assume the behavior policy is also a Tsallis policy, then the support constraint $a:\datasetPolicy(a|s)>0$ can be naturally replaced to the truncation criterion $a\in K(s)$; i.e.,
 \begin{align}
-    \datasetPolicy(a|s) = \AdaRectBracket{\frac{Q_{\datasetPolicy}(s,a)}{\tau_\mathcal{D}} - \psi\AdaBracket{\frac{Q_{\datasetPolicy}(s, \cdot)}{\tau_\mathcal{D}}}}_{+},  \quad \sum_{a\in K_\mathcal{D}(s)} \pi_\mathcal{D}(a|s)= 1,
+    \datasetPolicy(a|s) \propto \exp_q\AdaBracket{\frac{Q_{\datasetPolicy}(s,a)}{\tau_\mathcal{D}}},  \quad \sum_{a\in K_\mathcal{D}(s)} \pi_\mathcal{D}(a|s)= 1,
 \end{align}
-where $\tau_\mathcal{D}$ is a unknown coefficient and $K_\mathcal{D}(s)$ denotes the set of allowable actions.
-Now let us inspect the Bellman equation for the behavior policy:
-\begin{align}
-    \begin{split}
-    Q_{\datasetPolicy}(s,a) &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \sum_{a'} \datasetPolicy(a'|s') \AdaBracket{Q_{\datasetPolicy}(s',a') - \tau_{\mathcal{D}} S_2(\datasetPolicy)(\cdot|s')}}.
+where $\tau_\mathcal{D}$ is an unknown coefficient and $K_\mathcal{D}(s)$ denotes the set of actions present in the dataset.
+% Now let us inspect the Bellman equation for the behavior policy:
+% \begin{align}
+%     \begin{split}
+%     Q_{\datasetPolicy}(s,a) &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \sum_{a'} \datasetPolicy(a'|s') \AdaBracket{Q_{\datasetPolicy}(s',a') - \tau_{\mathcal{D}} S_2(\datasetPolicy)(\cdot|s')}}.
 \\
-    &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \tau_{\mathcal{D}} \AdaBracket{ \sum_{a' \in K_{\mathcal{D}}(s')} \AdaBracket{\frac{Q_{\datasetPolicy}(s',a')}{\tau_{\mathcal{D}}}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy}(s', \cdot)}{\tau_{\mathcal{D}}}}^2 + 1 }}\\
-    &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \tau_{\mathcal{D}} \AdaBracket{ \sum_{a': \datasetPolicy(a'|s') > 0} \AdaBracket{\frac{Q_{\datasetPolicy}(s',a')}{\tau_{\mathcal{D}}}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy}(s', \cdot)}{\tau_{\mathcal{D}}}}^2 + 1 }}.
-    \label{eq:insample_sparsemax}
-    \end{split}
-\end{align}
-Here, $\max_{\pi \in \Delta(\mathcal{A})}\pi(a|s)Q(s,a) + \tau S_2(\pi(\cdot|s)) $ attains its maximum at $V(s) = \tau \sum_{a \in K(s)}\AdaBracket{ \frac{Q(s,a)}{\tau} }^2 - \psi\AdaBracket{\frac{Q(s, \cdot)}{\tau}}^2 + \tau$ \cite{Lee2018-TsallisRAL,Martins16-sparsemax}.
-It is also possible to assume that $\datasetPolicy(a|s) = \exp_q\AdaBracket{\frac{Q_{\datasetPolicy}(s,a)}{\tau_\mathcal{D}} - \tilde{\psi}\AdaBracket{\frac{Q_{\datasetPolicy}(s, \cdot)}{\tau_\mathcal{D}}}}$ for general $q>2$, though in these cases the policy does not have a closed-form expression we have to resort to approximation.
+%     &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \tau_{\mathcal{D}} \AdaBracket{ \sum_{a' \in K_{\mathcal{D}}(s')} \AdaBracket{\frac{Q_{\datasetPolicy}(s',a')}{\tau_{\mathcal{D}}}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy}(s', \cdot)}{\tau_{\mathcal{D}}}}^2 + 1 }}\\
+%     &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \tau_{\mathcal{D}} \AdaBracket{ \sum_{a': \datasetPolicy(a'|s') > 0} \AdaBracket{\frac{Q_{\datasetPolicy}(s',a')}{\tau_{\mathcal{D}}}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy}(s', \cdot)}{\tau_{\mathcal{D}}}}^2 + 1 }}.
+%     \label{eq:insample_sparsemax}
+%     \end{split}
+% \end{align}
+% Here, $\max_{\pi \in \Delta(\mathcal{A})}\pi(a|s)Q(s,a) + \tau S_2(\pi(\cdot|s)) $ attains its maximum at $V(s) = \tau \sum_{a \in K(s)}\AdaBracket{ \frac{Q(s,a)}{\tau} }^2 - \psi\AdaBracket{\frac{Q(s, \cdot)}{\tau}}^2 + \tau$ \cite{Lee2018-TsallisRAL,Martins16-sparsemax}.
+% It is also possible to assume that $\datasetPolicy(a|s) = \exp_q\AdaBracket{\frac{Q_{\datasetPolicy}(s,a)}{\tau_\mathcal{D}} - \tilde{\psi}\AdaBracket{\frac{Q_{\datasetPolicy}(s, \cdot)}{\tau_\mathcal{D}}}}$ for general $q>2$, though in these cases the policy does not have a closed-form expression we have to resort to approximation.
 Under the Tsallis behavior policy assumption, the dataset support constraint $a: \datasetPolicy(a|s) > 0$ coincides with the condition $a \in K_{\mathcal{D}}(s)$.
-Moreover, the assumption actually assumes a forward process of learning a set of allowable actions $K_\mathcal{D}(s)$ given $Q_{\datasetPolicy}(s,a), \forall s,a$;
- and it naturally raises the possibility of \emph{backward Tsallis learning}: given the set of allowable actions $K_\mathcal{D}(s)$, learn an optimal policy that resembles $\datasetPolicy$ in the sense of $\pi_* \preceq \datasetPolicy$, i.e.
+Revisiting the in-sample hard-max Bellman equation \eq{\ref{eq:hardmax_offline}}:
+\begin{align*}
+    \datasetOptimalQ(s,a) = r(s,a) + \gamma \expectation{s'\sim P(\cdot | s,a)}{\max_{a'\in K_{\mathcal{D}}(s)} \datasetOptimalQ(s',a')}.
+\end{align*}
+However, the assumption of the Tsallis behavior policy alone is not useful.
+% Moreover, the assumption actually assumes a forward process of learning a set of allowable actions $K_\mathcal{D}(s)$ given $Q_{\datasetPolicy}(s,a), \forall s,a$;
+%  and it naturally raises the possibility of \emph{backward Tsallis learning}: given the set of allowable actions $K_\mathcal{D}(s)$, learn an optimal policy that resembles $\datasetPolicy$ in the sense of $\pi_* \preceq \datasetPolicy$, i.e.
 putting probability mass only on the actions in $K_\mathcal{D}$.
-
- The backward Tsallis learning consists in learning $\pi_{t, \datasetPolicy} \preceq \datasetPolicy$ given $K_\mathcal{D}$.
+The power of the assumption manifests when we use also Tsallis regularized policy learning:
 % However, it should be noted that the learned policy no longer requires a thresholding function $\tilde{\psi}$ since the set of allowable actions $K_\mathcal{D}$ is given, hence no action need to be explicitly truncated.
@@
 \begin{align}
-    \pi_{t+1, \datasetPolicy}(a|s) = \exp_q\AdaBracket{\frac{Q_{t, \datasetPolicy}(s,a)}{\tau} - \tilde{\psi}{\AdaBracket{\frac{Q_{t, \datasetPolicy}(s,\cdot)}{\tau}}}},  \quad \sum_{a\in K_{\mathcal{D}}(s)} \pi_{t+1, \datasetPolicy} = 1.
+    & \pi_{t+1, \datasetPolicy}(a|s) \propto \exp_q\AdaBracket{\frac{Q_{t, \datasetPolicy}(s,a)}{\tau}},  \quad \sum_{a\in K_{\mathcal{D}}(s)} \pi_{t+1, \datasetPolicy} = 1.\\
+    Q_{t+1, \datasetPolicy}(s,a) &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \sum_{a' \in K_{\mathcal{D}}(s)} \!\!\!\!\!
+\pi_{t}(a'|s') \AdaBracket{Q_{t, \datasetPolicy}(s',a') - \tau S_q(\pi_{t}(\cdot|s'))}}.
     \label{eq:proposal_policy}
 \end{align}
-The thresholding function together with the given $K_\mathcal{D}$ implies during the backward learning the new set of allowable actions $K_{t, q}$, where we make clear the dependence on $q$ and iteration $t$.
-The set should satisfy the condition $K_{t, q} \preceq K_\mathcal{D}$:
- i.e. the cardinality $|K_{t, q}| \leq |K_\mathcal{D}|$ and support constraint $\pi_t \preceq \datasetPolicy$.
-Another advantage lies in that for $q>1$, the policy is a variant of categorical distributions which is less susceptible to numerical issues than exponential functions \cite{Tsai2021-selfsupervisedRelativePredictiveCoding}.
-
+The learned Tsallis policy maintains a new set of allowable actions $K_{t, q}$ depending on $q$ and iteration $t$.
+The set satisfies the condition $K_{t, q} \preceq K_\mathcal{D}$:
+ i.e. $|K_{t, q}| \leq |K_\mathcal{D}|$ and support constraint $\pi_t \preceq \datasetPolicy$.
 % That is, the backward Bellman optimality equation becomes:
@@
 
-\textbf{Why use Tsallis backward learning? }
-If the Tsallis behavior policy assumption holds, and Tsallis backward learning is adopted, then the KL divergence between the learned policy and the behavior policy can be uppper-bounded.
-That being said, the learned policy is guaranteed to better respect the behavior policy.
+We now show that, if the Tsallis behavior policy assumption holds and we use Tsallis regularized policy learning, then the learned policy is guaranteed to stay close to the behavior policy in the sense that their KL divergence is uppper-bounded.
 \begin{theorem}
-    Suppose the dataset $\mathcal{D}$ is generated by a Tsallis behavior policy of any entropic index $q^*$ (it is not important).
+    Suppose the dataset $\mathcal{D}$ is generated by a Tsallis behavior policy of entropic index $q$.
     Let $K_{t, q}(s) \preceq K_\mathcal{D}(s)$ denote the set of allowable actions at $t$-th iteration whose cardinality is smaller than $K_\mathcal{D}(s)$.
     Also let $\pi_{t}(a|s) \propto \exp_q\AdaBracket{\frac{Q_{t-1}(s,a)}{\tau}}$ denote the learned policy. 
-    Then the KL divergence between the $\pi_t$ and the behavior policy can be upper bounded:
+    Then the KL divergence between $\pi_t$ and the behavior policy can be upper bounded:
 \begin{align}
     \begin{split}
-    &\KLany{\pi_t(\cdot|s)}{\datasetPolicy(\cdot|s)}  \leq  |K_{t}(s)|\!
+    &\KLany{\pi_t(\cdot|s)}{\datasetPolicy(\cdot|s)}  \leq  |K_{t,q}(s)|\!
 \AdaRectBracket{  { \pi_{t}^{q}(a|s) - \pi_{t}^{q-1} (a|s) } + \pi_t(a|s) - \frac{q-3}{q-1}  + \datasetPolicy^{q-2}(a|s) - \datasetPolicy^{q-1}(a|s)}.
@@
 \end{theorem}
-The bound is suggesting.
-It explains why using in-sample softmax $q=1$ may not be a good choice as in this case the KL divergence to the behavior policy is unbounded.
-On the other hand, choosing $q=2$ which recovers sparsemax is bounded by $K_{t, q}(s)\AdaBracket{\pi_t(a|s) - \pi_\mathcal{D}(a|s) + 2}$.
+The bound is suggestive.
+For $q=1$ (the in-sample softmax case) the bound is not useful and simply states the KL divergence may be unbounded.
+On the other hand, choosing any $q>1$ brings an upper bound of at most $4 |K_{t,q}(s)|$.
+When  $q=2$, the in-sample sparsemax has KL divergence to the behavior policy bounded by $|K_{t, q}(s)|\AdaBracket{\pi_t(a|s) - \pi_\mathcal{D}(a|s) + 2}$.
 However, it should be noted that there is a trade-off between the power of policies $\pi^{q}$ and the cardinality of $K_{t,q}(s)$: $K_{t,q}(s)$ tends to collect all actions when $q \rightarrow \infty$.
 \begin{proof}
-    Before we state the main result, we prove the following lemma:
+    We first prove the following lemma:
 \begin{lemma}\label{lemma:log_qlog_diff}
@@
 \end{align}
-where we leveraged $\ln x \leq x - 1$ and  the fact that $\forall a\in K(s), \,\, -\frac{1}{q-1} \leq \frac{Q_{t-1}(s,a)}{\tau} - \tilde{\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}} \leq 0$. 
-If $a \notin K(s)$, then $\ln\pi_t(a|s) = -\infty$ and the KL term is unbounded.
+where we leveraged $\ln x \leq x - 1$ and  $\ln_q\exp_q(x) = x$ both when $x>0$.
+Considering the definition of $\pi_t(a|s) \propto \exp_q\AdaBracket{\frac{Q_{t-1}(s,a)}{\tau} - {\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}}}$ and $\exp_q(x) = \AdaRectBracket{1 + (q-1)x}_{+}^{\frac{1}{q-1}}$, we must have  $\pi_t(a|s) > 0 \Leftrightarrow a\in K_{t-1,q}(s) \Leftrightarrow -\frac{1}{q-1} \leq \frac{Q_{t-1}(s,a)}{\tau} - {\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}} \leq 0$.
+If $a \notin K_{t-1,q}(s)$, then $\ln\pi_t(a|s) = -\infty$ and the KL term is unbounded.
 The same fact is exploited to yield an upper bound $\frac{1}{q-1}$  for (2).
@@
 
-\textbf{Difference with In-sample Softmax. }
-Our proposal is seemingly similar to in-sample softmax \cite{Xiao2023-InSampleSoftmax}.
-However, the two methods are radically different: in-sample softmax requires satisfying the dataset support constraint during learning to alleviate the out-of-distribution error, and is motivated by the observation that taking softmax is easier than the hard max operator \eq{\ref{eq:hardmax_offline}}.
-On the other hand, we assume the behavior policies are Tsallis policies and as such the actions not present in the dataset are those being truncated.
-This viewpoint leads to the natural satisfaction of the dataset support constraint.
+The assumption of Tsallis behavior policy is not restrictive.
+If all actions present, then it corresponds to the case of $q=1$ where the policy has full support.
+On the other hand, different levels of missingness can be simulated by both $q$ and $\tau$.
 
-Furthermore, the assumption naturally leads to the choice of Tsallis regularized backward learning which recovers in-sample softmax as a special case when $q=1$.
-Since softmax policies have full support, there always exists non-negligible probabilities for suboptimal actions; while for \eq{\ref{eq:proposal_policy}} the optimal policy can switch between greedy policy and multimodal policy.
+Another advantage of \eq{\ref{eq:proposal_policy}} is that for $q>1$, the policy is a variant of categorical distributions which is less susceptible to numerical issues than exponential functions \cite{Tsai2021-selfsupervisedRelativePredictiveCoding}.
+Furthermore, \eq{\ref{eq:proposal_policy}} can switch between greedy policy and multimodal policy.
 The former is achieved when all but one action have values lower than the threshold; while vice versa for the latter.
+
+\textbf{In-sample Tsallis Policy Estimation}
+Similar to \cite{Xiao2023-InSampleSoftmax}, we want to directly use the actions from the dataset to estimate our policy.
+We do a similar step to \eq{\ref{eq:inac_policy}}:
+\begin{align}
+    \begin{split}
+        &\pi_{\datasetPolicy, k+1} (a|s) \propto \exp_q \AdaBracket{\frac{1}{\tau} Q_{t, \datasetPolicy}(s,a)} = \datasetPolicy(a|s) \datasetPolicy(a|s)^{-1} \exp_q\AdaBracket{\frac{1}{\tau}Q_{t, \datasetPolicy}(s,a)} \\
+        &= \datasetPolicy(a|s)  \exp_q{\AdaBracket{\ln_q{\frac{1}{\datasetPolicy(a|s)}}}} \exp_q\AdaBracket{\frac{1}{\tau}Q_{t, \datasetPolicy}(s,a)}\\
+        &=  \datasetPolicy(a|s) {\AdaBracket{ \exp_q\AdaBracket{\frac{1}{\tau}Q_{t, \datasetPolicy}(s,a) + \ln_2\frac{1}{\datasetPolicy(a|s)}}^{q-1} - (q-1)^2 \frac{1}{\tau}Q_{t, \datasetPolicy}(s,a)  \ln_q{\frac{1}{\datasetPolicy(a|s)}} }^{\frac{1}{q-1}}}.
+        \label{eq:tsallis_inac_policy_thoery}
+    \end{split}
+\end{align}
+In the last step we made use of the relationship $\AdaBracket{\expq{x}\cdot \expq{y}}^{q-1} = \expq{\AdaBracket{x+y}}^{q-1} + (q-1)^2 xy$ \cite{Yamano2004-properties-qlogexp}.
 
@@
     One of the main reasons is the sparsemax policies are not suitable for online RL since the exploration is handicapped resulted from the action truncation.
-    However, in offline RL this drawback vanishes, and theoretically it provides a better in-sample algorithm that the constraint is satisfied implicitly. 
+    However, in offline RL this drawback vanishes, and theoretically it allows for bounding the distance to the behavior policy, guaranteeing less OOD actions.
 \end{remark}
@@
 We follow the same setup here, but replacing every appearance of $\ln, \exp$ to their \qlog and $q$-exponential counterpart.
-In practice, computing the thresholding function $\psi$ requires sorting in the discrete control setting and it remains unclear how to accurately estimate it for the continuous control problems.
-We found it improves the performance by simply removing $\psi$ and renormalize the policy.
-In this case, thresholding can still be achieved by simply learning $Q_{\theta} \rightarrow 0$.
-That is, the learned policy form becomes $\pi_{t+1}(a|s) \propto \AdaRectBracket{Q_{\theta}(s,a)}_{+}$.
-In order to also sample actions from the dataset when updating the policy, we repeat the derivation in \eq{\ref{eq:insample_policy}} but for the sparsemax policy.
-% Let us define $q_{\theta}(s,a) := \frac{Q_{\datasetPolicy, k}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,\cdot)}{\tau}}$ for convenience:
-\begin{align}
-    \begin{split}
-        \pi_{\datasetPolicy, k+1} (a|s) &= \datasetPolicy(a|s) \datasetPolicy(a|s)^{-1} \exp_q{Q_{\theta}(s,a)} \\
-        &= \datasetPolicy(a|s)  \exp_q{\AdaBracket{\ln_q{\frac{1}{\datasetPolicy(a|s)}}}} \exp_q{Q_{\theta}(s,a)}\\
-        &=  \datasetPolicy(a|s) \sqrt[q-1]{\AdaBracket{ \exp_q\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\datasetPolicy(a|s)}}^{q-1} - (q-1)^2 Q_{\theta}(s,a)  \ln_q{\frac{1}{\datasetPolicy(a|s)}} }}.
-        \label{eq:tsallis_inac_policy}
-    \end{split}
-\end{align}
-In the last step we made use of the relationship $\AdaBracket{\expq{x}\cdot \expq{y}}^{q-1} = \expq{\AdaBracket{x+y}}^{q-1} + (q-1)^2 xy$ \cite{Yamano2004-properties-qlogexp}.
+% In practice, computing the thresholding function $\psi$ requires sorting in the discrete control setting and it remains unclear how to accurately estimate it for the continuous control problems.
+% We found it improves the performance by simply removing $\psi$ and renormalize the policy.
+% In this case, thresholding can still be achieved by simply learning $Q_{\theta} \rightarrow 0$.
+% That is, the learned policy form becomes $\pi_{t+1}(a|s) \propto \AdaRectBracket{Q_{\theta}(s,a)}_{+}$.
+% In order to also sample actions from the dataset when updating the policy, we repeat the derivation in \eq{\ref{eq:insample_policy}} but for the sparsemax policy.
+% % Let us define $q_{\theta}(s,a) := \frac{Q_{\datasetPolicy, k}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,\cdot)}{\tau}}$ for convenience:
+% \begin{align}
+%     \begin{split}
+%         \pi_{\datasetPolicy, k+1} (a|s) &= \datasetPolicy(a|s) \datasetPolicy(a|s)^{-1} \exp_q{Q_{\theta}(s,a)} \\
+%         &= \datasetPolicy(a|s)  \exp_q{\AdaBracket{\ln_q{\frac{1}{\datasetPolicy(a|s)}}}} \exp_q{Q_{\theta}(s,a)}\\
+%         &=  \datasetPolicy(a|s) \sqrt[q-1]{\AdaBracket{ \exp_q\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\datasetPolicy(a|s)}}^{q-1} - (q-1)^2 Q_{\theta}(s,a)  \ln_q{\frac{1}{\datasetPolicy(a|s)}} }}.
+%         \label{eq:tsallis_inac_policy}
+%     \end{split}
+% \end{align}
+% In the last step we made use of the relationship $\AdaBracket{\expq{x}\cdot \expq{y}}^{q-1} = \expq{\AdaBracket{x+y}}^{q-1} + (q-1)^2 xy$ \cite{Yamano2004-properties-qlogexp}.
 Similar to the discussion after \eq{\ref{eq:inac_policy}}, the Tsallis policy we derived here can be seen as the result from Tsallis KL regularization, plus another regularization that gives rise to the additional term inside the $\exp_q$ function.
-In the implementation, we choose the sparsemax parametrization, which gives the following Tsallis in-sample sparsemax actor-critic update rule:
+In the implementation, we choose the sparsemax parametrization $q=2$, which gives the following Tsallis in-sample sparsemax actor-critic update rule:
 % In the discrete case, the truncation threshold or the normalization function $\psi$ is computed by sorting the action values.
@@
     % & Q_{\theta}(s,a) \leftarrow \frac{Q_{\theta}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\theta}(s,\cdot)}{\tau}},\\
-    &\mathcal{L}_\text{actor}(\phi) = -\expectation{s,a \sim \mathcal{D}}{ \AdaBracket{ \exp_q\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\pi_{\omega}(a|s)}} -  Q_{\theta}(s,a)  \logq{\frac{1}{\pi_{\omega}(a|s)}} } \ln\pi_{\phi}(a|s) }, \\
+    &\mathcal{L}_\text{actor}(\phi) = -\expectation{s,a \sim \mathcal{D}}{ \AdaBracket{ \exp_2\AdaBracket{\frac{1}{\tau}Q_{\theta}(s,a) + \ln_2\frac{1}{\pi_{\omega}(a|s)}} -  \frac{1}{\tau}Q_{\theta}(s,a)  \logq{\frac{1}{\pi_{\omega}(a|s)}} } \ln\pi_{\phi}(a|s) }, \\
     & \mathcal{L}_\text{baseline}(\zeta) = \expectation{s\sim\mathcal{D}, a\sim \pi_{\phi}(s)}{\AdaBracket{ v_{\zeta}(s) - Q_{\theta}(s,a) - \tau\ln_2\pi_{\phi}(a|s)}^2},\\
@@
 
+\clearpage
+
+\appendix
+\section{Derivation}\label{apdx:derivation}
+
+It is worth noting that, for $q\neq 1$ the regularized policy is given by 
+\begin{align}
+    \pi^{*}(a|s) = \sqrt[\leftroot{-2}\uproot{16}q-1]{\AdaRectBracket{\frac{Q_{\pi}(s,a)}{q\tau}  - \psi\AdaBracket{\frac{Q_{\pi}(s,\cdot)}{\tau}} }_{+}  (q-1) },
+    \label{eq:approximate_policy}
+\end{align}
+the normalization function $\psi$ can only be analytically solved when $q=2, \infty$.  
+When $q\neq 1,2$, the closed-form expression of $\pi, \psi$ might not exist.
+Following \citep{chen2018-TsallisApproximate}, we leverage the first order Taylor expansion $f(z) + f'(z)(x-z)$ on the policy \eq{\ref{eq:approximate_policy}},
+where we let $z=1$, $x= \AdaRectBracket{\frac{Q_{\pi}(s,a)}{q\tau}  - \psi\AdaBracket{\frac{Q_{\pi}(s,\cdot)}{q\tau}} }_{+}  \frac{q-1}{p} $,  $f(x) = x^{\frac{1}{q-1}}$, $f'(x) = \frac{1}{q-1}x^\frac{2-q}{q-1}$.
+So that
+\begin{align}
+    \begin{split}
+        \tilde{\pi}^{*}(a|s) &\approx f(z) + f'(z)(x-z) \\
+        & = 1 + \frac{1}{q-1}\AdaBracket{\AdaBracket{\frac{Q_{\pi}(s,a)}{q\tau} - \tilde{\psi}\AdaBracket{\frac{Q_{\pi}(s,\cdot)}{q\tau}} } \frac{q-1}{p}  - 1} .
+    \end{split}
+    \label{eq:apdx_approximate_policy}
+\end{align}
+
 \end{document}

### 2023-04-22 paper/main.tex
added more details

@@
 
-We can write out the regularized policy for general Tsallis entropy regularization for any $q$\footnotemark\footnotetext{Different to prior works \cite{Lee2018-TsallisRAL,Nachum18a-tsallis}, our definition of $\psi$ has an additional $\frac{1}{q-1}$ term in order to be accommodated in the $q$-exponential function, see Appendix \ref{apdx:derivation} for derivation.}:
+We can write out the regularized policy for general Tsallis entropy regularization for any $q$\footnotemark\footnotetext{Different to prior works \cite{Lee2018-TsallisRAL,Nachum18a-tsallis}, our definition of $\psi$ has an additional $\frac{1}{q-1}$ term to accommodate the $q$-exponential policy, see Appendix \ref{apdx:derivation} for derivation.}:
 \begin{align}
@@
 Instead, many of the existing works propose to force the learned policy to be close to the behavior policy \cite{Dadashi2021-pseudoMetricOffline,Fujimoto2019-InSampleMax,Fujimoto2021-minimalist,Nair2021-awac}.
-This is often achieved by adding a regularizer penalizing deviating from the behavior policy \cite{Ghasemipour2021-EMaQ,Kumar2019-BootstrapErrorQlearningMMD,Jaques2020-OfflineDialog,Wu2020-BehaviorRegularizedAC}, e.g.
-incorporating the KL penalty $\tau\KLany{\pi(\cdot|s)}{\datasetPolicy(\cdot|s)}$ leads to the policy form $\pi(a|s)\propto\datasetPolicy(a|s)\exp\AdaBracket{\tau^{-1}Q_{\pi}(s,a)}$ where the learned policy must have the same support of the behavior policy.
+This is often achieved by adding a regularizer in the policy update penalizing deviating from the behavior policy \cite{Ghasemipour2021-EMaQ,Kumar2019-BootstrapErrorQlearningMMD,Jaques2020-OfflineDialog,Wu2020-BehaviorRegularizedAC}, e.g.
+by incorporating KL divergence $\max_{\pi}\expectation{s\sim\mathcal{D}}{\expectation{a\sim\pi(\cdot|s)}{Q(s,a)} - \tau\KLany{\pi(\cdot|s)}{\datasetPolicy(\cdot|s)}}$.
+The critic is typically updated by minimizing the loss $\mathbb{E}_{(s,a,s')\sim \mathcal{D}}\expectation{a\sim\pi(\cdot|s)}{\AdaBracket{ r(s,a) + \gamma Q(s',a') - Q(s,a)}^2}$.
+The regularization leads to the policy form $\pi(a|s)\propto\datasetPolicy(a|s)\exp\AdaBracket{\tau^{-1}Q_{\pi}(s,a)}$ where the learned policy must have the same support of the behavior policy.
 This idea is shared by the \emph{in-sample} methods that enforce the support of the learned policy to be within the behavior policy $\pi \preceq \datasetPolicy$ \cite{Fujimoto2019-InSampleMax,Kostrikov2022-implicitQlearning,Xiao2023-InSampleSoftmax}, which is achieved by only updating for the actions in the dataset.
@@
 \begin{align}
-    \pi_{\datasetPolicy, k+1} \propto  \datasetPolicy(a|s) \exp\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,a)}{\tau} - \ln\hat{\pi}_{\mathcal{D}}(a|s)},
+    \pi_{t+1,\datasetPolicy} \propto  \datasetPolicy(a|s) \exp\AdaBracket{\frac{Q_{t, \datasetPolicy}(s,a)}{\tau} - \ln\hat{\pi}_{\mathcal{D}}(a|s)},
     \label{eq:inac_policy}
@@
 We now explain why \eq{\ref{eq:inac_policy}} can be regarded as a KL-regularized policy: it can be decomposed into two terms:
-the first term $\datasetPolicy(a|s) \exp\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,a)}{\tau} }$ acts as a KL-regularized policy with respect to the behavior policy; 
+the first term $\datasetPolicy(a|s) \exp\AdaBracket{\frac{Q_{t, \datasetPolicy}(s,a)}{\tau} }$ acts as a KL-regularized policy with respect to the behavior policy; 
 the second term $\exp\AdaBracket{-\ln\hat{\pi}_{\mathcal{D}}(a|s)}$ can be seen as induced by another regularization $-\sum_{a}{\pi(a|s)}{\ln{\hat{\pi}_{\mathcal{D}}(a|s)}}$, which is the cross entropy between the in-sample softmax policy and the (learned) behavior policy.
@@
 
-Furthermore, Tsallis policies provide an means to formulate the common assumption that actions not present in the dataset are with low probability of being selected.
+Furthermore, Tsallis policies provide an means to formulate the common assumption that actions not present in the dataset are with low probability of being selected \cite{Kostrikov2021-fisherCriticReg}.
 If we assume the behavior policy is also a Tsallis policy, then the support constraint $a:\datasetPolicy(a|s)>0$ can be naturally replaced to the truncation criterion $a\in K(s)$; i.e.,
@@
     Q_{t+1, \datasetPolicy}(s,a) &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \sum_{a' \in K_{\mathcal{D}}(s)} \!\!\!\!\!
-\pi_{t}(a'|s') \AdaBracket{Q_{t, \datasetPolicy}(s',a') - \tau S_q(\pi_{t}(\cdot|s'))}}.
+\pi_{t+1}(a'|s') \AdaBracket{Q_{t, \datasetPolicy}(s',a') - \tau S_q(\pi_{t+1}(\cdot|s'))}}.
     \label{eq:proposal_policy}
 \end{align}
-The learned Tsallis policy maintains a new set of allowable actions $K_{t, q}$ depending on $q$ and iteration $t$.
+The term inside the bracket is known as $q$-maximum \cite{Lee2020-generalTsallisRSS}.
+Similar to the softmax operators \cite{asadi17a,azar2012dynamic}, $q$-maximum is an approximation to the maximum with degree controlled by $q$.
+
+More importantly, the support constraint is naturally satisfied since 
+the learned Tsallis policy learns a new set of allowable actions $K_{t, q}$ from $K_\mathcal{D}(s)$ depending on $q$ and iteration $t$.
 The set satisfies the condition $K_{t, q} \preceq K_\mathcal{D}$:
  i.e. $|K_{t, q}| \leq |K_\mathcal{D}|$ and support constraint $\pi_t \preceq \datasetPolicy$.
+ Let us take $q=2$ for example:
+ \begin{align}
+    \begin{split}
+    &Q_{t+1, \datasetPolicy}(s,a) = r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \sum_{a' \in K_{\mathcal{D}}(s)} \!\!\!\!\!
+\pi_{t+1}(a'|s') \AdaBracket{Q_{t, \datasetPolicy}(s',a') - \tau S_2(\pi_{t+1}(\cdot|s'))}}\\
+    &\qquad = r(s,a) + \gamma \expectation{s'\sim P(\cdot | s,a)}{ \tau\sum_{a' \in K_{\mathcal{D}}(s)} \!\!
+\AdaBracket{\frac{Q_{t, \datasetPolicy}(s',a')}{\tau}}^2 - \AdaBracket{ \psi\AdaBracket{\frac{Q_{t, \datasetPolicy}(s',a')}{\tau}} - 1 }^2 + \tau}\\
+    &\qquad = r(s,a) + \gamma \expectation{s'\sim P(\cdot | s,a)}{ \tau\sum_{a' \in K_{t,q}(s)} \!\!
+\AdaBracket{\frac{Q_{t, \datasetPolicy}(s',a')}{\tau}}^2 - \AdaBracket{ \psi\AdaBracket{\frac{Q_{t, \datasetPolicy}(s',a')}{\tau}} - 1 }^2 + \tau}.
+    \end{split}
+    \label{eq:tsallis_backward_q2}
+ \end{align}
+ The second equation is because $\max_{\pi}\sum_{a\in\pi(\cdot|s)}\pi(a|s)Q(s,a) - \tau S_2(\pi(\cdot|s))$ attains its maximum at $\tau \sum_{a \in K(s)}\AdaBracket{ \frac{Q(s,a)}{\tau} }^2 - \psi\AdaBracket{\frac{Q(s, \cdot)}{\tau}}^2 + \tau$ \cite{Lee2018-TsallisRAL,Martins16-sparsemax}, and the last equation is due to $K_{t, q} \preceq K_\mathcal{D}$.
+ \eq{\ref{eq:tsallis_backward_q2}} states that, one needs not to explicitly enforce the support constraint since it is automatically fulfilled by learning a new subset of allowable actions.
 % That is, the backward Bellman optimality equation becomes:
@@
 \begin{align*}
-    \KLany{\pi_t(\cdot|s)}{\datasetPolicy(\cdot|s)} \leq K_{t}(s)\!
-\AdaRectBracket{  { \pi_{t}^{q}(a|s) - \pi_{t}^{q-1} (a|s) } + \pi_t(a|s) - \frac{q-3}{q-1}  + \datasetPolicy^{q-2}(a|s) - \datasetPolicy^{q-1}(a|s)}.
+    \KLany{\pi_t(\cdot|s)}{\datasetPolicy(\cdot|s)} &\leq \!\!\!\!\!\!
+\sum_{a\in K_{t,q}(s)}\!\!\!\!\!
+\pi_t(a|s)\AdaRectBracket{  { \pi_{t}^{q}(a|s) - \pi_{t}^{q-1} (a|s) } + \pi_t(a|s) - \frac{q-3}{q-1}  + \datasetPolicy^{q-2}(a|s) - \datasetPolicy^{q-1}(a|s)} \\
+    &\leq K_{t,q}(s)\!\AdaRectBracket{  { \pi_{t}^{q}(a|s) - \pi_{t}^{q-1} (a|s) } + \pi_t(a|s) - \frac{q-3}{q-1}  + \datasetPolicy^{q-2}(a|s) - \datasetPolicy^{q-1}(a|s)}.
 \end{align*}
@@
     \begin{split}
-        &\pi_{\datasetPolicy, k+1} (a|s) \propto \exp_q \AdaBracket{\frac{1}{\tau} Q_{t, \datasetPolicy}(s,a)} = \datasetPolicy(a|s) \datasetPolicy(a|s)^{-1} \exp_q\AdaBracket{\frac{1}{\tau}Q_{t, \datasetPolicy}(s,a)} \\
+        &\pi_{t+1,\datasetPolicy} (a|s) \propto \exp_q \AdaBracket{\frac{1}{\tau} Q_{t, \datasetPolicy}(s,a)} = \datasetPolicy(a|s) \datasetPolicy(a|s)^{-1} \exp_q\AdaBracket{\frac{1}{\tau}Q_{t, \datasetPolicy}(s,a)} \\
         &= \datasetPolicy(a|s)  \exp_q{\AdaBracket{\ln_q{\frac{1}{\datasetPolicy(a|s)}}}} \exp_q\AdaBracket{\frac{1}{\tau}Q_{t, \datasetPolicy}(s,a)}\\
-        &=  \datasetPolicy(a|s) {\AdaBracket{ \exp_q\AdaBracket{\frac{1}{\tau}Q_{t, \datasetPolicy}(s,a) + \ln_2\frac{1}{\datasetPolicy(a|s)}}^{q-1} - (q-1)^2 \frac{1}{\tau}Q_{t, \datasetPolicy}(s,a)  \ln_q{\frac{1}{\datasetPolicy(a|s)}} }^{\frac{1}{q-1}}}.
+        &=  \datasetPolicy(a|s) {\AdaBracket{ \exp_q\AdaBracket{\frac{1}{\tau}Q_{t, \datasetPolicy}(s,a) + \ln_q\frac{1}{\datasetPolicy(a|s)}}^{q-1} - (q-1)^2 \frac{1}{\tau}Q_{t, \datasetPolicy}(s,a)  \ln_q{\frac{1}{\datasetPolicy(a|s)}} }^{\frac{1}{q-1}}}.
         \label{eq:tsallis_inac_policy_thoery}

### 2023-04-22 paper/main.tex
added more technical details

@@
   Intuitively, the policy first sorts actions by $a_{(1)}, \dots, a_{(|\mathcal{A}|)}$ and then compute the threshold $\psi$.
-  Suppose $Q(s, a_{(j+1)}) \leq \psi \leq Q(s, a_{(j)})$, then all actions from $j+1$ are truncated and have zero probability of being selected.
+  Suppose $Q(s, a_{(j+1)}) \leq \psi \leq Q(s, a_{(j)})$, then actions from $a_{(j+1)}, \dots, a_{(|\mathcal{A}|)}$ are truncated and have zero probability of being selected.
   The truncation effect can be controlled by $\tau$.
@@
 \begin{align}
-    & \pi_{t+1, \datasetPolicy}(a|s) \propto \exp_q\AdaBracket{\frac{Q_{t, \datasetPolicy}(s,a)}{\tau}},  \quad \sum_{a\in K_{\mathcal{D}}(s)} \pi_{t+1, \datasetPolicy} = 1.\\
+    & \pi_{t+1, \datasetPolicy}(a|s) \propto \exp_q\AdaBracket{\frac{Q_{t, \datasetPolicy}(s,a)}{\tau}},  \quad \sum_{a\in K_{\mathcal{D}}(s)} \pi_{t+1, \datasetPolicy} (a|s) = 1.\\
     Q_{t+1, \datasetPolicy}(s,a) &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \sum_{a' \in K_{\mathcal{D}}(s)} \!\!\!\!\!

### 2023-04-23 core/agent/tsallis_inac.py
added option for normalize

@@ -31,7 +31,8 @@ class TsallisInAC(base.Agent):
                  target_network_update_freq,
                  evaluation_criteria,
                  logger,
-                 q
+                 q,
+                 normalize,
                  ):
         super(TsallisInAC, self).__init__(
             exp_path=exp_path,
@@ -102,6 +103,7 @@ class TsallisInAC(base.Agent):
         self.polyak = polyak
 
         self.q = q
+        self.normalize = normalize
         self.fill_offline_data_to_buffer(offline_data)
         self.offline_param_init(offline_data)
 
@@ -163,9 +165,8 @@ class TsallisInAC(base.Agent):
         '''
         with torch.no_grad():
             beh_prob = torch.clip(torch.exp(self.beh_pi.get_logprob(states, actions)), min=self.eps)
-            # value = self.get_state_value(states)
-            # Psi = torch.sqrt(torch.clip((min_Q/self.tau)**2 - (value - 0.5*self.tau)/0.5*self.tau, min=0))
-        # x = min_Q/self.tau - Psi
+            value = self.get_state_value(states)
+            Psi = torch.sqrt(torch.clip((min_Q/self.tau)**2 - (value - self.tau)/self.tau, min=0))
         '''why we do not use Psi 
         we are assuming that actions not in the dataset are truncated 
         by Tsallis behavior policies. So when presented with these actions in the dataset,
@@ -177,7 +178,11 @@ class TsallisInAC(base.Agent):
         '''
         x = Q, y = ln_q pi_D^-1
         '''                
-        x = min_Q / self.tau
+        if self.normalize:
+            x = min_Q/self.tau - Psi
+        else:
+            x = min_Q / self.tau
+
         y = self.ac.pi.logq_x(beh_prob**(-1), self.q)
 
         tsallis_policy = torch.pow(self.ac.pi.expq_x(x + y, self.q)**(self.q-1) + (self.q - 1)**2 * x * y, 1/(self.q-1))

### 2023-04-23 run_ac_offline.py

@@ -29,7 +29,7 @@ if __name__ == '__main__':
     parser.add_argument('--evaluation_criteria', default='return', type=str)
     parser.add_argument('--device', default='cpu', type=str)
     parser.add_argument('--info', default='0', type=str)
-    # parser.add_argument('--mode', default='reverse_jensen_shannon', type=str)
+    parser.add_argument('--normalize', default=True, type=bool)
     parser.add_argument('--q', default=2.0, type=float)
     cfg = parser.parse_args()
 
@@ -49,9 +49,7 @@ if __name__ == '__main__':
     logger.log_config(cfg)
 
     # Initializing the agent and running the experiment
-    # agent_obj = InSampleAC(
     agent_obj = TsallisInAC(
-    # agent_obj = TsallisFdivInAC(
         device=cfg.device,
         discrete_control=cfg.discrete_control,
         state_dim=cfg.state_dim,
@@ -72,6 +70,6 @@ if __name__ == '__main__':
         evaluation_criteria=cfg.evaluation_criteria,
         logger=cfg.logger,
         q=cfg.q,
-        # mode=cfg.mode,
+        normalize=cfg.normalize,
     )
     run_funcs.run_steps(agent_obj, cfg.max_steps, cfg.log_interval, exp_path)

### 2023-04-23 core/construct.py
Fixing configs for new constructor. Adding new config.

@@ -38,7 +38,9 @@ def construct_TsallisInAC(config, exp_path, env_fn, offline_data, logger):
 
         tau=config["tau"],
         polyak=config["polyak"],
-        gamma=config["gamma"])
+        gamma=config["gamma"],
+        q=config["q"],
+        normalize=config["normalize"])
 
 
 def construct_InSampleAC(config, exp_path, env_fn, offline_data, logger):

### 2023-04-23 run_from_config.py

@@ -48,7 +48,7 @@ def run_experiment(config_file, job_id, base_save_dir, num_threads):
     # Save Path
     exp_path = cfg.get_save_dir_and_save_config(
         parsed.base_save_dir,
-        preformat_args=["env_name", "dataset"],
+        preformat_args=["env_name", "dataset", "agent_name"],
         postformat_args=["run"],
         arg_hash=True,
         extra_hash_ignore=["seed", "run"])
@@ -71,7 +71,6 @@ def run_experiment(config_file, job_id, base_save_dir, num_threads):
         offline_data=offline_data,
         logger=lggr)
 
-    # here we need to capture errors and still return an error if we fail (for gnu-parallel).
     try:
         run_funcs.run_steps(agent_obj,
                             cfg["max_steps"],

### 2023-04-23 run_from_config.py
Can't make change to save yet.

@@ -48,7 +48,7 @@ def run_experiment(config_file, job_id, base_save_dir, num_threads):
     # Save Path
     exp_path = cfg.get_save_dir_and_save_config(
         parsed.base_save_dir,
-        preformat_args=["env_name", "dataset", "agent_name"],
+        preformat_args=["env_name", "dataset"],
         postformat_args=["run"],
         arg_hash=True,
         extra_hash_ignore=["seed", "run"])

### 2023-04-24 core/agent/tsallis_inac.py
made normalization by value and added behavior policy clipping argument

@@ -33,6 +33,7 @@ class TsallisInAC(base.Agent):
                  logger,
                  q,
                  normalize,
+                 beh_clip,
                  ):
         super(TsallisInAC, self).__init__(
             exp_path=exp_path,
@@ -104,6 +105,7 @@ class TsallisInAC(base.Agent):
 
         self.q = q
         self.normalize = normalize
+        self.beh_clip = beh_clip
         self.fill_offline_data_to_buffer(offline_data)
         self.offline_param_init(offline_data)
 
@@ -164,9 +166,9 @@ class TsallisInAC(base.Agent):
         Psi should be learned. Estimating it using value could render policy zero everywhere
         '''
         with torch.no_grad():
-            beh_prob = torch.clip(torch.exp(self.beh_pi.get_logprob(states, actions)), min=self.eps)
+            beh_prob = torch.clip(torch.exp(self.beh_pi.get_logprob(states, actions)), min=self.beh_clip)
             value = self.get_state_value(states)
-            Psi = torch.sqrt(torch.clip((min_Q/self.tau)**2 - (value - self.tau)/self.tau, min=0))
+            # Psi = torch.sqrt(torch.clip((min_Q/self.tau)**2 - (value - self.tau)/self.tau, min=0))
         '''why we do not use Psi 
         we are assuming that actions not in the dataset are truncated 
         by Tsallis behavior policies. So when presented with these actions in the dataset,
@@ -179,7 +181,7 @@ class TsallisInAC(base.Agent):
         x = Q, y = ln_q pi_D^-1
         '''                
         if self.normalize:
-            x = min_Q/self.tau - Psi
+            x = min_Q/self.tau - value
         else:
             x = min_Q / self.tau
 
@@ -187,6 +189,7 @@ class TsallisInAC(base.Agent):
 
         tsallis_policy = torch.pow(self.ac.pi.expq_x(x + y, self.q)**(self.q-1) + (self.q - 1)**2 * x * y, 1/(self.q-1))
         clipped = torch.clip(tsallis_policy, self.eps, self.exp_threshold)
+        # print(f"x : {x}, value: {value}, y: {y}, policy: {tsallis_policy}")
         pi_loss = -(clipped * log_probs).mean()
         return pi_loss, ""
     

### 2023-04-24 run_ac_offline.py

@@ -31,9 +31,10 @@ if __name__ == '__main__':
     parser.add_argument('--info', default='0', type=str)
     parser.add_argument('--normalize', default=True, type=bool)
     parser.add_argument('--q', default=2.0, type=float)
+    parser.add_argument('--beh_clip', default=1e-8, type=float)
     cfg = parser.parse_args()
 
-    torch_utils.set_one_thread()
+    torch_utils.set_thread_count()
 
     torch_utils.random_seed(cfg.seed)
     
@@ -71,5 +72,6 @@ if __name__ == '__main__':
         logger=cfg.logger,
         q=cfg.q,
         normalize=cfg.normalize,
+        beh_clip=cfg.beh_clip,
     )
     run_funcs.run_steps(agent_obj, cfg.max_steps, cfg.log_interval, exp_path)

### 2023-04-25 analysis/data_analysis.py
Updating data_analysis code

@@ -74,9 +74,13 @@ def analyze_data(dir_name):
             row["mean_arr"] = np.mean(_data, axis=0)
             row["var_arr"] = np.var(_data, axis=0)
             row["stderr_arr"] = np.sqrt(row["var_arr"]/len(_data))
+            row["end_max"] = np.max([np.mean(d[-10:]) for d in _data])
+            row["end_min"] = np.min([np.mean(d[-10:]) for d in _data])
             row["end_mean"] = np.mean(row["mean_arr"][-10:])
             row["end_var"] = np.var([np.mean(d[-10:]) for d in _data])
             row["end_stderr"] = np.sqrt(row["end_var"]/len(_data))
+            row["all_max"] = np.max([np.mean(d) for d in _data])
+            row["all_min"] = np.min([np.mean(d) for d in _data])
             row["all_mean"] = np.mean([np.mean(d) for d in _data])
             row["all_var"] = np.var([np.mean(d) for d in _data])
             row["all_stderr"] = np.sqrt(row["all_var"]/len(_data))

### 2023-05-04 paper/main.tex
updated the paper a bit

@@
 % [preprint] option:
-\usepackage[preprint]{nips_2018}
+\usepackage{nips2023}
 
@@
 
-We focus on discrete-time discounted Markov Decision Processes (MDPs) expressed by the tuple $(\mathcal{S}, \mathcal{A}, P,  r,  \gamma)$, where $\mathcal{S}$ and $\mathcal{A}$ denote state space and finite action space, respectively.
+RL problems are usually modelled by Markov Decision Processes (MDPs) expressed by the tuple $(\mathcal{S}, \mathcal{A}, P,  r,  \gamma)$, where $\mathcal{S}$ and $\mathcal{A}$ respectively denote state space and action space.
 % $d(\cdot)$ denotes the initial state distribution.
@@
 
-\begin{figure}
-    \centering
-    \begin{subfigure}[b]{0.3\textwidth}
-        \includegraphics[width=\textwidth]{img/q_stat_illus.png}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.3\textwidth}
-        \includegraphics[width=\textwidth]{img/tsallis_entropy.png}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.3\textwidth}
-        \includegraphics[width=\textwidth]{img/tsallis_kl.png}
-    \end{subfigure}
-    \caption{(Left) Behavior of \qlog and $q$-exponential functions.
-    When $q=1$ they respectively recover the standard logarithm and exponential.
-    (Mid) Tsallis entropy of the Gaussian policy $\mathcal{N}(2, 1)$. (Right) Tsallis KL divergence between two Gaussian policies $\mathcal{N}(2.75, 1)$ and $\mathcal{N}(3.25, 1)$.
-    }
-    \label{fig:q_stats}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=0.33\textwidth]{img/Gaussian_ks.png}
+%     \caption{xs}
+% \end{figure}
+% \begin{figure}
+%     \centering
+%     \begin{subfigure}[b]{0.3\textwidth}
+%         \includegraphics[width=\textwidth]{img/q_stat_illus.png}
+%     \end{subfigure}
+%     \begin{subfigure}[b]{0.3\textwidth}
+%         \includegraphics[width=\textwidth]{img/tsallis_entropy.png}
+%     \end{subfigure}
+%     \begin{subfigure}[b]{0.3\textwidth}
+%         \includegraphics[width=\textwidth]{img/tsallis_kl.png}
+%     \end{subfigure}
+%     \caption{(Left) Behavior of \qlog and $q$-exponential functions.
+%     When $q=1$ they respectively recover the standard logarithm and exponential.
+%     (Mid) Tsallis entropy of the Gaussian policy $\mathcal{N}(2, 1)$. (Right) Tsallis KL divergence between two Gaussian policies $\mathcal{N}(2.75, 1)$ and $\mathcal{N}(3.25, 1)$.
+%     }
+%     \label{fig:q_stats}
+% \end{figure}
 
-In this paper we consider a broad class of less studied entropic regularizer known as the Tsallis entropy $S_q(\pi(\cdot|s)):= \frac{1}{q-1} \AdaBracket{\sum_a \pi^q(a|s) - 1}$, where $q \in \mathbb{R}_{+}$.
+In this paper we consider a broad class of less studied entropic regularizer known as the Tsallis entropy $S_q(\pi(\cdot|s)):= \frac{1}{q-1} \AdaBracket{1-\sum_a \pi^q(a|s)}$, where $q \in \mathbb{R}_{+}$.
 It is also called generalized entropy since it generalizes the Shannon entropy \cite{TsallisEntropy,tsallis2009introduction}.
@@
 $S_2(\pi(\cdot | s)):= \pi(a|s)\AdaBracket{1 - \pi(a|s)}$ \cite{Nachum18a-tsallis,Lee2018-TsallisRAL}. 
-The name sparse entropy comes from the fact that the regularizer leads to sparse support for the resulting policy \cite{Blondel-2020LearningFenchelYoundLoss,Martins16-sparsemax}.
+The name sparse entropy comes from the fact that the regularizer leads to sparse support for the resulting \emph{sparsemax} policy \cite{Blondel-2020LearningFenchelYoundLoss,Martins16-sparsemax}.
 When $q=\infty$, the regularizer vanishes.
-In Figure \ref{fig:q_stats} we plot the behavior of $q$-statistics under different $q$.
-
-We can write out the regularized policy for general Tsallis entropy regularization for any $q$\footnotemark\footnotetext{Different to prior works \cite{Lee2018-TsallisRAL,Nachum18a-tsallis}, our definition of $\psi$ has an additional $\frac{1}{q-1}$ term to accommodate the $q$-exponential policy, see Appendix \ref{apdx:derivation} for derivation.}:
+% In Figure \ref{fig:q_stats} we plot the behavior of $q$-statistics under different $q$.
+In this paper, we call the regularized policy sparsemax for all $q$ , and express them using the $q$-exponential\footnotemark\footnotetext{Different to prior works \cite{Lee2018-TsallisRAL,Nachum18a-tsallis}, our definition of $\psi$ has an additional $\frac{1}{q-1}$ term to accommodate the $q$-exponential policy, see Appendix \ref{apdx:derivation} for derivation.}:
 \begin{align}
@@
   Suppose $Q(s, a_{(j+1)}) \leq \psi \leq Q(s, a_{(j)})$, then actions from $a_{(j+1)}, \dots, a_{(|\mathcal{A}|)}$ are truncated and have zero probability of being selected.
-  The truncation effect can be controlled by $\tau$.
+  The actions $a_{(1)}, \dots, a_{(j)}$ are called allowable actions and collected in the set $K(s)$.
+  The degree of truncation and the size of $K(s)$ can be controlled by $\tau$. The truncation is visualized in Figure \ref{fig:simplex_truncate}.
 
   Note that for $q\neq 1,2,\infty$ the policy does not have a closed-form solution. 
-  Nonetheless we can resort to the Taylor's expansion to obtain an approximate policy, see Appendix \ref{apdx:derivation} for details. 
+  Nonetheless we can resort to the Taylor's expansion to obtain approximate sparsemax policies, see Appendix \ref{apdx:derivation} for details. 
   For those cases \eq{\ref{eq:tsallis_policy}} still holds, but may need an additional renormalization step to guarantee a valid probability distribution. 
@@
   Indeed, we can show that as $q \rightarrow \infty$, the unnormalized Taylor's expansion for the policy $\pi(a|s) \rightarrow 1, \, \forall a$, that is, after normalization it becomes a uniform distribution.
-We plot the behavior of action truncation in Figure \ref{fig:q_truncation} for two commonly used policy classes, and indeed as $q$ gets larger the truncation becomes weaker.
+  We compare argmax, softmax, sparsemax and plot the behavior of sparsemax truncation in Figure \ref{fig:simplex_truncate}.
+  We focus on Gaussian policies which are common for continuous-action algorithms.
 % \begin{figure}
@@
 
-Very recently \citet{zhu2023generalized} proposed in RL to exploit Tsallis KL divergence which generalizes Tsallis entropy.  
+Very recently \citet{zhu2023generalized} proposed in RL to exploit Tsallis KL divergence \citep{Furuichi2004-fundamentals-qKL} which generalizes Tsallis entropy.  
 They proved that the regularized optimal policy takes a similar form to the optimal KL-regularized policy:
@@
     \begin{split}
-        \qKLany{\pi(\cdot|s)}{\mu(\cdot|s)} = \sum_a \pi(a|s) \ln_{q} \frac{\pi(a|s)}{\mu(a|s)}, \quad \pi(a|s) = \mu(a|s) \exp_q\AdaBracket{\frac{Q_{\pi}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\pi}(s,a)}{\tau}} }.
+        \qKLany{\pi(\cdot|s)}{\mu(\cdot|s)} \!=\!
+-\sum_a \pi(a|s) \ln_{q} \frac{\mu(a|s)}{\pi(a|s)}, \,\, \pi(a|s) = \mu(a|s) \exp_q\AdaBracket{\frac{Q_{\pi}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\pi}(s,a)}{\tau}} \!}.
     \end{split}
@@
 \end{align}
-We can also resort to Taylor's expansion and renormalization to obtain approximate policies, hence the policy can be written as $\pi(a|s)\propto \mu(a|s)\exp_q\AdaBracket{\frac{Q_{\pi}(s,a)}{\tau}}$.
+We also write the policy as $\pi(a|s)\propto \mu(a|s)\exp_q\AdaBracket{\frac{Q_{\pi}(s,a)}{\tau}}$.
 It is worth noting that unlike the KL divergence, the Tsallis KL divergence cannot be decomposed as $\ln_q\!\frac{\pi(a|s)}{\mu(a|s)} \neq \ln_q \pi(a|s) - \ln_q \mu(a|s)$.
@@
 
-\begin{figure}[t]
+\begin{figure}
     \centering
-    \begin{subfigure}[b]{0.245\textwidth}
-        \includegraphics[width=\textwidth]{img/q2_Gaussian_sparsemax.png}
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{img/2-simplex.png}
     \end{subfigure}
-    \begin{subfigure}[b]{0.245\textwidth}
-        \includegraphics[width=\textwidth]{img/q50_scaled_Gaussian_sparsemax.png}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.245\textwidth}
-        \includegraphics[width=\textwidth]{img/q2.0_Boltzmann_sparsemax.png}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.245\textwidth}
-        \includegraphics[width=\textwidth]{img/q50_Boltzmann_sparsemax.png}
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{img/Gaussian_ks.png}
     \end{subfigure}
     \caption{
-        The sparsemax operator acting upon Gaussian and Boltzmann policies for $q=2$ and $q=50$ by truncating actions with values lower than $\psi$. 
-    As $q\rightarrow \infty$ the policy tends toward uniform distribution.
-    We assume the The dataset is generated by Tsallis policies such that the actions absent from the offline dataset are those being truncated.
+        (Left) Comparison between argmax, softmax and sparsemax on the probability simplex.
+Argmax produces a deterministic policy residing on the vertices, while a softmax policy lies inside the simplex.
+By contrast, a sparsemax policy lives on the border.
+        (Right) Sparsemax operator acting on a Gaussian policy by truncating actions with value below the threshold $\psi$.
+        Actions with value larger than $\psi$ are collected in the set $K(s)$.
     }
-    \label{fig:q_truncation}
+    \label{fig:simplex_truncate}
 \end{figure}
+
+% \begin{figure}[t]
+%     \centering
+%     \begin{subfigure}[b]{0.245\textwidth}
+%         \includegraphics[width=\textwidth]{img/q2_Gaussian_sparsemax.png}
+%     \end{subfigure}
+%     \begin{subfigure}[b]{0.245\textwidth}
+%         \includegraphics[width=\textwidth]{img/q50_scaled_Gaussian_sparsemax.png}
+%     \end{subfigure}
+%     \begin{subfigure}[b]{0.245\textwidth}
+%         \includegraphics[width=\textwidth]{img/q2.0_Boltzmann_sparsemax.png}
+%     \end{subfigure}
+%     \begin{subfigure}[b]{0.245\textwidth}
+%         \includegraphics[width=\textwidth]{img/q50_Boltzmann_sparsemax.png}
+%     \end{subfigure}
+%     \caption{
+%         The sparsemax operator acting upon Gaussian and Boltzmann policies for $q=2$ and $q=50$ by truncating actions with values lower than $\psi$. 
+%     As $q\rightarrow \infty$ the policy tends toward uniform distribution.
+%     We assume the The dataset is generated by Tsallis policies such that the actions absent from the offline dataset are those being truncated.
+%     }
+%     \label{fig:q_truncation}
+% \end{figure}
 
@@
  \end{align}
- The second equation is because $\max_{\pi}\sum_{a\in\pi(\cdot|s)}\pi(a|s)Q(s,a) - \tau S_2(\pi(\cdot|s))$ attains its maximum at $\tau \sum_{a \in K(s)}\AdaBracket{ \frac{Q(s,a)}{\tau} }^2 - \psi\AdaBracket{\frac{Q(s, \cdot)}{\tau}}^2 + \tau$ \cite{Lee2018-TsallisRAL,Martins16-sparsemax}, and the last equation is due to $K_{t, q} \preceq K_\mathcal{D}$.
+ The second equation is because $\max_{\pi}\sum_{a\in\pi(\cdot|s)}\pi(a|s)Q(s,a) + \tau S_2(\pi(\cdot|s))$ attains its maximum at $\tau \sum_{a \in K(s)}\AdaBracket{ \frac{Q(s,a)}{\tau} }^2 - \psi\AdaBracket{\frac{Q(s, \cdot)}{\tau}}^2 + \tau$ \cite{Lee2018-TsallisRAL,Martins16-sparsemax}, and the last equation is due to $K_{t, q} \preceq K_\mathcal{D}$.
  \eq{\ref{eq:tsallis_backward_q2}} states that, one needs not to explicitly enforce the support constraint since it is automatically fulfilled by learning a new subset of allowable actions.

### 2023-05-05 analysis/data_analysis.py
Adding kldiv logging

@@ -5,6 +5,7 @@ import toml
 import os
 import pathlib
 import pandas as pd
+from analysis.log2data import extract_from_single_run
 
 def get_configs_and_data(dir_name):
     param_fldrs = os.listdir(dir_name)
@@ -33,10 +34,11 @@ def compress_configs_and_data(config_and_data):
     diff_ks = set()
     for idx, k1 in enumerate(ks):
         for k2 in ks[(idx+1):]:
-            d = dict(set(config_and_data[k1]["params"].items()) ^ set(config_and_data[k2]["params"].items()))
+            d = dict(set(config_and_data[k1]["params"].items()) ^
+                     set(config_and_data[k2]["params"].items()))
             for k in d.keys():
                 diff_ks.add(k)
-    diff_dict = {k:set() for k in diff_ks}
+    diff_dict = {k: set() for k in diff_ks}
     for k, v in config_and_data.items():
         for p in diff_ks:
             diff_dict[p].add(v["params"][p])
@@ -71,6 +73,7 @@ def analyze_data(dir_name):
 
         if any(d is not None for d in data):
             _data = [d for d in data if d is not None]
+            row["arr"] = _data
             row["mean_arr"] = np.mean(_data, axis=0)
             row["var_arr"] = np.var(_data, axis=0)
             row["stderr_arr"] = np.sqrt(row["var_arr"]/len(_data))
@@ -87,3 +90,25 @@ def analyze_data(dir_name):
             
         p_a_ds.append(row)
     return pd.DataFrame(p_a_ds)
+
+def transform_best_over(df, sort_perf_by, group_by):
+    sdf = df.sort_values([sort_perf_by]).groupby(group_by).tail(1).sort_values(group_by)
+    return sdf
+
+def transform_to_json(dir_name, csv_save_name, sort_perf_by, group_by):
+    # Honestly it would be really nice to grab data and compress it.
+    # param_fldrs = os.listdir(dir_name)
+    df = analyze_data(dir_name)
+    sdf = df.sort_values([sort_perf_by]).groupby(group_by).tail(1).sort_values(group_by)
+    sdf.to_json(csv_save_name)
+
+def transform_to_csv(dir_name, csv_save_name, sort_perf_by, group_by):
+    # Honestly it would be really nice to grab data and compress it.
+    # param_fldrs = os.listdir(dir_name)
+    df = analyze_data(dir_name)
[…]

### 2023-05-05 analysis/log2data.py

@@ -1,0 +1,47 @@
+
+def extract_from_single_run(file, key, label=None, before_step=None, testset=None, xlim=[]):
+    with open(file, "r") as f:
+        content = f.readlines()
+        # print(file)
+    returns = []
+    
+    for num, l in enumerate(content):
+        if "|" in l:
+            info = l.split("|")[1].strip()
+            i_list = info.split(" ")
+            # print('i_list: ', i_list)
+            if key == "learning_rate" and "learning_rate:" in i_list:
+                returns = i_list[1]
+                return returns
+            if key == "tau" and "tau:" in i_list:
+                returns = i_list[1]
+                return returns
+            elif key == "id" and "id:" in i_list:
+                returns = i_list[1]
+                return returns
+            elif key == "train_loss" and "training" in i_list and "loss" in i_list:
+                returns.append(float(i_list[i_list.index("training")+2].split("/")[0].strip()))
+            elif key == "test_loss" and "test" in i_list and "loss" in i_list:
+                returns.append(float(i_list[i_list.index("test")+2].split("/")[0].strip()))
+            if "EVAL:" == i_list[0] or "total" == i_list[0] or "TRAIN" == i_list[0] or "TEST" == i_list[0] or "Normalized" == i_list[0]:
+                if key == "train_return" and "returns" in i_list and "TRAIN" == i_list[0]:
+                    returns.append(float(i_list[i_list.index("returns") + 1].split("/")[0].strip()))  # mean
+                elif key == "test_return" and "returns" in i_list and "TEST" == i_list[0]:
+                    returns.append(float(i_list[i_list.index("returns") + 1].split("/")[0].strip()))  # mean
+                    # returns.append(float(i_list[i_list.index("returns")+1].split("/")[1].strip())) # median
+                    # returns.append(float(i_list[i_list.index("returns")+1].split("/")[2].strip())) # min
+                    # returns.append(float(i_list[i_list.index("returns")+1].split("/")[3].strip())) # max
+                elif key == "normalized_return" and "returns" in i_list and "Normalized" == i_list[0]:
+                    returns.append(float(i_list[i_list.index("returns") + 1].split("/")[0].strip()))  # mean
+
+    # Sanity Check
+    if not isinstance(returns, int):
+        if len(returns) in [0]:  # , 1] :
+            print('Empty returns when extracting: {} '.format(key), returns)
+            print('File Name: ', file)
+            
+    if xlim!=[]:
+        return returns[xlim[0]: xlim[1]]
+    else:
+        return returns
+

### 2023-05-05 analysis/plot.py

@@ -1,0 +1,34 @@
+from analysis.log2data import extract_from_single_run
+import matplotlib.pyplot as plt
+import numpy as np
+import tol_colors as tc
+
+def plot_data(hans_dir, hans_param_setting, my_data):
+    hans_hc_y, hans_hc_error = prep_hans_data(hans_dir, hans_param_setting)
+    y, error = prep_my_data(my_data)
+
+    colorset = tc.colorsets["bright"]
+    x = range(len(y))
+
+    plt.plot(x, y, color=colorset[1], label="Tsallis")
+    plt.fill_between(x, y-error, y+error, color=colorset[1], alpha=0.4)
+
+    plt.plot(range(len(hans_hc_y)), hans_hc_y, color=colorset[2], label="InAC")
+    plt.fill_between(range(len(hans_hc_y)),
+                     hans_hc_y - hans_hc_error,
+                     hans_hc_y + hans_hc_error, color=colorset[2], alpha=0.4)
+    plt.legend()
+    plt.show()
+
+
+def prep_hans_data(dir_name, param_setting):
+    hans_hc = [extract_from_single_run(dir_name + "/" + "{}_run/{}_param_setting/log".format(r, param_setting), "normalized_return") for r in range(4)]
+    hans_smoothed_hc_y = [np.mean(np.lib.stride_tricks.sliding_window_view(d, 10), axis=1) for d in hans_hc]
+    y = np.mean(hans_smoothed_hc_y, axis=0)
+    error = 1.97*(np.sqrt(np.var(hans_smoothed_hc_y, axis=0))/4)
+    return y, error
+
+def prep_my_data(my_data):
+    smoothed_y = [np.mean(np.lib.stride_tricks.sliding_window_view(d, 10), axis=1) for d in my_data]
+    y, error = np.mean(smoothed_y, axis=0), 1.97*(np.sqrt(np.var(smoothed_y, axis=0))/4)
+    return y, error

### 2023-05-05 core/agent/base.py

@@ -319,6 +319,11 @@ class Agent:
                 pass
         return mean, median, min_, max_
 
+    def get_kl_div(self, data):
+        states, actions = data['obs'], data['act']
+        return (self.ac.pi.get_logprob(states, actions) -
+                self.beh_pi.get_logprob(states, actions)).mean()
+
     def policy(self, o, eval=False):
         o = torch_utils.tensor(self.state_normalizer(o), self.device)
         with torch.no_grad():

### 2023-05-05 core/utils/Runner.py

@@ -1,62 +1,4 @@
 
-
-# def run_steps(agent, max_steps, log_interval, eval_pth):
-#     t0 = time.time()
-#     evaluations = []
-#     agent.populate_returns(initialize=True)
-#     while True:
-#         if log_interval and not agent.total_steps % log_interval:
-#             mean, median, min_, max_ = agent.log_file(
-#                 elapsed_time=log_interval / (time.time() - t0), test=True)
-#             evaluations.append(mean)
-#             t0 = time.time()
-#         if max_steps and agent.total_steps >= max_steps:
-#             break
-#         agent.step()
-#     agent.save()
-#     np.save(Path(eval_pth, "evaluations.npy"), np.array(evaluations))
-
-# def run_experiment(config_file, job_id, base_save_dir, num_threads):
-#     # Parse Config
-#     cfg = config.Config(config_file, job_id)
-
-#     # Setup
-#     torch.use_deterministic_algorithms(True)
-#     torch_utils.set_thread_count(num_threads)
-
-#     # set seed
-#     random.seed(cfg["run"])
-#     seed = random.randint(1, 1000000000)
-#     cfg.set_seed(seed)
-#     torch_utils.random_seed(cfg["seed"])
-
-#     # Save Path
-#     exp_path = cfg.get_save_dir_and_save_config(
-#         parsed.base_save_dir,
-#         preformat_args=["env_name", "dataset"],
-#         postformat_args=["run"],
-#         arg_hash=True,
-#         extra_hash_ignore=["seed", "run"])
-#     torch_utils.ensure_dir(exp_path)
-
-#     # DataSet and Environment loading
-#     env_fn = environment.EnvFactory.create_env_fn(cfg)
-#     offline_data = run_funcs.load_testset(
-#         cfg["env_name"], cfg["dataset"], cfg["seed"])
-
-#     # Setting up the logger
-#     lggr = logger.Logger(cfg, exp_path)
-#     cfg.log(lggr)
-
-#     # Initializing the agent and running the experiment
-#     agent_obj = construct_agent(
-#         config=cfg,
-#         exp_path=exp_path,
-#         env_fn=env_fn,
-#         offline_data=offline_data,
-#         logger=lggr)
-
-
 import os
 import time
 import random
@@ -113,6 +55,7 @@ class Runner(object):
         self._initialized_from_checkpoint = False
         # misc setup
         self.evaluations = []
+        self.kldiv_eval = []
 
         # Save Path
 
@@ -120,7 +63,7 @@ class Runner(object):
 
[…]

### 2023-05-05 paper/nips2023.tex

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2023

% ready for submission
\usepackage{nips2023}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2023}

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2023}

% to avoid loading the natbib package, add option nonatbib:
%    \usepackage[nonatbib]{neurips_2023}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors

\title{Formatting Instructions For NeurIPS 2023}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{%
  David S.~Hippocampus\thanks{Use footnote for providing further information
    about author (webpage, alternative address)---\emph{not} for acknowledging
    funding agencies.} \\
  Department of Computer Science\\
  Cranberry-Lemon University\\
  Pittsburgh, PA 15213 \\
  \texttt{hippo@cs.cranberry-lemon.edu} \\
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

\begin{document}

\maketitle

\begin{abstract}
  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
  both the left- and right-hand margins. Use 10~point type, with a vertical
  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
  bold, and in point size 12. Two line spaces precede the abstract. The abstract
  must be limited to one paragraph.
\end{abstract}

\section{Submission of papers to NeurIPS 2023}

Please read the instructions below carefully and follow them faithfully. \textbf{Important:} This year the checklist will be submitted separately from the main paper in OpenReview, please review it well ahead of the submission deadline: \url{https://neurips.cc/public/guides/PaperChecklist}.

\subsection{Style}

Papers to be submitted to NeurIPS 2023 must be prepared according to the
instructions presented here. Papers may only be up to {\bf nine} pages long,
including figures. Additional pages \emph{containing only acknowledgments and
references} are allowed. Papers that exceed the page limit will not be
reviewed, or in any other way considered for presentation at the conference.

The margins in 2023 are the same as those in previous years.

Authors are required to use the NeurIPS \LaTeX{} style files obtainable at the
NeurIPS website as indicated below. Please make sure you use the current files
and not previous versions. Tweaking the style files may be grounds for
rejection.

\subsection{Retrieval of style files}

The style files for NeurIPS and other conference information are available on
the website at
\begin{center}
  \url{http://www.neurips.cc/}
\end{center}
The file \verb+neurips_2023.pdf+ contains these instructions and illustrates the
various formatting requirements your NeurIPS paper must satisfy.

The only supported style file for NeurIPS 2023 is \verb+neurips_2023.sty+,
rewritten for \LaTeXe{}.  \textbf{Previous style files for \LaTeX{} 2.09,
  Microsoft Word, and RTF are no longer supported!}

The \LaTeX{} style file contains three optional arguments: \verb+final+, which
creates a camera-ready copy, \verb+preprint+, which creates a preprint for
submission to, e.g., arXiv, and \verb+nonatbib+, which will not load the
\verb+natbib+ package for you in case of package clash.

\paragraph{Preprint option}
If you wish to post a preprint of your work online, e.g., on arXiv, using the
NeurIPS style, please use the \verb+preprint+ option. This will create a
nonanonymized version of your work with the text ``Preprint. Work in progress.''
in the footer. This version may be distributed as you see fit, as long as you do not say which conference it was submitted to. Please \textbf{do
  not} use the \verb+final+ option, which should \textbf{only} be used for
papers accepted to NeurIPS.

At submission time, please omit the \verb+final+ and \verb+preprint+
options. This will anonymize your submission and add line numbers to aid
review. Please do \emph{not} refer to these line numbers in your paper as they
will be removed during generation of camera-ready copies.

The file \verb+neurips_2023.tex+ may be used as a ``shell'' for writing your
paper. All you have to do is replace the author, title, abstract, and text of
the paper with your own.

The formatting instructions contained in these style files are summarized in
Sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.

\section{General formatting instructions}
\label{gen_inst}

The text must be confined within a rectangle 5.5~inches (33~picas) wide and
9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).  Use 10~point
type with a vertical spacing (leading) of 11~points.  Times New Roman is the
preferred typeface throughout, and will be selected for you by default.
Paragraphs are separated by \nicefrac{1}{2}~line space (5.5 points), with no
indentation.

The paper title should be 17~point, initial caps/lower case, bold, centered
between two horizontal rules. The top rule should be 4~points thick and the
bottom rule should be 1~point thick. Allow \nicefrac{1}{4}~inch space above and
below the title to rules. All pages should start at 1~inch (6~picas) from the
top of the page.

For the final version, authors' names are set in boldface, and each name is
centered above the corresponding address. The lead author's name is to be listed
first (left-most), and the co-authors' names (if different address) are set to
follow. If there is only one co-author, list both author and co-author side by
side.

Please pay special attention to the instructions in Section \ref{others}
regarding figures, tables, acknowledgments, and references.

\section{Headings: first level}
\label{headings}

All headings should be lower case (except for first word and proper nouns),
flush left, and bold.

First-level headings should be in 12-point type.

\subsection{Headings: second level}

Second-level headings should be in 10-point type.

\subsubsection{Headings: third level}

Third-level headings should be in 10-point type.

\paragraph{Paragraphs}

There is also a \verb+\paragraph+ command available, which sets the heading in
bold, flush left, and inline with the text, with the heading followed by 1\,em
of space.

\section{Citations, figures, tables, references}
\label{others}

These instructions apply to everyone.

\subsection{Citations within the text}

The \verb+natbib+ package will be loaded for you by default.  Citations may be
author/year or numeric, as long as you maintain internal consistency.  As to the
format of the references themselves, any style is acceptable as long as it is
used consistently.

The documentation for \verb+natbib+ may be found at
\begin{center}
  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
\end{center}
Of note is the command \verb+\citet+, which produces citations appropriate for
use in inline text.  For example,
\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}
produces
\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

If you wish to load the \verb+natbib+ package with options, you may add the
following before loading the \verb+neurips_2023+ package:
\begin{verbatim}
   \PassOptionsToPackage{options}{natbib}
\end{verbatim}

If \verb+natbib+ clashes with another package you load, you can add the optional
argument \verb+nonatbib+ when loading the style file:
\begin{verbatim}
   \usepackage[nonatbib]{neurips_2023}
\end{verbatim}

As submission is double blind, refer to your own published work in the third
person. That is, use ``In the previous work of Jones et al.\ [4],'' not ``In our
previous work [4].'' If you cite your other papers that are not widely available
(e.g., a journal paper under review), use anonymous author names in the
citation, e.g., an author of the form ``A.\ Anonymous'' and include a copy of the anonymized paper in the supplementary material.

\subsection{Footnotes}

Footnotes should be used sparingly.  If you do require a footnote, indicate
footnotes with a number\footnote{Sample of the first footnote.} in the
text. Place the footnotes at the bottom of the page on which they appear.
Precede the footnote with a horizontal rule of 2~inches (12~picas).

Note that footnotes are properly typeset \emph{after} punctuation
marks.\footnote{As in this example.}

\subsection{Figures}

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
\end{figure}

All artwork must be neat, clean, and legible. Lines should be dark enough for
purposes of reproduction. The figure number and caption always appear after the
figure. Place one line space before the figure caption and one line space after
the figure. The figure caption should be lower case (except for first word and
proper nouns); figures are numbered consecutively.

You may use color figures.  However, it is best for the figure captions and the
paper body to be legible if the paper is printed in either black/white or in
color.

\subsection{Tables}

All tables must be centered, neat, clean and legible.  The table number and
title always appear before the table.  See Table~\ref{sample-table}.

Place one line space before the table title, one line space after the
table title, and one line space after the table. The table title must
be lower case (except for first word and proper nouns); tables are
numbered consecutively.

Note that publication-quality tables \emph{do not contain vertical rules.} We
strongly suggest the use of the \verb+booktabs+ package, which allows for
typesetting high-quality, professional tables:
\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}
This package was used to typeset Table~\ref{sample-table}.

\begin{table}
  \caption{Sample table title}
  \label{sample-table}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Math}
Note that display math in bare TeX commands will not create correct line numbers for submission. Please use LaTeX (or AMSTeX) commands for unnumbered display math. (You really shouldn't be using \$\$ anyway; see \url{https://tex.stackexchange.com/questions/503/why-is-preferable-to} and \url{https://tex.stackexchange.com/questions/40492/what-are-the-differences-between-align-equation-and-displaymath} for more information.)

\subsection{Final instructions}

Do not change any aspects of the formatting parameters in the style files.  In
particular, do not modify the width or length of the rectangle the text should
fit into, and do not change font sizes (except perhaps in the
\textbf{References} section; see below). Please note that pages should be
numbered.

\section{Preparing PDF files}

Please prepare submission files with paper size ``US Letter,'' and not, for
example, ``A4.''

Fonts were the main cause of problems in the past years. Your PDF file must only
contain Type 1 or Embedded TrueType fonts. Here are a few instructions to
achieve this.

\begin{itemize}

\item You should directly generate PDF files using \verb+pdflatex+.

\item You can check which fonts a PDF files uses.  In Acrobat Reader, select the
  menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
  also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
  available out-of-the-box on most Linux machines.

\item \verb+xfig+ "patterned" shapes are implemented with bitmap fonts.  Use
  "solid" shapes instead.

\item The \verb+\bbold+ package almost always uses bitmap fonts.  You should use
  the equivalent AMS Fonts:
\begin{verbatim}
   \usepackage{amsfonts}
\end{verbatim}
followed by, e.g., \verb+\mathbb{R}+, \verb+\mathbb{N}+, or \verb+\mathbb{C}+
for $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.  You can also use the following
workaround for reals, natural and complex:
\begin{verbatim}
   \newcommand{\RR}{I\!\!R} %real numbers
   \newcommand{\Nat}{I\!\!N} %natural numbers
   \newcommand{\CC}{I\!\!\!\!C} %complex numbers
\end{verbatim}
Note that \verb+amsfonts+ is automatically loaded by the \verb+amssymb+ package.

\end{itemize}

If your file contains type 3 fonts or non embedded TrueType fonts, we will ask
you to fix it.

\subsection{Margins in \LaTeX{}}

Most of the margin problems come from figures positioned by hand using
\verb+\special+ or other commands. We suggest using the command
\verb+\includegraphics+ from the \verb+graphicx+ package. Always specify the
figure width as a multiple of the line width as in the example below:
\begin{verbatim}
   \usepackage[pdftex]{graphicx} ...
   \includegraphics[width=0.8\linewidth]{myfile.pdf}
\end{verbatim}
See Section 4.4 in the graphics bundle documentation
(\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})

A number of width problems arise when \LaTeX{} cannot properly hyphenate a
line. Please give LaTeX hyphenation hints using the \verb+\-+ command when
necessary.

\begin{ack}
Use unnumbered first level headings for the acknowledgments. All acknowledgments
go at the end of the paper before the list of references. Moreover, you are required to declare
funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2023/PaperInformation/FundingDisclosure}.

Do {\bf not} include this section in the anonymized submission, only in the final paper. You can use the \texttt{ack} environment provided in the style file to autmoatically hide this section in the anonymized submission.
\end{ack}

\section{Supplementary Material}

Authors may wish to optionally include extra information (complete proofs, additional experiments and plots) in the appendix. All such materials should be part of the supplemental material (submitted separately) and should NOT be included in the main submission.

\section*{References}

References follow the acknowledgments in the camera-ready paper. Use unnumbered first-level heading for
the references. Any choice of citation style is acceptable as long as you are
consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
when listing the references.
Note that the Reference section does not count towards the page limit.
\medskip

{
\small

[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
(eds.), {\it Advances in Neural Information Processing Systems 7},
pp.\ 609--616. Cambridge, MA: MIT Press.

[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
TELOS/Springer--Verlag.

[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
recall at excitatory recurrent synapses and cholinergic modulation in rat
hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2023-05-06 core/utils/Runner.py
Adding configs, tested new kldiv code.

@@ -179,7 +179,7 @@ class Runner(object):
 
         in_ = torch_utils.tensor(self.agent.state_normalizer(self.offline_data["env"]["states"]),
                                  self.agent.device)
-        self.kldiv_eval.append(self.agent.get_kl_div({"obs": in_, "act": self.offline_data["env"]["actions"]}))
+        self.kldiv_eval.append(self.agent.get_kl_div({"obs": in_, "act": self.offline_data["env"]["actions"]}).detach())
         self.evaluations.append(mean)
 
     def _run(self):

### 2023-05-07 core/utils/Runner.py
Fixed KLDiv memory issues

@@ -63,8 +63,10 @@ class Runner(object):
 
         # DataSet and Environment loading
         env_fn = environment.EnvFactory.create_env_fn(self.cfg)
-        self.offline_data = load_testset(
+        offline_data = load_testset(
             self.cfg["env_name"], self.cfg["dataset"], self.cfg["seed"])
+        self.kldiv_data = {"states": offline_data["env"]["states"],
+                           "actions": offline_data["env"]["actions"]}
 
         # Setting up the logger
         self.lggr = logger.Logger(self.cfg, self.save_path)
@@ -75,7 +77,7 @@ class Runner(object):
             config=self.cfg,
             exp_path=self.save_path,
             env_fn=env_fn,
-            offline_data=self.offline_data,
+            offline_data=offline_data,
             logger=self.lggr)
 
     def _initialize_comp_env(self, num_threads):
@@ -137,7 +139,7 @@ class Runner(object):
         self.cleanup_checkpoints()
 
     def _initialize_from_checkpoint(self,
-                                   chkpt_file):
+                                    chkpt_file):
         with open(chkpt_file, 'rb') as f:
             cp_dict = pickle.load(f)
 
@@ -146,8 +148,10 @@ class Runner(object):
         random.setstate(cp_dict["rngs"]["python"])
         np.random.set_state(cp_dict["rngs"]["numpy"])
 
-        self.offline_data = load_testset(
+        offline_data = load_testset(
             self.cfg["env_name"], self.cfg["dataset"], self.cfg["seed"])
+        self.kldiv_data = {"states": offline_data["env"]["states"],
+                           "actions": offline_data["env"]["actions"]}
         # initialize other
         self.evaluations = cp_dict["evaluations"]
         self.kldiv_eval = cp_dict["kldiv_eval"]
@@ -177,9 +181,15 @@ class Runner(object):
             elapsed_time=self.cfg.log_interval / (time.time() - t0),
             test=True)
 
-        in_ = torch_utils.tensor(self.agent.state_normalizer(self.offline_data["env"]["states"]),
-                                 self.agent.device)
-        self.kldiv_eval.append(self.agent.get_kl_div({"obs": in_, "act": self.offline_data["env"]["actions"]}).detach())
+        sze = self.kldiv_data["states"].shape[0]
+        r_idx = np.random.randint(0, sze, size=1000)
+        in_ = torch_utils.tensor(
+            self.agent.state_normalizer(self.kldiv_data["states"][r_idx, :]),
[…]

### 2023-05-07 core/utils/load_datasets.py

@@ -1,6 +1,7 @@
 import pickle
 import os
 import gym
+import d4rl # This registers all the envs to gym
 
 os.environ['D4RL_SUPPRESS_IMPORT_ERROR'] = '1'
 EARLYCUTOFF = "EarlyCutOff"
